---
title: "TEST Global-Restore-Project GOE DataViz"
author: "Wendy Anthony"
date: "2025/09/18"
output:
  html_document:
    highlight: zenburn
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries
```{r load_libraries}
library(shiny)
library(tidyverse)
library(plotly)
library(htmlwidgets)
```

# Load Data
- data from Nancy Shackelford at Global Restore Project https://globalrestoreproject.com/open_science/
```{r load-data}
OM_veg_1 <- read.csv("data/OM_vegResults1.csv", header = TRUE, sep = ",")
OM_veg_2 <- read.csv("data/OM_vegResults2.csv", header = TRUE, sep = ",")
OM_veg_3 <- read.csv("data/OM_vegResults3.csv", header = TRUE, sep = ",")
OM_veg_4 <- read.csv("data/OM_vegResults4.csv", header = TRUE, sep = ",")
OM_veg_5 <- read.csv("data/OM_vegResults5.csv", header = TRUE, sep = ",")
OM_site_data <- read.csv("data/OM_projectdata-site.csv", header = TRUE, sep = ",")
OM_treatments_data <- read.csv("data/OM_projectdata-treatments.csv", header = TRUE, sep = ",")
```

# Data Structure
## OM_veg_1
```{r structure-1}
str(OM_veg_1)
```

## OM_veg_2

### Structure OM_veg_2
```{r structure-2}
str(OM_veg_2)
```
### Unique species OM_veg_2
<details><summary><b><i>Click open to see `structure of OM_veg_2`</i></b></summary><p>
```{r unique-species-2}
str(OM_veg_2)
unique(OM_veg_2$speciesid) # 315 unique species
```
</p></details>


## OM_veg_3
```{r structure-3}
str(OM_veg_3)
```

## OM_veg_4
```{r structure-4}
str(OM_veg_4)
```

## OM_veg_5
```{r structure-5}
str(OM_veg_5)
```
## OM_veg_1_5 cbind()
```{r cbind}
OM_veg_1_5 <- rbind(OM_veg_1, OM_veg_2, OM_veg_3, OM_veg_4, OM_veg_5)
```




## Filter OM_veg_1_5
```{r filter-OM_veg_1_5}
OM_veg_1_5_t_1_19 <- OM_veg_1_5[OM_veg_1_5$treatmentid == "1_19", ] #131 ibs
# OM_veg_1_5_t_1_19 <- OM_veg_1_5[OM_veg_1_5$treatmentid == "1_19" & OM_veg_1_5$tsr !=0, ] #131 ibs
```


### str OM_veg_1_5
<details><summary><b><i>Click open to see `structure of OM_veg_1_5`</i></b></summary><p>
```{r str-OM-veg-1-5}
str(OM_veg_1_5)
```
</p></details>


### sort unique species to df 
```{r OM_veg_1_5-unique-species}
library(dplyr)

colnames(OM_veg_1_5)
OM_veg_1_5_sp_all_or <- subset(OM_veg_1_5, select = c("id", "DB", "treatmentid", "year", "tsr", "block", "replicate", "speciesid", "speciesorigin", "response", "responselevel", "responsemetric", "measurementscale", "measurementmetric"))
OM_veg_1_5_sp_all_or_sort <- OM_veg_1_5_sp_all_or %>%
                          arrange(speciesid)



OM_veg_1_5_sp_or <- subset(OM_veg_1_5, select = c("speciesid", "speciesorigin"))
OM_veg_1_5_sp_or_sort <- OM_veg_1_5_sp_or %>%
                          arrange(speciesid)
# str(OM_veg_1_5_sp_or_sort)

# list of unique species values
OM_veg_1_5_sp_or_sort_unique <- unique(OM_veg_1_5_sp_or_sort$speciesid)
```


```{r save-OM_veg_1_5_sp_or_sort-csv}
write.csv(OM_veg_1_5_sp_all_or_sort, "data/OM_veg_1_5_sp_all_or_sort.csv", row.names = FALSE)
```



### add speciesorigin to speciesid for OM_veg_2
<details><summary><b><i>Click open to see `list of species and origin OM_veg_2`</i></b></summary><p>
```{r}
OM_veg_2
```
</p></details>

### Order and sort OM_veg_2
<details><summary><b><i>Click open to see `sort order of OM_veg_2`</i></b></summary><p>
```{r order-sort-OM_veg_2}
OM_veg_1_5_sp_or_sort
```
</p></details>

### OM_veg_2 percent species changed from total species
```{r percent-OM_veg_2}
length(unique(OM_veg_2$speciesid)) # 316
length(OM_veg_2$speciesid) #15,230 records
length(OM_veg_1_5_sp_or_sort$speciesid) # 136364 records

length(OM_veg_2$speciesid) / length(OM_veg_1_5_sp_or_sort$speciesid) # 11.17%
```


## OM_veg_1_5_sp_or_sort_unique
<details><summary><b><i>Click open to see results for structure of `OM_veg_1_5_sp_or_sort_unique`</i></b></summary><p>

```{r OM_veg_1_5-unique-species_str}
OM_veg_1_5_sp_or_sort_unique
```
</p></details>



## OM_treatments_data
```{r structure-OM-treatments-data}
str(OM_treatments_data)
colnames(OM_treatments_data)
unique(OM_treatments_data$restorationtype)
unique(OM_treatments_data$treatment_type)
unique(OM_treatments_data$treatment_category)
```
### Treatments fire
```{r treatments-fire}
OM_treatments_data_fire <- OM_treatments_data[OM_treatments_data$treatment_category == "fire", ]
OM_treatments_data_fire
```

### Treatments mowing
```{r treatments-mowing}
OM_treatments_data_mowing <- OM_treatments_data[OM_treatments_data$treatment_type == "mowing", ]
OM_treatments_data_mowing
```

### Treatments herbicide
```{r treatments-herbicide}
OM_treatments_data_herbicide <- OM_treatments_data[OM_treatments_data$treatment_category == "herbicide", ]
OM_treatments_data_herbicide
```



## OM_site_data
```{r structure-OM_site_data}
str(OM_site_data)
```

### Round Temp values to 1 decimal OM_site_data
```{r round-temp}
OM_site_data$temp
OM_site_data$temp <- round(OM_site_data$temp, digits = 1)
# OM_site_data$temp
```

### Round Aspect values to 2 decimal OM_site_data
```{r round-aspect}
OM_site_data$aspect
OM_site_data$aspect <- round(OM_site_data$aspect, digits = 2)
# OM_site_data$aspect
```

### Sites: Choose Columns for Site `OM_site_data`
Note: `OM_projectdata-site.csv` was showing 1709 of empty rows at end of df that shows as NA    
- deleted empty rows from bottom of `OM_projectdata-2025-Ago-25.xlsx` before saving to csv file

```{r site-columns-table}
# OM_site_data <- read.csv("data/OM_projectdata-site.csv", header = TRUE, sep = ",")
str(OM_site_data)
colnames(OM_site_data)
# Removes to leave chosen ones to make new dataframe
# keep [2] "siteid" [3] "projectid" [4]"sitename" [5] "Subregion" [6] "latitude" [7] "longitude" [8] "refecosystem" [11] "landcover" [19] "elevation" [20] "aspect" [21]  "slope" [25] "tempc"  [36] "precip" 
GRP_Site_data <- OM_site_data[-c(1,9:11,13:19,23:25,27:36,38:49)]
```
### GRP_Site_data
<details><summary><b><i>Click open to see `GRP_Site_data`</i></b></summary><p>
```{r}
GRP_Site_data
```
</p></details>




### Sites: Save chosen columns as csv to input into Shiny drop down list of Sites
```{r save-sites-csv}
write.csv(GRP_Site_data, "data/GRP_Site_data.csv", row.names = FALSE)
```




# SQL Create Database
```{r libraries-created-db}
# https://www.youtube.com/watch?v=c3e8m3K0eC4
# 
# install.packages("RSQL")
# install.packages("RSQLite")
# install.packages("tidyverse") # v.1.3.2 to 2.0.0
library(RSQL)
library(RSQLite)
library(tidyverse)
# tidyverse_packages(include_self = TRUE)
```

## SQL  GRP Database
### create GRP database and connection
```{r create-database-connection-grp}
# builds empty db in R memory
con <- dbConnect(drv = RSQLite::SQLite(),
                 dbname = ":memory:")
dbListTables(con)
```

### Populate GRP tables
- dbWriteTable
```{r dbWriteTable}
# Load OM_veg_1_5 table
dbWriteTable(conn = con,
             name = "OM_veg_1_5",
             value = OM_veg_1_5)
str(OM_veg_1_5)

# Load OM_site_data table
dbWriteTable(conn = con,
             name = "OM_site_data",
             value = OM_site_data)
str(OM_site_data)

# Load OM_treatments_data table
dbWriteTable(conn = con,
             name = "OM_treatments_data",
             value = OM_treatments_data)
str(OM_treatments_data)

dbListTables(con)
```

## Preview tables
```{r preview-tables}
tbl(src = con, "OM_site_data")
tbl(src = con, "OM_treatments_data")
tbl(src = con, "OM_veg_1_5")
```


  ## SQL Query for db


### 1-Results query one tables
```{sql connection = con, output.var = "results_projid"}
SELECT *
FROM OM_site_data
WHERE projectid = 1
```

#### View query results

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid}
results_projid
```
</p></details>

### 2-Results query two tables
```{sql connection = con, output.var = "results_projid_2"}
SELECT s.siteid, t.treatmentid, s.Site
FROM OM_site_data AS s 
INNER JOIN OM_treatments_data AS t ON s.siteid = t.siteid
WHERE projectid = 1
```

#### View query results
<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_2}
results_projid_2
```
</p></details>

### 3-Results query three tables - one species & 1 year
- SELECT DISTINCT for no duplicates
```{sql connection = con, output.var = "results_projid_3"}
SELECT DISTINCT s.siteid, t.treatmentid, s.Site, v.year, v.speciesid
FROM OM_site_data AS s 
INNER JOIN OM_treatments_data AS t ON s.siteid = t.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE speciesid = "Ant_odo" AND year = 2008
```

#### View query results
<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_3}
results_projid_3
```
</p></details>

### 3-Results query three tables - one Site
- SELECT DISTINCT for no duplicates
```{sql connection = con, output.var = "results_projid_4"}
SELECT DISTINCT s.siteid, s.Site, s.Subregion, t.treatmentid, v.year, v.speciesid, v.tsr, v.replicate, v.response
FROM OM_site_data AS s 
INNER JOIN OM_treatments_data AS t ON s.siteid = t.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE Site = "Christmas Hill"
```

#### View query results
<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_4}
results_projid_4
```
</p></details>

### 3-Results query three tables - all Sites
- SELECT DISTINCT for no duplicates
- should be 196 sites - not showing ones with same Sitename but different siteid
- include t.siteid to see that the correspondence works
- should be 196 sites (query results 149)
- WHERE s.DB = "OM"
```{sql connection = con, output.var = "results_projid_5"}
SELECT DISTINCT s.siteid, s.projectid, s.Site, s.Subregion, s.temp, s.precip, s.latitude, s.longitude, s.refecosystem, s.landcover, s.elevation, s.aspect
FROM OM_site_data AS s 
INNER JOIN OM_treatments_data AS t ON s.siteid = t.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE s.siteid BETWEEN 1 AND 196
```

#### View query results
- this doesn't have all sites, and some duplicate siteid  

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_5}
results_projid_5
```
</p></details>

#### Save query as csv file with all Site inof
```{r create-csv-results_projid_5}
write.csv(results_projid_5, "data/GRP_all_site_data.csv", row.names = FALSE)
```


### 3-Results query three tables - all Sites - 
```{sql connection = con, output.var = "results_projid_6"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, s.latitude, s.longitude, s.refecosystem, s.landcover, s.elevation, s.aspect, s.slope, s.temp, s.precip, s.disturbance
FROM OM_site_data AS s 
WHERE s.DB = "OM"
```

#### View query results

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_6}
results_projid_6
length(unique(results_projid_6$Site)) # 201 with 5 duplicates
```
</p></details>

#### Save query as csv file with all 206 Sites
```{r create-csv-results_projid_6}
write.csv(results_projid_6, "data/GRP_all_site_data_sql.csv", row.names = FALSE)
```



### 7 Query Results Treatment + Sites
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576  
```{sql connection = con, output.var = "results_projid_7"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, t.treatmentid, t.restorationtype, t.tsr_start_year, t.trt_year, t.treatmentmonth, t.treatment_category, t.treatment_type
FROM OM_treatments_data AS t 
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE s.DB = "OM"
```

#### View query results
- Query Results Treatment + Sites

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_7}
results_projid_7
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576   ... too large to view in Excel
```{r create-csv-results_projid_7}
write.csv(results_projid_7, "data/GRP-Site-Treatment-SQL.csv", row.names = FALSE)
```

### 8 Query Results Treatment + Sites
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576 
- WHERE s.Subregion = "US-GI" >>> change Subregion and save csv for each  
- US-GI, US-Oreg, US-Wash, C-GI, C-EVI, C-EVII, C-SP
- add year from v.year as some treatments done include year start end  

```{sql connection = con, output.var = "results_projid_8"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, t.treatmentid, t.restorationtype, t.tsr_start_year, t.trt_year, t.treatmentmonth, t.treatment_category, t.treatment_type
FROM OM_treatments_data AS t 
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE s.Subregion = "C-SP"
```

#### View query results
- Query Results Treatment + Sites

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_8}
results_projid_8
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites
```{r create-csv-results_projid_8}
write.csv(results_projid_8, "data/GRP-Site-Treatment-SQL-C-SP.csv", row.names = FALSE)
```


### 9 Query Results Species + Sites 
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576 
- WHERE s.Subregion = "US-GI" >>> change Subregion and save csv for each  
- US-GI, US-Oreg, US-Wash, C-GI, C-EVI, C-EVII, C-SP

```{sql connection = con, output.var = "results_projid_9"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, v.treatmentid, v.year AS "v_year_measure", v.replicate, v.speciesid, v.speciesorigin, v.response, v.responselevel, v.responsemetric, v.measurementscale, v.measurementmetric
FROM OM_veg_1_5 AS v
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_treatments_data AS t  ON v.treatmentid = t.treatmentid
WHERE s.Subregion = "US-Wash"
```

#### View query results
- Query Results Treatment + Sites

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_9}
results_projid_9
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites
```{r create-csv-results_projid_9}
write.csv(results_projid_9, "data/GRP-Site-SpeciesCover-SQL-US-Wash.csv", row.names = FALSE)
```

### 10 Query Results Species + Sites 
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576 
- WHERE s.Subregion = "US-GI" >>> change Subregion and save csv for each  
- US-GI, US-Oreg, US-Wash, C-GI, C-EVI, C-EVII, C-SP

```{sql connection = con, output.var = "site_species_C_GI"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, v.treatmentid, v.year AS "v_year_measure", v.replicate, v.speciesid, v.speciesorigin, v.response, v.responselevel, v.responsemetric, v.measurementscale, v.measurementmetric
FROM OM_veg_1_5 AS v
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_treatments_data AS t  ON v.treatmentid = t.treatmentid
WHERE s.Subregion = "C-GI"
```

#### View query results
- Query Results Treatment + Sites

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-site-species-C-GI}
site_species_C_GI
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites
```{r create-csv-results-site-species-C-GI}
write.csv(site_species_C_GI, "data/GRP-Site-SpeciesCover-SQL-C-GI.csv", row.names = FALSE)
```


### 11 Query Results Species + Sites 
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576 
- WHERE s.Subregion = "US-GI" >>> change Subregion and save csv for each  
- US-GI, US-Oreg, US-Wash, C-GI, C-EVI, C-EVII, C-SP

```{sql connection = con, output.var = "site_species_C_SP"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, v.treatmentid, v.year AS "v_year_measure", v.replicate, v.speciesid, v.speciesorigin, v.response, v.responselevel, v.responsemetric, v.measurementscale, v.measurementmetric
FROM OM_veg_1_5 AS v
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_treatments_data AS t  ON v.treatmentid = t.treatmentid
WHERE s.Subregion = "C-SP"
```

#### View query results
- Query Results Treatment + Sites

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-site-species-C-SP}
site_species_C_SP
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites
```{r create-csv-results-site-species-C-SP}
write.csv(site_species_C_SP, "data/GRP-Site-SpeciesCover-SQL-C-SP.csv", row.names = FALSE)
```

### 12 Query Results Species Specific Sites Mt. Tolmie
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576 
- WHERE s.Subregion = "US-GI" >>> change Subregion and save csv for each  
- US-GI, US-Oreg, US-Wash, C-GI, C-EVI, C-EVII, C-SP

```{sql connection = con, output.var = "species_site_62"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, v.treatmentid, v.year AS "v_year_measure", v.replicate, v.speciesid, v.speciesorigin, v.response, v.responselevel, v.responsemetric, v.measurementscale, v.measurementmetric
FROM OM_veg_1_5 AS v
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_treatments_data AS t  ON v.treatmentid = t.treatmentid
WHERE s.siteid = 62
```

#### View query results
- Query Results Treatment + Sites

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-species-siteid-62}
species_site_62
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites
```{r create-csv-results-species-siteid-62}
write.csv(species_site_62, "data/GRP-Site-SpeciesCover-SQL-species-site-62.csv", row.names = FALSE)
```


#### Warning message:
```call dbDisconnect() when finished working with a connection``` 




#### Mt Tolmie Species Data
```{r mt-tolmie-species-data}
colnames(species_site_62)
str(species_site_62)
species_site_62_native <- species_site_62[species_site_62$speciesorigin == "native", ]
species_site_62_exotic <- species_site_62[species_site_62$speciesorigin == "exotic", ]

species_site_62_native_pres <- subset(species_site_62, (speciesorigin == 'native' & responsemetric == "presence"))

species_site_62_native_cov <- subset(species_site_62, (speciesorigin == 'native' & responsemetric == "cover"))

species_site_62_exotic_pres <- subset(species_site_62, (speciesorigin == 'exotic' & responsemetric == "presence"))

species_site_62_exotic_cov <- subset(species_site_62, (speciesorigin == 'exotic' & responsemetric == "cover"))

```

#### Mt Tolmie Cover Data
```{r mt-tolmie-cover-data}
species_site_62_by_replicate <- species_site_62 %>% group_by(replicate)
```

#### Mt Tolmie number Native & Exotic Species
```{r mt-tolmie-native}
nat_num_62 <- length(species_site_62_native_pres$speciesid) #45
nat_num_62
exo_num_62 <- length(species_site_62_exotic_pres$speciesid) #71
exo_num_62

tot_sp_62 <- nat_num_62 + exo_num_62
tot_sp_62

# percentage native species to total 
nat_num_62/tot_sp_62 * 100 # 38.7931

# percentage exotic species to total 
exo_num_62/tot_sp_62 * 100 # 61.2069

```

#### Mt Tolmie Exotic Species
```{r mt-tolmie-exotic}
species_site_62_by_replicate <- species_site_62 %>% group_by(replicate)
```

#### Save queried as csv file 
- Query Mt Tolmie species cover data
```{r create-csv-results-mt-tolmie-cover-data}
write.csv(species_site_62_by_replicate, "data/GRP-Site-SpeciesCover-SQL-species-site-62-cover-data.csv", row.names = FALSE)

write.csv(species_site_62_native_pres, "data/GRP-Site-SpeciesCover-SQL-species-site-62-cover-data-native-pres.csv", row.names = FALSE)
write.csv(species_site_62_native_cov, "data/GRP-Site-SpeciesCover-SQL-species-site-62-cover-data-native-cov.csv", row.names = FALSE)
write.csv(species_site_62_exotic_pres, "data/GRP-Site-SpeciesCover-SQL-species-site-62-cover-data-exotic-pres.csv", row.names = FALSE)
write.csv(species_site_62_exotic_cov, "data/GRP-Site-SpeciesCover-SQL-species-site-62-cover-data-exotic-cov.csv", row.names = FALSE)
```


## SQL query function
```{r SQL-query-function}
# Load some sample data (optional, for demonstration)
  # In a real scenario, you would connect to an existing database with your data

# Error: Table OM_veg_1_5 exists in database, and both overwrite and append are FALSE
  # Load OM_veg_1_5 table
# dbWriteTable(conn = con,
#              name = "OM_veg_1_5",
#              value = OM_veg_1_5)
# 
# 
# # Load OM_site_data table
# dbWriteTable(conn = con,
#              name = "OM_site_data",
#              value = OM_site_data)
# 
# 
# # Load OM_treatments_data table
# dbWriteTable(conn = con,
#              name = "OM_treatments_data",
#              value = OM_treatments_data)
# get_data_by_site <- function(con, Site) {
#   
#   sql_query_site <- paste0('SELECT s.siteid, s.projectid, s.Site, s.Subregion, v.treatmentid, v.year AS "v_year_measure", v.replicate, v.speciesid, v.speciesorigin, v.response, v.responselevel, v.responsemetric, v.measurementscale, v.measurementmetric
# FROM OM_veg_1_5 AS v
# INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
# INNER JOIN OM_treatments_data AS t  ON v.treatmentid = t.treatmentid
# WHERE s.siteid = "target_site"')
# 
#   # Execute the query and return the result as a data frame
#   dbGetQuery(con, sql_query_site)
# }
# 
#   # Establish a database connection.
#   # Example: Connect to an in-memory SQLite database
#   con <- dbConnect(RSQLite::SQLite(), ":memory:")
# 
#   # Error: Table OM_veg_1_5 exists in database, and both overwrite and append are FALSE
#   # Load OM_veg_1_5 table
# dbWriteTable(conn = con,
#              name = "OM_veg_1_5",
#              value = OM_veg_1_5)
# 
# 
# # Load OM_site_data table
# dbWriteTable(conn = con,
#              name = "OM_site_data",
#              value = OM_site_data)
# 
# 
# # Load OM_treatments_data table
# dbWriteTable(conn = con,
#              name = "OM_treatments_data",
#              value = OM_treatments_data)
# 
# # Call the function to get sales data for 2022
#   site_MtTolmie <- get_sales_by_year(con, "	
# Mt. Tolmie")
#   print(site_MtTolmie)


```

# For Loop
## for loop prep
- create a for loop for separate Site df  

<details><summary><b><i>Click open to see `data prep`</i></b></summary><p>
```{r for-loop-prep}
site_species_C_SP
length(site_species_C_SP$Site)
length(results_projid_6$Site) # 206
length(unique(results_projid_6$Subregion)) # 7

colnames(results_projid_6)
sites_C_SP <- subset(results_projid_6, Subregion == "C-SP")
# same as 
sites_C_SP <- results_projid_6[results_projid_6$Subregion == "C-SP", ]
sites_C_SP
sites_C_SP$Site

unique(sites_C_SP$Site) # 86 s/b 88

# sort alphabelticaly
sites_C_SP_sorted <- sites_C_SP[order(sites_C_SP$Site), ]
sites_C_SP_sorted

site_species_C_SP_sorted <- site_species_C_SP[order(site_species_C_SP$Site), ]
site_species_C_SP_sorted
```
</p></details>

##  Separate df for each Site of C_SP
<details><summary><b><i>Click open to see `for loop for separate sites`</i></b></summary><p>
```{r site-df-C_SP}

# sites_C_SP_sorted
site_species_C_SP_sorted

# create list of data frames for each C_SP Site
site_species_C_SP_df <- split( site_species_C_SP_sorted, site_species_C_SP_sorted$Site )
site_species_C_SP_MtTolmie <- site_species_C_SP_df$`Mt. Tolmie`

str(site_species_C_SP_df)

### WORKS !!!
# for (i in 1:length(sites_df)) {
#   assign(paste0("sites_df", i), as.data.frame(sites_df[[i]]))
# }
for (i in 1:length(site_species_C_SP_df)) {
  assign(paste0("site_species_C_SP_df", i, sep="_"), as.data.frame(site_species_C_SP_df[[i]]))
}

# sites_df[["Albert Head"]][["Site"]]

# EXTRACT LIST OF names of list dataframes
sites_df_namelist <- names(site_species_C_SP_df) # 76 
sites_df_namelist


# desired name derived from first 
desired_name_value <- site_species_C_SP_df$Site[1] # Taking the first value as an example
# this works to rename 1 df
site_species_C_SP_df_test <- site_species_C_SP_df
assign(paste("site_species_C_SP_df", desired_name_value, sep = "_"), site_species_C_SP_df_test)

```

</p></details>

### Rename dataframes
?? how can I do a for loop to rename all these  
??? Take care ... I crashed and corrupted this files predecessor after t
```{r rename-df}

# site_species_C_SP_df1_ <- assign(paste("site_species_C_SP_df1_", site_species_C_SP_df1_$Site, sep = ""), site_species_C_SP_df1_)
# 
# site_species_C_SP_df2_ <- assign(paste("site_species_C_SP_df2_", site_species_C_SP_df2_$Site, sep = ""), site_species_C_SP_df2_)
# 
# site_species_C_SP_df10_ <- assign(paste("site_species_C_SP_df10_", site_species_C_SP_df10_$Site, sep = ""), site_species_C_SP_df10_)
```



# Plots




## Treatments 
### Treatments Fire & Seeding

```{r treatments-fire-seed-plot}
colnames(OM_treatments_data_fire)
OM_treatments_data_fire_seed <- OM_treatments_data_fire[OM_treatments_data_fire$restorationtype == "seeding", ] 
unique(OM_treatments_data_fire_seed$siteid) #10 129:138

# display.brewer.all(n=10, exact.n=FALSE)

OM_treatments_data_fire_seed_plot <- ggplot(OM_treatments_data_fire_seed, 
                      # text = paste("Site:", Site), 
                    aes(x = factor(siteid), y = trt_year, color = treatmentid)) +
                    geom_point(shape = 20, size = 1) +
                    scale_y_continuous(breaks = c(2006, 2008)) +
                    geom_jitter() +
    #scale_x_continuous(limits = c(500, 2000)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_color_brewer(palette = "Accent") +
  
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="none") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Treatments with Fire and Seeding",
                         subtitle = "(First Treatment 2005)",
                      caption = "Chart by Wendy Anthony \n 2025-09-25",
                      x = "Site", y = "Treatment Year")

OM_treatments_data_fire_seed_plot
## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), OM_treatments_data_fire_seed_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Treatments with Fire and Seeding (First Treatment 2005)<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Site'),
        yaxis = list(title = 'Treatment Year'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-25"
              # no legend
             #               ,
             # showarrow = F, xref='paper', yref='paper',
             # xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             # font = list(size=8, color="grey")
             )
        )

```


### Treatments Mowing & Seeding

```{r treatments-mowing-seed-plot}
#colnames(OM_treatments_data_mowing)
#unique(OM_treatments_data_mowing$trt_year) 
# 2005 2006 2007 2008 2010 2011 2012 2013 2014 2015 2016 2009 2017 2018 2019 2020 2021 2022
#unique(OM_treatments_data_mowing$tsr_start_year) # 2005 2010 2009

OM_treatments_data_mowing_seed <- OM_treatments_data_mowing[OM_treatments_data_mowing$restorationtype == "seeding", ] 
unique(OM_treatments_data_mowing_seed$siteid) #10 129:138

# display.brewer.all(n=10, exact.n=FALSE)

OM_treatments_data_mowing_seed_plot <- ggplot(OM_treatments_data_mowing_seed, 
                      # text = paste("Site:", Site), 
                      
                    aes(x = factor(siteid), y = trt_year, color = treatmentid)) +
                    geom_point(shape = 20, size = 1) +
                    geom_jitter() +
#    scale_x_continuous(limits = c(129, 138)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_color_brewer(palette = "Accent") +
  
                    theme_minimal() + #get rid of grey background and tick marks
                    #theme(legend.position="bottom") +
                     theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Treatments with Mowing and Seeding",
                         subtitle = "(First Treatment 2005)",
                      caption = "Chart by Wendy Anthony \n 2025-09-25",
                      x = "Site", y = "Treatment Year")


## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), OM_treatments_data_mowing_seed_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Treatments with Mowing and Seeding (First Treatment 2005)<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Site'),
        yaxis = list(title = 'Treatment Year'),
        
        
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-25"              # no legend
             #               ,
             # showarrow = F, xref='paper', yref='paper',
             # xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             # font = list(size=8, color="grey")
             )
        ) 

```


### Treatments Herbicide

```{r treatments-herbicide-plot}
#colnames(OM_treatments_data_mowing)
#unique(OM_treatments_data_mowing$trt_year) 
# 2005 2006 2007 2008 2010 2011 2012 2013 2014 2015 2016 2009 2017 2018 2019 2020 2021 2022
#unique(OM_treatments_data_mowing$tsr_start_year) # 2005 2010 2009

OM_treatments_data_herbicide_seed <- OM_treatments_data_herbicide[OM_treatments_data_herbicide$restorationtype == "seeding", ] 
unique(OM_treatments_data_herbicide_seed$siteid) #10 129:138

OM_treatments_data_herbicide_seed_plot <- ggplot(OM_treatments_data_herbicide_seed, 
                      # text = paste("Site:", Site), 
                      
                    aes(x = factor(siteid), y = trt_year, color = treatmentid)) +
                    geom_point(shape = 20, size = 1) +
                    geom_jitter() +
#    scale_x_continuous(limits = c(129, 138)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_color_brewer(palette = "Accent") +
  
                    theme_minimal() + #get rid of grey background and tick marks
                    #theme(legend.position="bottom") +
                     theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Treatments with Herbicide and Seeding",
                         subtitle = "(First Treatment 2005)",
                      caption = "Chart by Wendy Anthony \n 2025-09-25",
                      x = "Site", y = "Treatment Year")


## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), OM_treatments_data_mowing_seed_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Treatments with Herbicide and Seeding (First Treatment 2005)<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Site'),
        yaxis = list(title = 'Treatment Year'),
        
        
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-25"              # no legend
             #               ,
             # showarrow = F, xref='paper', yref='paper',
             # xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             # font = list(size=8, color="grey")
             )
        ) 

```

## Template ggplot geom_point with plotly
```{r template-ggplot-geom_point-plotly}
OM_treatments_data_fire_seed_plot <- ggplot(OM_treatments_data_fire_seed, 
                      # text = paste("Site:", Site), 
                    aes(x = factor(siteid), y = trt_year, color = treatmentid)) +
                    geom_point(shape = 20, size = 1) +
                    scale_y_continuous(breaks = c(2006, 2008)) +
                    geom_jitter() +
    #scale_x_continuous(limits = c(500, 2000)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_color_brewer(palette = "Accent") +
  
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="none") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Treatments with Fire and Seeding",
                         subtitle = "(First Treatment 2005)",
                      caption = "Chart by Wendy Anthony \n 2025-09-25",
                      x = "Site", y = "Treatment Year")

# OM_treatments_data_fire_seed_plot
## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), OM_treatments_data_fire_seed_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Treatments with Fire and Seeding (First Treatment 2005)<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Site'),
        yaxis = list(title = 'Treatment Year'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-25"
              # no legend
             #               ,
             # showarrow = F, xref='paper', yref='paper',
             # xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             # font = list(size=8, color="grey")
             )
        )

```

## Sites 129-138
- these sites had treatment done  

<details><summary><b><i>Click open to see `species data for sites 129-138`</i></b></summary><p>
```{r sites-species-129-138}
# [1] 128 129 134 135 136 137 138 130 131 132 133
# https://stackoverflow.com/questions/25647470/filter-multiple-values-on-a-string-column-in-dplyr
target_sites <- c("129", "130", "131", "132", "133", "134", "135", "136", "137", "138")
# filter
sites_species_129_138 <- dplyr::filter(results_projid_9, siteid %in% target_sites)
sites_species_129_138
```
</p></details>


## Filter Single Site Species Cover & Presence
- Mt Tolmie siteid = 62  
- can't use pivot longer
```{r plot-site-species-cover-presence-MtTolmie}
colnames(species_site_62)
# remove columns for Site 
ss_cover_62 <- species_site_62[, -c(2,3,4)]

# filter
ss_cover_62_presence <- ss_cover_62[ss_cover_62$responsemetric == "presence", ]
ss_cover_62_cover <- ss_cover_62[ss_cover_62$responsemetric == "cover", ]

ss_cover_62_cover_native <- ss_cover_62_cover[ss_cover_62_cover$speciesorigin == "native", ]

ss_cover_62_cover_exotic <- ss_cover_62_cover[ss_cover_62_cover$speciesorigin == "exotic", ]

ss_cover_62_cover_native_response_order <- ss_cover_62_cover_native[order(ss_cover_62_cover_native$response), ]

ss_cover_62_cover_exotic_response_order <- ss_cover_62_cover_exotic[order(ss_cover_62_cover_exotic$response), ]
```


### plot Mt. Tolmie
- not showing all species - includes multiple replicates/species


### plot Mt. Tolmie native vs exotic species
- not showing all species
### Mt Tolmie Native 
```{r ggplot-ss_cover_62_cover_native_response_order}
ss_cover_62_cover_native_response_order_plot <- ggplot(ss_cover_62_cover_native_response_order, 
                      # text = paste("Site:", Site), 
                    aes(y = speciesid, x = response, color = replicate)) +
                    geom_point(shape = 20, size = 1) +
                    geom_jitter() +
    #scale_x_continuous(limits = c(500, 2000)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_fill_brewer(palette = "Dark2") +
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="bottom") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Mt. Tolmie Native Species Cover %",
                      caption = "Chart by Wendy Anthony \n 2025-09-25",
                      y = "Species", x = "% Cover")

## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), ss_cover_62_cover_native_response_order_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Mt. Tolmie Native Plant % Cover<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Cover (%)'),
        yaxis = list(title = 'Species'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-25",
             showarrow = F, xref='paper', yref='paper',
             xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             font = list(size=8, color="grey"))
        )
```

### Mt Tolmie Exotic 
```{r ggplot-ss_cover_62_cover_exotic_response_order}
ss_cover_62_cover_exotic_response_order_plot <- ggplot(ss_cover_62_cover_exotic_response_order, 
                      # text = paste("Site:", Site), 
                    aes(y = speciesid, x = response, color = replicate)) +
                    geom_point(shape = 20, size = 1) +
                    geom_jitter() +
    #scale_x_continuous(limits = c(500, 2000)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_fill_brewer(palette = "Dark2") +
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="bottom") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Mt. Tolmie Exotic Species Cover %",
                      caption = "Chart by Wendy Anthony \n 2025-09-25",
                      y = "Species", x = "% Cover")

## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), ss_cover_62_cover_exotic_response_order_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Mt. Tolmie Exotic Plant % Cover<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Cover (%)'),
        yaxis = list(title = 'Species'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-25",
             showarrow = F, xref='paper', yref='paper',
             xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             font = list(size=8, color="grey"))
        )
```



## 1-point plot - Temp & Precip
- I think this needs to be first use average of subregion temp and precip - showing too many here ???
```{r GRP-temp-precip-all-point}

GRP_all_site_temp_prec <- ggplot(results_projid_6, 
                      # text = paste("Site:", Site), 
                    aes(x = precip, y = temp, colour = Subregion)) +
                    geom_point(shape = 18, size = 2) +
    scale_x_continuous(limits = c(500, 2000)) +
    scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    scale_fill_brewer(palette = "Dark2") +
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="bottom") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "GRP Precipitation & Temperature",
                       subtitle = "By Subregion",
                      caption = "Chart by Wendy Anthony \n 2025-09-21",
                      x = "Precipitation", y = "Temperature")

## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), GRP_all_site_temp_prec) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "GRP Annual Precipitation & Temperature by Subregion<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Precipitation (mm)'),
        yaxis = list(title = 'Temperature (°C)'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-21",
             showarrow = F, xref='paper', yref='paper',
             xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             font = list(size=8, color="grey"))
        )
```

## 2-point plot - Temp & Precip
- I think this needs to be first use mean of subregion temp and precip - showing too many here ???

### mean temp-subregion
```{r GRP-temp-subregion-mean}
results_projid_6_test <- results_projid_6
colnames(results_projid_6_test)

subregion_mean_temp <- aggregate(results_projid_6_test[, 12], list(results_projid_6_test$Subregion), mean)
subregion_mean_temp # no results for C-GI, C-SP, US-GI

# doesn't work
#results_projid_6 %>% group_by(Subregion) %>% mutate(mean_temp=mean(temp))
```

### mean precip-subregion
- results_projid_6  all 206 site data
```{r GRP-precip-subregion-mean}
results_projid_6
```

### 2-point plot - Temp & Precip
- results_projid_6  all 206 site data
```{r GRP-temp-precip-all-point-plot-mean}
# results_projid_6  all 206 site data
GRP_all_site_temp_prec <- ggplot(results_projid_6, 
                      # text = paste("Site:", Site), 
                    aes(x = precip, y = temp, colour = Subregion)) +
                    geom_point(shape = 18, size = 2) +
    scale_x_continuous(limits = c(500, 2000)) +
    scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    scale_fill_brewer(palette = "Dark2") +
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="bottom") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "GRP Precipitation & Temperature",
                       subtitle = "By Subregion",
                      caption = "Chart by Wendy Anthony \n 2025-09-21",
                      x = "Precipitation", y = "Temperature")

## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), GRP_all_site_temp_prec) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "GRP Annual Precipitation & Temperature by Subregion<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Precipitation (mm)'),
        yaxis = list(title = 'Temperature (°C)'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-21",
             showarrow = F, xref='paper', yref='paper',
             xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             font = list(size=8, color="grey"))
        )
```

## violin plot - Temp 
```{r GRP-temp-sub-violin}

GRP_all_site_temp_violin <- ggplot(results_projid_6,
      aes(x = Subregion, y = temp)) +
      geom_violin(fill = "seagreen2") +
      geom_boxplot(width = 0.1, fill = "sandybrown") +
      theme_minimal() + #get rid of grey background and tick marks
      theme(legend.position="none") + #remove legend
      theme(axis.text.x = element_text(vjust = 1, size = 9)) +
      labs(title='Plotting Subregion by Temperature Range',
           subtitle='Data: GRP',
           caption = "Chart by Wendy Anthony \n 2025-09-22")
      
GRP_all_site_temp_violin

## ggplotly doesn't seem to work with violin plot??
    # ggplotly(tooltip = c("x", "y", "colour"), GRP_all_site_temp_violin) %>%
    #   config(displayModeBar = FALSE) %>% # This line disables the modebar
    #   layout(title = list(
    #     text = "GRP Annual Temperature by Subregion<br><sup>Data Source: GRP</sup>",
    #     y = 0.95), # Adjust vertical position if needed
    #     yaxis = list(title = 'Temperature (°C)'),
    #     annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-22",
    #          showarrow = F, xref='paper', yref='paper',
    #          xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
    #          font = list(size=8, color="grey"))
    #     )
```


## violin plot - Precip
```{r GRP-precip-subreg-violin}

GRP_all_site_prec_violin <- ggplot(results_projid_6,
      aes(x = Subregion, y = precip)) +
      geom_violin(fill = "seagreen2") +
      geom_boxplot(width = 0.1, fill = "sandybrown") +
      theme_minimal() + #get rid of grey background and tick marks
      theme(legend.position="none") + #remove legend
      theme(axis.text.x = element_text(vjust = 1, size = 9)) +
      labs(title='Plotting Subregion by Precipitation Range',
           subtitle='Data: GRP',
           caption = "Chart by Wendy Anthony \n 2025-09-22")
      
GRP_all_site_prec_violin
```


# Leaflet Maps

## Leaflet Site Map
```{r leaflet-grp-sites}
colnames(GRP_Site_data)

grp_site_data <- read.csv("data/GRP_all_site_data_sql.csv", header = TRUE, sep = ",")

# library(dplyr)
library(leaflet)

# Title control
# (Shackelford, et al., 2024). Global Restore Project. Restoration Futures Lab at the University of Victoria
 title <- '<p style="text-align: center; height: 18px; ">
<span style="font-size:9px;font-weight:bold; background-color: rgba(255, 255, 255, 0.9;");>Global Restore Project GOE Monitoring Sites</span><br>
<span style="font-size:6px;font-style:italic; background-color: rgba(255, 255, 255, 0.9;">(Data from Shackelford, et al., 2005-2022)</span></p>'

    # https://rstudio.github.io/leaflet/articles/markers.html
    # Create a palette that maps factor levels to colors
    #####
    # pal <- colorFactor(c("#d7191c", "#2c7bb6"), domain = c("Saanich Peninsula", "Gulf Islands"))
#     pal <- colorFactor(c("#7fc97f", "#beaed4", "#fdc086", "#ffff99", "#386cb0", "#f0027f", "#bf5b17"), domain = c("C-SP", "C-GI", "C-EVII", "C-EVI", "US-GI", "US-Wash", "US-Oreg"
# ))
 # #d95f02 orange; #7570b3 purple; #1b9e77 green
    pal <- colorFactor(c("#1b9e77", "#7570b3", "#7570b3", "#1b9e77", "#d95f02", "#d95f02", "#d95f02"), domain = c("C-SP", "C-GI", "C-EVII", "C-EVI", "US-GI", "US-Wash", "US-Oreg"))
GRP_site_map  <-  leaflet(grp_site_data) %>%
      addProviderTiles("Esri.WorldImagery") %>%
      addCircleMarkers(
        ~ longitude, ~ latitude,
        #
        color = ~pal(Subregion),
        # color = "yellow",
        weight = 1, # size of circle border
        stroke = TRUE, fillOpacity = 1, #fillOpacity = 0.5
        radius = 2,
        # OM_site_data$
        # "<b>Global Restore Project GOE Monitoring Sites</b>", "<br>", "<i>(Shackelford, et. al., 2005-2022)</i>", "<br><br>",
        popup = paste0(
          "<b>Site:</b> ", "<b>", grp_site_data$Site, "</b>",  "<br>",
          "<b>Subregion:</b> ", "<b>", grp_site_data$Subregion, "</b>",  "<br>",
          "<b>Project id:</b> ", "<b>", grp_site_data$projectid, "</b>",  "<br>",
          "<b>Landcover:</b> ", "<b>", grp_site_data$landcover, "</b>",  "<br>",
          "<b>Elevation:</b> ", "<b>", grp_site_data$elevation, " m</b>",  "<br>",
          "<b>Aspect:</b> ", "<b>", grp_site_data$aspect, " °</b>",  "<br>",
          "<b>Precipitation:</b> ", "<b>", grp_site_data$precip, " mm</b>",  "<br>",
          "<b>Temperature:</b> ", "<b>", grp_site_data$temp, " °C</b>"
          )) %>%

      setView(-123.44799, 48.64919, 10) %>%
      # add controls
      #  addMiniMap(width = 150, height = 150, zoomLevelOffset = -4) %>%
      addControl(title, position = "topright")

# Display map
GRP_site_map

# Save map
saveWidget(GRP_site_map, "output/GRP_site_map.html")
```






