---
title: "TEST Global-Restore-Project GOE DataViz"
author: "Wendy Anthony"
date: "2025/09/18"
output:
  html_document:
    highlight: zenburn
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries
```{r load_libraries}
library(shiny)
library(tidyverse)
library(plotly)
library(htmlwidgets)
```

# Load Data
- data from Nancy Shackelford at Global Restore Project https://globalrestoreproject.com/open_science/
```{r load-data}
OM_veg_1 <- read.csv("data/OM_vegResults1.csv", header = TRUE, sep = ",")
OM_veg_2 <- read.csv("data/OM_vegResults2.csv", header = TRUE, sep = ",")
OM_veg_3 <- read.csv("data/OM_vegResults3.csv", header = TRUE, sep = ",")
OM_veg_4 <- read.csv("data/OM_vegResults4.csv", header = TRUE, sep = ",")
OM_veg_5 <- read.csv("data/OM_vegResults5.csv", header = TRUE, sep = ",")
OM_site_data <- read.csv("data/OM_projectdata-site.csv", header = TRUE, sep = ",")
OM_treatments_data <- read.csv("data/OM_projectdata-treatments.csv", header = TRUE, sep = ",")
```

# Data Structure
## OM_veg_1
```{r structure-1}
str(OM_veg_1)
```

## OM_veg_2
```{r structure-2}
str(OM_veg_2)
```

## OM_veg_3
```{r structure-3}
str(OM_veg_3)
```

## OM_veg_4
```{r structure-4}
str(OM_veg_4)
```

## OM_veg_5
```{r structure-5}
str(OM_veg_5)
```
## OM_veg_1_5 cbind()
```{r cbind}
OM_veg_1_5 <- rbind(OM_veg_1, OM_veg_2, OM_veg_3, OM_veg_4, OM_veg_5)
```

## Filter OM_veg_1_5
```{r filter-OM_veg_1_5}
OM_veg_1_5_t_1_19 <- OM_veg_1_5[OM_veg_1_5$treatmentid == "1_19", ] #131 ibs
# OM_veg_1_5_t_1_19 <- OM_veg_1_5[OM_veg_1_5$treatmentid == "1_19" & OM_veg_1_5$tsr !=0, ] #131 ibs
```


### str OM_veg_1_5
<details><summary><b><i>Click open to see `structure of OM_veg_1_5`</i></b></summary><p>
```{r str-OM-veg-1-5}
str(OM_veg_1_5)
```
</p></details>


# sort unique species to df 
```{r OM_veg_1_5-unique-species}
library(dplyr)

colnames(OM_veg_1_5)
OM_veg_1_5_sp_or <- subset(OM_veg_1_5, select = c("speciesid", "speciesorigin"))
OM_veg_1_5_sp_or_sort <- OM_veg_1_5_sp_or %>%
                          arrange(speciesid)
# str(OM_veg_1_5_sp_or_sort)

# list of unique species values
OM_veg_1_5_sp_or_sort_unique <- unique(OM_veg_1_5_sp_or_sort$speciesid)
```

## OM_veg_1_5_sp_or_sort_unique
<details><summary><b><i>Click open to see results for structure of `OM_veg_1_5_sp_or_sort_unique`</i></b></summary><p>

```{r OM_veg_1_5-unique-species_str}
OM_veg_1_5_sp_or_sort_unique
```
</p></details>



## OM_treatments_data
```{r structure-OM-treatments-data}
str(OM_treatments_data)
colnames(OM_treatments_data)
```


## OM_site_data
```{r structure-OM_site_data}
str(OM_site_data)
```

## Round Temp values to 1 decimal OM_site_data
```{r round-temp}
OM_site_data$temp
OM_site_data$temp <- round(OM_site_data$temp, digits = 1)
OM_site_data$temp
```

### Sites: Choose Columns for Site `OM_site_data`
Note: `OM_projectdata-site.csv` was showing 1709 of empty rows at end of df that shows as NA    
- deleted empty rows from bottom of `OM_projectdata-2025-Ago-25.xlsx` before saving to csv file

```{r site-columns-table}
# OM_site_data <- read.csv("data/OM_projectdata-site.csv", header = TRUE, sep = ",")
str(OM_site_data)
colnames(OM_site_data)
# Removes to leave chosen ones to make new dataframe
# keep [2] "siteid" [3] "projectid" [4]"sitename" [5] "Subregion" [6] "latitude" [7] "longitude" [8] "refecosystem" [11] "landcover" [19] "elevation" [20] "aspect" [21]  "slope" [25] "tempc"  [36] "precip" 
GRP_Site_data <- OM_site_data[-c(1,9:11,13:19,23:25,27:36,38:49)]
```
### GRP_Site_data
<details><summary><b><i>Click open to see `GRP_Site_data`</i></b></summary><p>
```{r}
GRP_Site_data
```
</p></details>




### Sites: Save chosen columns as csv to input into Shiny drop down list of Sites
```{r save-sites-csv}
write.csv(GRP_Site_data, "data/GRP_Site_data.csv", row.names = FALSE)
```




# SQL Create Database
```{r libraries-created-db}
# https://www.youtube.com/watch?v=c3e8m3K0eC4
# 
# install.packages("RSQL")
# install.packages("RSQLite")
# install.packages("tidyverse") # v.1.3.2 to 2.0.0
library(RSQL)
library(RSQLite)
library(tidyverse)
# tidyverse_packages(include_self = TRUE)
```


```{r tidyverse-db-ex}
# ## Tidyverse db ex
# data("population")
# population
# 
# data("who")
# who
```


```{r create-database-connection-ex}
# ### create database and connection
# # builds empty db in R memory
# con <- dbConnect(drv = RSQLite::SQLite(),
#                  dbname = ":memory:")
# dbListTables(con)
```

```{r dbWriteTable-ex}
### Populate tables
# - dbWriteTable
# 
# # Load population table
# dbWriteTable(conn = con,
#              name = "population",
#              value = population)
# str(population)
# 
# # Load population table
# dbWriteTable(conn = con,
#              name = "who",
#              value = who)
# str(who)
# 
# #remove the local data from the environment
# rm(who, population)
# dbListTables(con)
```


```{r query-db-ex}
# ### Query for db
# res <- DBI::dbGetQuery(conn = con,
#          statement = "
#          SELECT who.country, who.year, who.new_sp_m014, population.population
#          FROM population, who
#          WHERE who.country = 'Afghanistan' AND 
#          who.year > 1995 AND
#          who.country = population.country AND
#          who.year = population.year
#          ")
# res

```

## GRP Database
### create GRP database and connection
```{r create-database-connection-grp}
# builds empty db in R memory
con <- dbConnect(drv = RSQLite::SQLite(),
                 dbname = ":memory:")
dbListTables(con)
```

### Populate GRP tables
- dbWriteTable
```{r dbWriteTable}
# Load OM_veg_1_5 table
dbWriteTable(conn = con,
             name = "OM_veg_1_5",
             value = OM_veg_1_5)
str(OM_veg_1_5)

# Load OM_site_data table
dbWriteTable(conn = con,
             name = "OM_site_data",
             value = OM_site_data)
str(OM_site_data)

# Load OM_treatments_data table
dbWriteTable(conn = con,
             name = "OM_treatments_data",
             value = OM_treatments_data)
str(OM_treatments_data)

dbListTables(con)
```

## Preview tables
```{r preview-tables}
tbl(src = con, "OM_site_data")
tbl(src = con, "OM_treatments_data")
tbl(src = con, "OM_veg_1_5")
```


## Query for db


### 1-Results query one tables
```{sql connection = con, output.var = "results_projid"}
SELECT *
FROM OM_site_data
WHERE projectid = 1
```

#### View query results

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid}
results_projid
```
</p></details>

### 2-Results query two tables
```{sql connection = con, output.var = "results_projid_2"}
SELECT s.siteid, t.treatmentid, s.Site
FROM OM_site_data AS s 
INNER JOIN OM_treatments_data AS t ON s.siteid = t.siteid
WHERE projectid = 1
```

#### View query results
<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_2}
results_projid_2
```
</p></details>

```{r}
# OM_veg_1_5: treatmentid  
# OM_site_data: siteid   sitename  
# OM_treatments_data: treatmentid   siteid  
# <!-- 
# # SELECT who.country, who.year, who.new_sp_m014, population.population -->
# <!-- #          FROM population, who -->
# <!-- #          WHERE who.country = 'Afghanistan' AND  -->
# <!-- #          who.year > 1995 AND -->
# <!-- #          who.country = population.country AND -->
# <!-- #          who.year = population.year -->
# 
# SELECT OM_site_data.siteid, OM_site_data.sitename, OM_treatments_data.treatmentid, OM_treatments_data.siteid, OM_veg_1_5.treatmentid, OM_treatments_data.restorationtype, OM_veg_1_5.year
# FROM OM_veg_1_5, OM_site_data, OM_treatments_data
# LEFT JOIN OM_site_data ON OM_site_data.siteid = OM_treatments_data.siteid AND OM_veg_1_5.treatmentid = OM_veg_1_5.treatmentid
# WHERE OM_treatments_data.restorationtype = 'seeding' AND OM_veg_1_5.year = 2005
```


### 3-Results query three tables - one species & 1 year
- SELECT DISTINCT for no duplicates
```{sql connection = con, output.var = "results_projid_3"}
SELECT DISTINCT s.siteid, t.treatmentid, s.Site, v.year, v.speciesid
FROM OM_site_data AS s 
INNER JOIN OM_treatments_data AS t ON s.siteid = t.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE speciesid = "Ant_odo" AND year = 2008
```

#### View query results
<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_3}
results_projid_3
```
</p></details>

### 3-Results query three tables - one Site
- SELECT DISTINCT for no duplicates
```{sql connection = con, output.var = "results_projid_4"}
SELECT DISTINCT s.siteid, s.Site, s.Subregion, t.treatmentid, v.year, v.speciesid, v.tsr, v.replicate, v.response
FROM OM_site_data AS s 
INNER JOIN OM_treatments_data AS t ON s.siteid = t.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE Site = "Christmas Hill"
```

#### View query results
<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_4}
results_projid_4
```
</p></details>

### 3-Results query three tables - all Sites
- SELECT DISTINCT for no duplicates
- should be 196 sites - not showing ones with same Sitename but different siteid
- include t.siteid to see that the correspondence works
- should be 196 sites (query results 149)
- WHERE s.DB = "OM"
```{sql connection = con, output.var = "results_projid_5"}
SELECT DISTINCT s.siteid, s.projectid, s.Site, s.Subregion, s.temp, s.precip, s.latitude, s.longitude, s.refecosystem, s.landcover, s.elevation, s.aspect
FROM OM_site_data AS s 
INNER JOIN OM_treatments_data AS t ON s.siteid = t.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE s.siteid BETWEEN 1 AND 196
```

#### View query results
- this doesn't have all sites, and some duplicate siteid  

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_5}
results_projid_5
```
</p></details>

### Save query as csv file with all Site inof
```{r create-csv-results_projid_5}
write.csv(results_projid_5, "data/GRP_all_site_data.csv", row.names = FALSE)
```


### ?? 3-Results query three tables - all Sites - still not showing all
```{sql connection = con, output.var = "results_projid_6"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, s.latitude, s.longitude, s.refecosystem, s.landcover, s.elevation, s.aspect, s.slope, s.temp, s.precip, s.disturbance
FROM OM_site_data AS s 
WHERE s.DB = "OM"
```

#### View query results
- this doesn't have all sites, and some duplicate siteid  

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_6}
results_projid_6
length(unique(results_projid_6$Site))
```
</p></details>

### Save query as csv file with all 206 Sites
```{r create-csv-results_projid_6}
write.csv(results_projid_6, "data/GRP_all_site_data_sql.csv", row.names = FALSE)
```



# Plots
## Temp & Precip
```{r GRP-temp-precip-all}

GRP_all_site_temp_prec <- ggplot(results_projid_6, 
                      # text = paste("Site:", Site), 
                    aes(x = precip, y = temp, colour = Subregion)) +
                    geom_point(shape = 18, size = 2) +
    scale_x_continuous(limits = c(500, 2000)) +
    scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    scale_fill_brewer(palette = "Dark2") +
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="bottom") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "GRP Precipitation & Temperature",
                       subtitle = "By Subregion",
                      caption = "Chart by Wendy Anthony \n 2025-09-21",
                      x = "Precipitation", y = "Temperature")

## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), GRP_all_site_temp_prec) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "GRP Annual Precipitation & Temperature by Subregion<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Precipitation (mm)'),
        yaxis = list(title = 'Temperature (°C)'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-21",
             showarrow = F, xref='paper', yref='paper',
             xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             font = list(size=8, color="grey"))
        )
```


# Leaflet Maps

## Leaflet Site Map
```{r leaflet-grp-sites}
colnames(GRP_Site_data)

GRP_Site_data <- read.csv("data/GRP_Site_data.csv", header = TRUE, sep = ",")

# library(dplyr)
library(leaflet)

# Title control
# (Shackelford, et al., 2024). Global Restore Project. Restoration Futures Lab at the University of Victoria
 title <- '<p style="text-align: center; height: 18px; ">
<span style="font-size:9px;font-weight:bold; background-color: rgba(255, 255, 255, 0.9;");>Global Restore Project GOE Monitoring Sites</span><br>
<span style="font-size:6px;font-style:italic; background-color: rgba(255, 255, 255, 0.9;">(Data from Shackelford, et al., 2005-2022)</span></p>'

    # https://rstudio.github.io/leaflet/articles/markers.html
    # Create a palette that maps factor levels to colors
    #####
    # pal <- colorFactor(c("#d7191c", "#2c7bb6"), domain = c("Saanich Peninsula", "Gulf Islands"))
#     pal <- colorFactor(c("#7fc97f", "#beaed4", "#fdc086", "#ffff99", "#386cb0", "#f0027f", "#bf5b17"), domain = c("C-SP", "C-GI", "C-EVII", "C-EVI", "US-GI", "US-Wash", "US-Oreg"
# ))
 # #d95f02 orange; #7570b3 purple; #1b9e77 green
    pal <- colorFactor(c("#1b9e77", "#7570b3", "#7570b3", "#1b9e77", "#d95f02", "#d95f02", "#d95f02"), domain = c("C-SP", "C-GI", "C-EVII", "C-EVI", "US-GI", "US-Wash", "US-Oreg"))
GRP_site_map  <-  leaflet(GRP_Site_data) %>%
      addProviderTiles("Esri.WorldImagery") %>%
      addCircleMarkers(
        ~ longitude, ~ latitude,
        #
        color = ~pal(Subregion),
        # color = "yellow",
        weight = 1, # size of circle border
        stroke = TRUE, fillOpacity = 1, #fillOpacity = 0.5
        radius = 2,
        # OM_site_data$
        # "<b>Global Restore Project GOE Monitoring Sites</b>", "<br>", "<i>(Shackelford, et. al., 2005-2022)</i>", "<br><br>",
        popup = paste0(
          "<b>Site:</b> ", "<b>", OM_site_data$Site, "</b>",  "<br>",
          "<b>Subregion:</b> ", "<b>", OM_site_data$Subregion, "</b>",  "<br>",
          "<b>Project id:</b> ", "<b>", OM_site_data$projectid, "</b>",  "<br>",
          "<b>Landcover:</b> ", "<b>", OM_site_data$landcover, "</b>",  "<br>",
          "<b>Elevation:</b> ", "<b>", OM_site_data$elevation, "</b>",  "<br>",
          "<b>Aspect:</b> ", "<b>", OM_site_data$aspect, "</b>")
      ) %>%

      setView(-123.44799, 48.64919, 10) %>%
      # add controls
      #  addMiniMap(width = 150, height = 150, zoomLevelOffset = -4) %>%
      addControl(title, position = "topright")

# Display map
GRP_site_map

# Save map
saveWidget(GRP_site_map, "output/_site_map.html")
```






