---
title: "TEST Global-Restore-Project GOE DataViz"
author: "Wendy Anthony"
date: "2025/09/18"
output:
  html_document:
    highlight: zenburn
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---
Modified 2025-09-28 09:39  
Modified 2025-10-01 07:52  
Modified 2025-10-02 17:00  
Working 2025-10-02 18:50   
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries
```{r load_libraries}
library(shiny)
library(tidyverse)
library(plotly)
library(htmlwidgets)
```

# Load Data
- data from Nancy Shackelford at Global Restore Project https://globalrestoreproject.com/open_science/
```{r load-data}
OM_veg_1 <- read.csv("data/OM_vegResults1.csv", header = TRUE, sep = ",")
OM_veg_2 <- read.csv("data/OM_vegResults2.csv", header = TRUE, sep = ",")
OM_veg_3 <- read.csv("data/OM_vegResults3.csv", header = TRUE, sep = ",")
OM_veg_4 <- read.csv("data/OM_vegResults4.csv", header = TRUE, sep = ",")
OM_veg_5 <- read.csv("data/OM_vegResults5.csv", header = TRUE, sep = ",")
OM_species_data <- read.csv("data/species_names-2024-Feb-28.csv", header = TRUE, sep = ",")
OM_site_data <- read.csv("data/OM_projectdata-site.csv", header = TRUE, sep = ",")
OM_treatments_data <- read.csv("data/OM_projectdata-treatments.csv", header = TRUE, sep = ",")
```

# Data Structure
## OM_veg_1
```{r structure-1}
str(OM_veg_1)
```

## OM_veg_2

### Structure OM_veg_2
```{r structure-2}
str(OM_veg_2)
```
### Unique species OM_veg_2
<details><summary><b><i>Click open to see `structure of OM_veg_2`</i></b></summary><p>
```{r unique-species-2}
str(OM_veg_2)
# unique(OM_veg_2$speciesid) # 315 unique species
```
</p></details>


## OM_veg_3
```{r structure-3}
str(OM_veg_3)
```

## OM_veg_4
```{r structure-4}
str(OM_veg_4)
```

## OM_veg_5
```{r structure-5}
str(OM_veg_5)
```
## OM_veg_1_5 cbind()
```{r cbind}
OM_veg_1_5 <- rbind(OM_veg_1, OM_veg_2, OM_veg_3, OM_veg_4, OM_veg_5)
```

### convert OM_veg_1_5$speciesorigin to lower case
```{r convert-lower-case-speciesorigin}
colnames(OM_veg_1_5)
unique(OM_veg_1_5$speciesorigin) #7 to #4
# [1] "exotic"  "native"  "unknown" NA        "Native"  "Exotic" [7] "Unknown"
OM_veg_1_5$speciesorigin <- tolower(OM_veg_1_5$speciesorigin)

unique(OM_veg_1_5$speciesorigin) #7 to #4
#  rename column
colnames(OM_veg_1_5)

str(OM_veg_1_5)
```



## Filter OM_veg_1_5
```{r filter-OM_veg_1_5}
OM_veg_1_5_t_1_19 <- OM_veg_1_5[OM_veg_1_5$treatmentid == "1_19", ] #131 ibs
# OM_veg_1_5_t_1_19 <- OM_veg_1_5[OM_veg_1_5$treatmentid == "1_19" & OM_veg_1_5$tsr !=0, ] #131 ibs
```


### str OM_veg_1_5
<details><summary><b><i>Click open to see `structure of OM_veg_1_5`</i></b></summary><p>
```{r str-OM-veg-1-5}
str(OM_veg_1_5)
```
</p></details>


# sort unique species to df 
```{r OM_veg_1_5-unique-species}

colnames(OM_veg_1_5)
OM_veg_1_5_sp_all_or <- subset(OM_veg_1_5, select = c("id", "DB", "treatmentid", "year", "tsr", "block", "replicate", "speciesid", "speciesorigin", "response", "responselevel", "responsemetric", "measurementscale", "measurementmetric"))
OM_veg_1_5_sp_all_or_sort <- OM_veg_1_5_sp_all_or %>%
                          arrange(speciesid)
OM_veg_1_5_sp_or <- subset(OM_veg_1_5, select = c("speciesid", "speciesorigin"))
OM_veg_1_5_sp_or_sort <- OM_veg_1_5_sp_or %>%
                          arrange(speciesid)
# str(OM_veg_1_5_sp_or_sort)

# list of unique species values
OM_veg_1_5_sp_or_sort_unique <- unique(OM_veg_1_5_sp_or_sort$speciesid)
```

## save csv
```{r save-OM_veg_1_5_sp_or_sort-csv}
# write.csv(OM_veg_1_5_sp_all_or_sort, "data/OM_veg_1_5_sp_all_or_sort.csv", row.names = FALSE)
```



## add speciesorigin to speciesid for OM_veg_2
<details><summary><b><i>Click open to see `list of species and origin OM_veg_2`</i></b></summary><p>
```{r}
OM_veg_2
```
</p></details>

### Order and sort OM_veg_2
<details><summary><b><i>Click open to see `sort order of OM_veg_2`</i></b></summary><p>
```{r order-sort-OM_veg_2}
OM_veg_1_5_sp_or_sort
```
</p></details>

### OM_veg_2 percent species changed from total species
```{r percent-OM_veg_2}
length(unique(OM_veg_2$speciesid)) # 316
length(OM_veg_2$speciesid) #15,230 records
length(OM_veg_1_5_sp_or_sort$speciesid) # 136364 records

length(OM_veg_2$speciesid) / length(OM_veg_1_5_sp_or_sort$speciesid) # 11.17%
```


## OM_veg_1_5_sp_or_sort_unique
<details><summary><b><i>Click open to see results for structure of `OM_veg_1_5_sp_or_sort_unique`</i></b></summary><p>
- use `, echo=T, results='hide'` to hide long results output
```{r OM_veg_1_5-unique-species_str}
# OM_veg_1_5_sp_or_sort_unique
```
</p></details>


## Add Species genus species & Form
- Add new columns to OM_veg_1_5 from OM_species_data
```{r species-genus-species-forms}
str(OM_veg_1_5)
str(OM_species_data)

OM_species_data_subset <- OM_species_data[, c("speciesid", "genus", "species", "lifeform")]
# OM_species_data_subset # 7310 observations

merged_OM_veg_1_5_species_form_df <- merge(OM_veg_1_5, OM_species_data_subset, by = "speciesid", all.x = TRUE)
#  merged_OM_veg_1_5_species_form_df
```



## OM_treatments_data
```{r structure-OM-treatments-data}
str(OM_treatments_data)
colnames(OM_treatments_data)
unique(OM_treatments_data$restorationtype)
unique(OM_treatments_data$treatment_type)
unique(OM_treatments_data$treatment_category)
```

### Treatments fire
<details><summary><b><i>Click open to see `fire treatment data`</i></b></summary><p>

```{r treatments-fire}

OM_treatments_data_fire <- OM_treatments_data[OM_treatments_data$treatment_category == "fire", ]
colnames(OM_treatments_data_fire)
OM_treatments_data_fire <- OM_treatments_data_fire[c(-1,-8, -10, -11, -12, -15, -16, -17)]

OM_treatments_data_fire
```
</p></details>

### Treatments mowing
<details><summary><b><i>Click open to see `mowing treatment data`</i></b></summary><p>
```{r treatments-mowing}
OM_treatments_data_mowing <- OM_treatments_data[OM_treatments_data$treatment_type == "mowing", ]
OM_treatments_data_mowing <- OM_treatments_data_mowing[c(-1,-8, -10, -11, -12, -15, -16, -17)]
OM_treatments_data_mowing
```
</p></details>

### Treatments herbicide
<details><summary><b><i>Click open to see `herbicide treatment data`</i></b></summary><p>
```{r treatments-herbicide}
OM_treatments_data_herbicide <- OM_treatments_data[OM_treatments_data$treatment_category == "herbicide", ]
OM_treatments_data_herbicide <- OM_treatments_data_herbicide[c(-1,-8, -10, -11, -12, -15, -16, -17)]

OM_treatments_data_herbicide
```
</p></details>
### Write Treatment csv
```{r write-treatment-csv}
# write.csv(OM_treatments_data_fire, "data/OM_treatments_data_fire.csv", row.names = FALSE)
# write.csv(OM_treatments_data_mowing, "data/OM_treatments_data_mowing.csv", row.names = FALSE)
# write.csv(OM_treatments_data_herbicide, "data/OM_treatments_data_herbicide.csv", row.names = FALSE)
```




## OM_site_data
```{r structure-OM_site_data}
str(OM_site_data)
```

# Temp
## Round Temp values to 1 decimal OM_site_data
```{r round-temp}
OM_site_data$temp
OM_site_data$temp <- round(OM_site_data$temp, digits = 1)
# OM_site_data$temp
```

## Mean Temp per site
```{r temp-mean-site}
site_temp_mean <- dplyr::distinct(OM_site_data) %>% group_by(Site) %>% summarise(temp = toString(temp))
site_temp_mean
```

## Mean Temp per Subregion
- this is listing all temp for each subregion
```{r temp-mean-Subregion}
subregion_temp <- dplyr::distinct(OM_site_data) %>% group_by(Subregion) %>% summarise(temp = toString(temp))
subregion_temp
str(subregion_temp) # temp is character thus can't take mean



# DOESN"T WORK
# mean(subregion_temp$temp)

# subregion_temp$temp <- as.numeric(as.character(subregion_temp$temp))
# subregion_temp$temp <- as.numeric(subregion_temp$temp)

# subregion_temp_mean <- aggregate(Subregion~temp, subregion_temp, FUN=mean)

# DOESN"T WORK
# subregion_temp_mean <- aggregate(subregion_temp, by=list(subregion_temp$temp), FUN=mean)
# subregion_temp_mean
```


# Precip
## Round Precip values to 1 decimal OM_site_data
```{r round-precip}
OM_site_data$precip
OM_site_data$precip <- round(OM_site_data$precip, digits = 1)
# OM_site_data$temp
```

## Mean precip per site
```{r precip-mean-site}
site_precip_mean <- dplyr::distinct(OM_site_data) %>% group_by(Site) %>% summarise(precip = toString(precip))
site_precip_mean
```

## Mean precip per Subregion
- this is listing all temp for each subregion
```{r precip-mean-Subregion}
subregion_precip_mean <- dplyr::distinct(OM_site_data) %>% group_by(Subregion) %>% summarise(precip = toString(precip))
subregion_precip_mean
```



# Aspect
## Round Aspect values to 2 decimal OM_site_data
```{r round-aspect}
OM_site_data$aspect
OM_site_data$aspect <- round(OM_site_data$aspect, digits = 2)
# OM_site_data$aspect
```

# Sites: 
## Choose Columns for Site `OM_site_data`
Note: `OM_projectdata-site.csv` was showing 1709 of empty rows at end of df that shows as NA    
- deleted empty rows from bottom of `OM_projectdata-2025-Ago-25.xlsx` before saving to csv file

```{r site-columns-table}
# OM_site_data <- read.csv("data/OM_projectdata-site.csv", header = TRUE, sep = ",")
str(OM_site_data)
colnames(OM_site_data)
# [1] "DB"                 "siteid"             "projectid"         
#  [4] "Site"               "Subregion"          "Subregion.an"      
#  [7] "latitude"           "longitude"          "refecosystem"      
# [10] "USDA.class"         "USDA.subclass"      "USDA.subsubclass"  
# [13] "landcover"          "growingseasonstart" "growingseasonend"  
# [16] "sand"               "silt"               "clay"              
# [19] "soildescription"    "soildepth"          "elevation"         
# [22] "aspect"             "slope"              "precipc"           
# [25] "tempc"              "aridity"            "temp"              
# [28] "drange"             "iso"                "tseason"           
# [31] "mtwarm"             "mtcold"             "trange"            
# [34] "twetq"              "tdryq"              "twarmq"            
# [37] "tcoldq"             "precip"             "pwet"              
# [40] "pdry"               "pseason"            "pwetq"             
# [43] "pdryq"              "pwarmq"             "pcoldq"            
# [46] "disturbance"        "minetype"           "ogtype"            
# [49] "invasivespe"        "invasiveform" 
# Removes to leave chosen ones to make new dataframe
# keep [2] "siteid" [3] "projectid" [4]"sitename" [5] "Subregion" [6] "latitude" [7] "longitude" [8] "refecosystem" [11] "landcover" [19] "elevation" [20] "aspect" [21]  "slope" [25] "tempc"  [36] "precip" 
GRP_Site_data <- OM_site_data[-c(6,12,14:21,24:26,28:37,39:50)]
```
## GRP_Site_data
<details><summary><b><i>Click open to see `GRP_Site_data`</i></b></summary><p>
```{r}
GRP_Site_data
```
</p></details>




### Sites: Save chosen columns as csv to input into Shiny drop down list of Sites
```{r save-sites-csv}
write.csv(GRP_Site_data, "data/GRP_Site_data.csv", row.names = FALSE)
```




# SQL Create Database
```{r libraries-created-db}
# https://www.youtube.com/watch?v=c3e8m3K0eC4
# 
# install.packages("RSQL")
# install.packages("RSQLite")
# install.packages("tidyverse") # v.1.3.2 to 2.0.0
library(RSQL)
library(RSQLite)
library(tidyverse)
# tidyverse_packages(include_self = TRUE)
```

# Sql GRP Database
## create GRP database and connection
- builds empty db in R memory  
- ## ## Warning message: call dbDisconnect() when finished working with a connection to avoid R memory overload and crash `dbDisconnect(con) `
```{r create-database-connection-grp}
# builds empty db in R memory
con <- dbConnect(drv = RSQLite::SQLite(),
                 dbname = ":memory:")
dbListTables(con)

## Warning message: call dbDisconnect() when finished working with a connection to avoid R memory overload and crashes e.g. after last query
# dbDisconnect(con) 
```

## Populate GRP tables
- dbWriteTable
```{r dbWriteTable}
# Load OM_veg_1_5 table
dbWriteTable(conn = con,
             name = "OM_veg_1_5",
             value = OM_veg_1_5)
str(OM_veg_1_5)

dbWriteTable(conn = con,
             name = "OM_species_data",
             value = OM_species_data)
str(OM_species_data)

# Load OM_site_data table
dbWriteTable(conn = con,
             name = "OM_site_data",
             value = OM_site_data)
str(OM_site_data)


# Load OM_treatments_data table
dbWriteTable(conn = con,
             name = "OM_treatments_data",
             value = OM_treatments_data)
str(OM_treatments_data)

dbListTables(con)
```

### Preview tables
```{r preview-tables}
tbl(src = con, "OM_site_data")
tbl(src = con, "OM_treatments_data")
tbl(src = con, "OM_species_data")
tbl(src = con, "OM_veg_1_5")
```


## SQL Query for db


### 1-Results query one tables
```{sql connection = con, output.var = "results_projid"}
SELECT *
FROM OM_site_data
WHERE projectid = 1
```

#### View query results

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid}
results_projid
```
</p></details>

### 2-Results query two tables
```{sql connection = con, output.var = "results_projid_2"}
SELECT s.siteid, t.treatmentid, s.Site
FROM OM_site_data AS s 
INNER JOIN OM_treatments_data AS t ON s.siteid = t.siteid
WHERE projectid = 1
```

#### View query results
<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_2}
results_projid_2
```
</p></details>

### 3-Results query three tables - one species & 1 year
- SELECT DISTINCT for no duplicates
```{sql connection = con, output.var = "results_projid_3"}
SELECT DISTINCT s.siteid, t.treatmentid, s.Site, v.year, v.speciesid
FROM OM_site_data AS s 
INNER JOIN OM_treatments_data AS t ON s.siteid = t.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE speciesid = "Ant_odo" AND year = 2008
```

#### View query results
<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_3}
results_projid_3
```
</p></details>

### 3-Results query three tables - one Site
- SELECT DISTINCT for no duplicates
```{sql connection = con, output.var = "results_projid_4"}
SELECT DISTINCT s.siteid, s.Site, s.Subregion, t.treatmentid, v.year, v.speciesid, v.tsr, v.replicate, v.response
FROM OM_site_data AS s 
INNER JOIN OM_treatments_data AS t ON s.siteid = t.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE Site = "Christmas Hill"
```

#### View query results
<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_4}
results_projid_4
```
</p></details>

### 3-Results query three tables - all Sites
- SELECT DISTINCT for no duplicates
- should be 196 sites - not showing ones with same Sitename but different siteid
- include t.siteid to see that the correspondence works
- should be 196 sites (query results 149)
- WHERE s.DB = "OM"
```{sql connection = con, output.var = "results_projid_5"}
SELECT DISTINCT s.siteid, s.projectid, s.Site, s.Subregion, s.temp, s.precip, s.latitude, s.longitude, s.refecosystem, s.landcover, s.elevation, s.aspect
FROM OM_site_data AS s 
INNER JOIN OM_treatments_data AS t ON s.siteid = t.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE s.siteid BETWEEN 1 AND 196
```

#### View query results
- this doesn't have all sites, and some duplicate siteid  

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_5}
results_projid_5
```
</p></details>

#### Save query as csv file with all Site inof
```{r create-csv-results_projid_5}
# write.csv(results_projid_5, "data/GRP_all_site_data.csv", row.names = FALSE)
```


### 3-Results query three tables - all Sites 
```{sql connection = con, output.var = "OM_all_sites"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, s.latitude, s.longitude, s.refecosystem, s.landcover, s.elevation, s.aspect, s.slope, s.temp, s.precip, s.disturbance
FROM OM_site_data AS s 
WHERE s.DB = "OM"
```

#### View query results

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_6}
OM_all_sites
length(unique(OM_all_sites$Site))
```
</p></details>

#### Save query as csv file with all 206 Sites
```{r create-csv-OM_all_sites}
# write.csv(OM_all_sites, "data/GRP_all_site_data_sql.csv", row.names = FALSE)
```



### 7 Query Results Treatment + Sites
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576  
```{sql connection = con, output.var = "results_projid_7"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, t.treatmentid, t.restorationtype, t.tsr_start_year, t.trt_year, t.treatmentmonth, t.treatment_category, t.treatment_type
FROM OM_treatments_data AS t 
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE s.DB = "OM"
```

#### View query results
- Query Results Treatment + Sites

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_7}
results_projid_7
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576   ... too large to view in Excel
```{r create-csv-results_projid_7}
# write.csv(results_projid_7, "data/GRP-Site-Treatment-SQL.csv", row.names = FALSE)
```

### 8 Query Results Treatment + Sites
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576 
- WHERE s.Subregion = "US-GI" >>> change Subregion and save csv for each  
- US-GI, US-Oreg, US-Wash, C-GI, C-EVI, C-EVII, C-SP
- add year from v.year as some treatments done include year start end  

```{sql connection = con, output.var = "results_projid_8"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, t.treatmentid, t.restorationtype, t.tsr_start_year, t.trt_year, t.treatmentmonth, t.treatment_category, t.treatment_type
FROM OM_treatments_data AS t 
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_veg_1_5 AS v ON t.treatmentid = v.treatmentid
WHERE s.Subregion = "C-SP"
```

#### View query results
- Query Results Treatment + Sites

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_8}
results_projid_8
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites
```{r create-csv-results_projid_8}
# write.csv(results_projid_8, "data/GRP-Site-Treatment-SQL-C-SP.csv", row.names = FALSE)
```


### 9 Query Results Species + Sites 
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576 
- WHERE s.Subregion = "US-GI" >>> change Subregion and save csv for each  
- US-GI, US-Oreg, US-Wash, C-GI, C-EVI, C-EVII, C-SP

```{sql connection = con, output.var = "results_projid_9"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, v.treatmentid, v.year AS "v_year_measure", v.replicate, v.speciesid, v.speciesorigin, v.response, v.responselevel, v.responsemetric, v.measurementscale, v.measurementmetric
FROM OM_veg_1_5 AS v
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_treatments_data AS t  ON v.treatmentid = t.treatmentid
WHERE s.Subregion = "US-Wash"
```

#### View query results
- Query Results Treatment + Sites

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-projid_9}
results_projid_9
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites
```{r create-csv-results_projid_9}
# write.csv(results_projid_9, "data/GRP-Site-SpeciesCover-SQL-US-Wash.csv", row.names = FALSE)
```

### 10 Query Results Species + Sites 
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576 
- WHERE s.Subregion = "US-GI" >>> change Subregion and save csv for each  
- US-GI, US-Oreg, US-Wash, C-GI, C-EVI, C-EVII, C-SP

```{sql connection = con, output.var = "site_species_C_GI"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, v.treatmentid, v.year AS "v_year_measure", v.replicate, v.speciesid, v.speciesorigin, v.response, v.responselevel, v.responsemetric, v.measurementscale, v.measurementmetric
FROM OM_veg_1_5 AS v
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_treatments_data AS t  ON v.treatmentid = t.treatmentid
WHERE s.Subregion = "C-GI"
```

#### View query results
- Query Results Treatment + Sites

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-site-species-C-GI}
site_species_C_GI
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites
```{r create-csv-results-site-species-C-GI}
# write.csv(site_species_C_GI, "data/GRP-Site-SpeciesCover-SQL-C-GI.csv", row.names = FALSE)
```


### 11 Query Results Species + Sites 
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576 
- WHERE s.Subregion = "US-GI" >>> change Subregion and save csv for each  
- US-GI, US-Oreg, US-Wash, C-GI, C-EVI, C-EVII, C-SP

```{sql connection = con, output.var = "site_species_C_SP"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, v.treatmentid, v.year AS "v_year_measure", v.replicate, v.speciesid, v.speciesorigin, v.response, v.responselevel, v.responsemetric, v.measurementscale, v.measurementmetric
FROM OM_veg_1_5 AS v
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_treatments_data AS t  ON v.treatmentid = t.treatmentid
WHERE s.Subregion = "C-SP"
```

#### View query results
- Query Results Treatment + Sites

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-site-species-C-SP}
site_species_C_SP
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites
```{r create-csv-results-site-species-C-SP}
# write.csv(site_species_C_SP, "data/GRP-Site-SpeciesCover-SQL-C-SP.csv", row.names = FALSE)
```



### 12a Query Add to `OM_veg_1_5` Results All Species + Sites  not all sites???
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576 
- WHERE s.Subregion = "US-GI" >>> change Subregion and save csv for each  
- US-GI, US-Oreg, US-Wash, C-GI, C-EVI, C-EVII, C-SP


```{sql connection = con, output.var = "site_species_all"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, v.treatmentid, t.tsr_start_year, t.trt_year, v.year AS "v_year_measure", v.replicate, v.speciesid, v.speciesorigin, v.response, v.responselevel, v.responsemetric, v.measurementscale, v.measurementmetric, f.genus, f.species, f.lifeform
FROM OM_veg_1_5 AS v
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_treatments_data AS t  ON v.treatmentid = t.treatmentid
INNER JOIN OM_species_data AS f ON v.speciesid = f.speciesid
WHERE v.DB = "OM"
```

#### View query results
- Query Results Treatment + Sites
- NOT ALL SITES !!
<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-site-species-all}
site_species_all
length(unique(site_species_all$siteid)) #139 s/b 206
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites >> 177MB filesize!!
- NOT ALL SITES !!
```{r create-csv-results-site-species-all}
# write.csv(site_species_all, "data/GRP-Site-SpeciesCover-SQL-all.csv", row.names = FALSE)
```



### 13 Query Results Species Specific Sites Mt. Tolmie
- dataframe *add Site & Subregion* to *treatment and species df* 
- `results_projid_7` R observation rows: 1,961,307  
- Excel maximum rows 1,048,576 
- WHERE s.Subregion = "US-GI" >>> change Subregion and save csv for each  
- US-GI, US-Oreg, US-Wash, C-GI, C-EVI, C-EVII, C-SP

```{sql connection = con, output.var = "species_site_62"}
SELECT s.siteid, s.projectid, s.Site, s.Subregion, v.treatmentid, v.year AS "v_year_measure", v.replicate, v.speciesid, v.speciesorigin, v.response, v.responselevel, v.responsemetric, v.measurementscale, v.measurementmetric
FROM OM_veg_1_5 AS v
INNER JOIN OM_site_data AS s ON t.siteid = s.siteid
INNER JOIN OM_treatments_data AS t  ON v.treatmentid = t.treatmentid
WHERE s.siteid = 62
```

#### View query results
- Query Results Treatment + Sites

<details><summary><b><i>Click open to see R code for `query results`</i></b></summary><p>
```{r preview-results-species-siteid-62}
species_site_62
```
</p></details>

#### Save query as csv file 
- Query Results Treatment + Sites
```{r create-csv-results-species-siteid-62}
# write.csv(species_site_62, "data/GRP-Site-SpeciesCover-SQL-species-site-62.csv", row.names = FALSE)
```


# dbDisconnect(con)
## Warning message:
```call dbDisconnect() when finished working with a connection``` 

```{r}
#con <- dbConnect(RSQLite::SQLite(), ":memory:")
dbDisconnect(con)
```




# Mt Tolmie 
#### Mt Tolmie Species Data
```{r mt-tolmie-species-data}
colnames(species_site_62)
str(species_site_62)
species_site_62_native <- species_site_62[species_site_62$speciesorigin == "native", ]
species_site_62_exotic <- species_site_62[species_site_62$speciesorigin == "exotic", ]

species_site_62_native_pres <- subset(species_site_62, (speciesorigin == 'native' & responsemetric == "presence"))

species_site_62_native_cov <- subset(species_site_62, (speciesorigin == 'native' & responsemetric == "cover"))

species_site_62_exotic_pres <- subset(species_site_62, (speciesorigin == 'exotic' & responsemetric == "presence"))

species_site_62_exotic_cov <- subset(species_site_62, (speciesorigin == 'exotic' & responsemetric == "cover"))

```

#### Mt Tolmie Cover Data
```{r mt-tolmie-cover-data}
species_site_62_by_replicate <- species_site_62 %>% group_by(replicate)
```

#### Mt Tolmie number Native & Exotic Species
```{r mt-tolmie-native}
nat_num_62 <- length(species_site_62_native_pres$speciesid) #45
nat_num_62
exo_num_62 <- length(species_site_62_exotic_pres$speciesid) #71
exo_num_62

tot_sp_62 <- nat_num_62 + exo_num_62
tot_sp_62

# percentage native species to total 
nat_num_62/tot_sp_62 * 100 # 38.7931

# percentage exotic species to total 
exo_num_62/tot_sp_62 * 100 # 61.2069

```
#### Mt Tolmie Exotic Species
```{r mt-tolmie-exotic}
species_site_62_by_replicate <- species_site_62 %>% group_by(replicate)
```

#### Save queried as csv file 
- Query Mt Tolmie species cover data
```{r create-csv-results-mt-tolmie-cover-data}
# write.csv(species_site_62_by_replicate, "data/GRP-Site-SpeciesCover-SQL-species-site-62-cover-data.csv", row.names = FALSE)

# write.csv(species_site_62_native_pres, "data/GRP-Site-SpeciesCover-SQL-species-site-62-cover-data-native-pres.csv", row.names = FALSE)
# write.csv(species_site_62_native_cov, "data/GRP-Site-SpeciesCover-SQL-species-site-62-cover-data-native-cov.csv", row.names = FALSE)
# write.csv(species_site_62_exotic_pres, "data/GRP-Site-SpeciesCover-SQL-species-site-62-cover-data-exotic-pres.csv", row.names = FALSE)
# write.csv(species_site_62_exotic_cov, "data/GRP-Site-SpeciesCover-SQL-species-site-62-cover-data-exotic-cov.csv", row.names = FALSE)
```


# Species Richness Calculation
- total number of species per Site

## Unique Sites
```{r unique}
length(unique(site_species_all$siteid))
```

## Calculate Species Richness 
```{r calculate-species-richness}
length(unique(site_species_all$speciesid)) # 626 unique species

site_species_all_native <- subset(site_species_all, speciesorigin == 'native')
site_species_all_native

site_species_all_exotic <- subset(site_species_all, speciesorigin == 'exotic')
site_species_all_exotic
 

# not enough sites
species_unique_site <- aggregate(speciesid ~ Site, data = site_species_all, FUN = function(x) length(unique(x)))
species_unique_site

# this would be better with Subregion added for better plotting
## Logic to add Subregion
# if site_species_all$Site  == species_unique_site$Site 
# then site_species_all$Subregion == species_unique_site$Subregion

# change column name to species richness
colnames(species_unique_site)
colnames(species_unique_site)[2] <- 'species_richness_all'

species_unique_site_exotic <- aggregate(speciesid ~ Site, data = site_species_all_exotic, FUN = function(x) length(unique(x)))
species_unique_site_exotic

colnames(species_unique_site_exotic)
colnames(species_unique_site_exotic)[2] <- 'species_richness_ex'

species_unique_site_native <- aggregate(speciesid ~ Site, data = site_species_all_native, FUN = function(x) length(unique(x)))
species_unique_site_native

colnames(species_unique_site_native)
colnames(species_unique_site_native)[2] <- 'species_richness_nat'


## should this be merge or cbind
# rounds2_final <- cbind(rounds2,rounds2_lower)
# rename columns
# Merge
species_unique_site_sp_rich <- merge(species_unique_site, species_unique_site_exotic, by = "Site")

species_unique_site_sp_rich <- merge(species_unique_site_sp_rich, species_unique_site_native, by = "Site")


# Merge with GRP_Site_data
# merged_df <- merge(df1, df2, by = "common_column_name") 
#GRP_Site_data_sp_rich <- merge(species_unique_site_sp_rich, GRP_Site_data, by = "Site")
GRP_Site_data <- merge(species_unique_site_sp_rich, GRP_Site_data, by = "Site")

```

# Functions
## Function Site All Species
`fun_site_species()`
```{r function-site-all-species}
fun_site_species <- function (fSite) {
subset(site_species_all, Site == fSite)  
}
length(unique(site_species_all$Site)) #148
```

## Function Subregion All Species
`fun_subreg_species()`
```{r function-site-all-species-subregion}
fun_subreg_species <- function (fSubregion) {
subset(site_species_all, Subregion == fSubregion)  
}
```

### Individual Site Species
#### Albert Head Individual Site Species
```{r one-site-species-fun}
species_AlbertHead <- fun_site_species("Albert Head")
```

##### Data explore
```{r species-AlbertHead-explore}
str(species_AlbertHead)
```

#### Plot Albert Head
```{r plot-albert-head}
str(species_AlbertHead)
colnames(species_AlbertHead)

length(unique(species_AlbertHead$speciesid)) # 74 unique species 

species_AlbertHead_native <- subset(species_AlbertHead, speciesorigin == 'native')
species_AlbertHead_native

species_AlbertHead_exotic <- subset(species_AlbertHead, speciesorigin == 'exotic')
species_AlbertHead_exotic
 

species_AlbertHead_plot <- ggplot(species_AlbertHead, 
                      # text = paste("Site:", Site), 
                    aes(x = v_year_measure, y = speciesid)) +
  # , color = speciesorigin
                    geom_point(shape = 20, size = 1) +
                    scale_x_continuous(breaks = c(2006, 2008)) +
                    geom_jitter() +
    #scale_x_continuous(limits = c(500, 2000)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_color_brewer(palette = "Accent") +
  
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="none") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Albert Head Species Data",
                         subtitle = "GRP",
                      caption = "Chart by Wendy Anthony \n 2025-09-28",
                      x = "Year", y = "Species")

 species_AlbertHead_plot
## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), species_AlbertHead_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Albert Head Species Data<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Year'),
        yaxis = list(title = 'Species'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-28"
              # no legend
             #               ,
             # showarrow = F, xref='paper', yref='paper',
             # xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             # font = list(size=8, color="grey")
             )
        )

```


### Individual Subregion Species
`fun_subreg_species()`
```{r one-subregion-species-fun}
species_C_GI <- fun_subreg_species("C-GI")
```





# Species Richness Plot 
- only 153 
```{r template-ggplot-geom_point-plotly-species-richness}
colnames(GRP_Site_data)
GRP_Site_data_plot <- ggplot(GRP_Site_data, 
                      # text = paste("Site:", Site), 
                    aes(x = factor(Site), y = species_richness_all, color = Subregion)) +
                    geom_point(shape = 20, size = 1) +
                    #scale_y_continuous(breaks = c(2006, 2008)) +
                    geom_jitter() +
    #scale_x_continuous(limits = c(500, 2000)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_color_brewer(palette = "Accent") +
  
                    theme_minimal() + #get rid of grey background and tick marks
#                    theme(legend.position="none") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Species Richness",
#                         subtitle = "(First Treatment 2005)",
                      caption = "Chart by Wendy Anthony \n 2025-10-02",
                      x = "Site", y = "Treatment Year")

# OM_treatments_data_fire_seed_plot
## ggplotly
    ggplotly(tooltip = c("x", "y", "color"), GRP_Site_data_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Species Richness by Site<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Site'),
        yaxis = list(title = 'Species Richness'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-10-02"
              # no legend
                           ,
             showarrow = F, xref='paper', yref='paper',
             xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             font = list(size=8, color="grey")
             )
        )

```




# For Loop
## for loop prep
- create a for loop for separate Site df  

<details><summary><b><i>Click open to see `data prep`</i></b></summary><p>
```{r for-loop-prep}
site_species_C_SP
length(site_species_C_SP$Site)
length(OM_all_sites$Site) # 206
length(unique(OM_all_sites$Subregion)) # 7

colnames(OM_all_sites)
sites_C_SP <- subset(OM_all_sites, Subregion == "C-SP")
# same as 
sites_C_SP <- OM_all_sites[OM_all_sites$Subregion == "C-SP", ]
sites_C_SP
sites_C_SP$Site

unique(sites_C_SP$Site) # 86 s/b 88

# sort alphabelticaly
sites_C_SP_sorted <- sites_C_SP[order(sites_C_SP$Site), ]
sites_C_SP_sorted

site_species_C_SP_sorted <- site_species_C_SP[order(site_species_C_SP$Site), ]
site_species_C_SP_sorted
```
</p></details>

##  Separate df for each Site of C_SP
<details><summary><b><i>Click open to see `for loop for separate sites`</i></b></summary><p>
```{r site-df-C_SP}

# sites_C_SP_sorted
site_species_C_SP_sorted

# create list of data frames for each C_SP Site
site_species_C_SP_df <- split( site_species_C_SP_sorted, site_species_C_SP_sorted$Site )
site_species_C_SP_MtTolmie <- site_species_C_SP_df$`Mt. Tolmie`

str(site_species_C_SP_df)

### WORKS !!!
# for (i in 1:length(sites_df)) {
#   assign(paste0("sites_df", i), as.data.frame(sites_df[[i]]))
# }
for (i in 1:length(site_species_C_SP_df)) {
  assign(paste0("site_species_C_SP_df", i, sep="_"), as.data.frame(site_species_C_SP_df[[i]]))
}

# sites_df[["Albert Head"]][["Site"]]

# EXTRACT LIST OF names of list dataframes
sites_df_namelist <- names(site_species_C_SP_df) # 76 
sites_df_namelist


# desired name derived from first 
desired_name_value <- site_species_C_SP_df$Site[1] # Taking the first value as an example
# this works to rename 1 df
site_species_C_SP_df_test <- site_species_C_SP_df
assign(paste("site_species_C_SP_df", desired_name_value, sep = "_"), site_species_C_SP_df_test)

```

</p></details>

### Rename dataframes
?? how can I do a for loop to rename all these  
??? Take care ... I crashed and corrupted this files predecessor after t
```{r rename-df}

# site_species_C_SP_df1_ <- assign(paste("site_species_C_SP_df1_", site_species_C_SP_df1_$Site, sep = ""), site_species_C_SP_df1_)
# 
# site_species_C_SP_df2_ <- assign(paste("site_species_C_SP_df2_", site_species_C_SP_df2_$Site, sep = ""), site_species_C_SP_df2_)
# 
# site_species_C_SP_df10_ <- assign(paste("site_species_C_SP_df10_", site_species_C_SP_df10_$Site, sep = ""), site_species_C_SP_df10_)
```



# Plots




## Treatments 
### Treatments Fire & Seeding

```{r treatments-fire-seed-plot}
colnames(OM_treatments_data_fire)
OM_treatments_data_fire_seed <- OM_treatments_data_fire[OM_treatments_data_fire$restorationtype == "seeding", ] 
unique(OM_treatments_data_fire_seed$siteid) #10 129:138

# display.brewer.all(n=10, exact.n=FALSE)

OM_treatments_data_fire_seed_plot <- ggplot(OM_treatments_data_fire_seed, 
                      # text = paste("Site:", Site), 
                    aes(x = factor(siteid), y = trt_year, color = treatmentid)) +
                    geom_point(shape = 20, size = 1) +
                    scale_y_continuous(breaks = c(2006, 2008)) +
                    geom_jitter() +
    #scale_x_continuous(limits = c(500, 2000)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_color_brewer(palette = "Accent") +
  
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="none") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Treatments with Fire and Seeding",
                         subtitle = "(First Treatment 2005)",
                      caption = "Chart by Wendy Anthony \n 2025-09-25",
                      x = "Site", y = "Treatment Year")

OM_treatments_data_fire_seed_plot
## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), OM_treatments_data_fire_seed_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Treatments with Fire and Seeding (First Treatment 2005)<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Site'),
        yaxis = list(title = 'Treatment Year'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-25"
              # no legend
             #               ,
             # showarrow = F, xref='paper', yref='paper',
             # xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             # font = list(size=8, color="grey")
             )
        )

```


### Treatments Mowing & Seeding

```{r treatments-mowing-seed-plot}
#colnames(OM_treatments_data_mowing)
#unique(OM_treatments_data_mowing$trt_year) 
# 2005 2006 2007 2008 2010 2011 2012 2013 2014 2015 2016 2009 2017 2018 2019 2020 2021 2022
#unique(OM_treatments_data_mowing$tsr_start_year) # 2005 2010 2009

OM_treatments_data_mowing_seed <- OM_treatments_data_mowing[OM_treatments_data_mowing$restorationtype == "seeding", ] 
unique(OM_treatments_data_mowing_seed$siteid) #10 129:138

# display.brewer.all(n=10, exact.n=FALSE)

OM_treatments_data_mowing_seed_plot <- ggplot(OM_treatments_data_mowing_seed, 
                      # text = paste("Site:", Site), 
                      
                    aes(x = factor(siteid), y = trt_year, color = treatmentid)) +
                    geom_point(shape = 20, size = 1) +
                    geom_jitter() +
#    scale_x_continuous(limits = c(129, 138)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_color_brewer(palette = "Accent") +
  
                    theme_minimal() + #get rid of grey background and tick marks
                    #theme(legend.position="bottom") +
                     theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Treatments with Mowing and Seeding",
                         subtitle = "(First Treatment 2005)",
                      caption = "Chart by Wendy Anthony \n 2025-09-25",
                      x = "Site", y = "Treatment Year")


## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), OM_treatments_data_mowing_seed_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Treatments with Mowing and Seeding (First Treatment 2005)<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Site'),
        yaxis = list(title = 'Treatment Year'),
        
        
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-25"              # no legend
             #               ,
             # showarrow = F, xref='paper', yref='paper',
             # xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             # font = list(size=8, color="grey")
             )
        ) 

```


### Treatments Herbicide

```{r treatments-herbicide-plot}
#colnames(OM_treatments_data_mowing)
#unique(OM_treatments_data_mowing$trt_year) 
# 2005 2006 2007 2008 2010 2011 2012 2013 2014 2015 2016 2009 2017 2018 2019 2020 2021 2022
#unique(OM_treatments_data_mowing$tsr_start_year) # 2005 2010 2009

OM_treatments_data_herbicide_seed <- OM_treatments_data_herbicide[OM_treatments_data_herbicide$restorationtype == "seeding", ] 
unique(OM_treatments_data_herbicide_seed$siteid) #10 129:138

OM_treatments_data_herbicide_seed_plot <- ggplot(OM_treatments_data_herbicide_seed, 
                      # text = paste("Site:", Site), 
                      
                    aes(x = factor(siteid), y = trt_year, color = treatmentid)) +
                    geom_point(shape = 20, size = 1) +
                    geom_jitter() +
#    scale_x_continuous(limits = c(129, 138)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_color_brewer(palette = "Accent") +
  
                    theme_minimal() + #get rid of grey background and tick marks
                    #theme(legend.position="bottom") +
                     theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Treatments with Herbicide and Seeding",
                         subtitle = "(First Treatment 2005)",
                      caption = "Chart by Wendy Anthony \n 2025-09-25",
                      x = "Site", y = "Treatment Year")


## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), OM_treatments_data_mowing_seed_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Treatments with Herbicide and Seeding (First Treatment 2005)<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Site'),
        yaxis = list(title = 'Treatment Year'),
        
        
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-25"              # no legend
             #               ,
             # showarrow = F, xref='paper', yref='paper',
             # xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             # font = list(size=8, color="grey")
             )
        ) 

```

## Template ggplot geom_point with plotly
```{r template-ggplot-geom_point-plotly}
OM_treatments_data_fire_seed_plot <- ggplot(OM_treatments_data_fire_seed, 
                      # text = paste("Site:", Site), 
                    aes(x = factor(siteid), y = trt_year, color = treatmentid)) +
                    geom_point(shape = 20, size = 1) +
                    scale_y_continuous(breaks = c(2006, 2008)) +
                    geom_jitter() +
    #scale_x_continuous(limits = c(500, 2000)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_color_brewer(palette = "Accent") +
  
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="none") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Treatments with Fire and Seeding",
                         subtitle = "(First Treatment 2005)",
                      caption = "Chart by Wendy Anthony \n 2025-09-25",
                      x = "Site", y = "Treatment Year")

# OM_treatments_data_fire_seed_plot
## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), OM_treatments_data_fire_seed_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Treatments with Fire and Seeding (First Treatment 2005)<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Site'),
        yaxis = list(title = 'Treatment Year'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-25"
              # no legend
             #               ,
             # showarrow = F, xref='paper', yref='paper',
             # xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             # font = list(size=8, color="grey")
             )
        )

```

# Sites
## Sites 129-138
- these sites had treatment done  

<details><summary><b><i>Click open to see `species data for sites 129-138`</i></b></summary><p>
```{r sites-species-129-138}
# [1] 128 129 134 135 136 137 138 130 131 132 133
# https://stackoverflow.com/questions/25647470/filter-multiple-values-on-a-string-column-in-dplyr
target_sites <- c("129", "130", "131", "132", "133", "134", "135", "136", "137", "138")
# filter
sites_species_129_138 <- dplyr::filter(results_projid_9, siteid %in% target_sites)
sites_species_129_138
```
</p></details>


## Filter Single Site Species Cover & Presence
- Mt Tolmie siteid = 62  
- can't use pivot longer
```{r plot-site-species-cover-presence-MtTolmie}
colnames(species_site_62)
# remove columns for Site 
ss_cover_62 <- species_site_62[, -c(2,3,4)]

# filter
ss_cover_62_presence <- ss_cover_62[ss_cover_62$responsemetric == "presence", ]
ss_cover_62_cover <- ss_cover_62[ss_cover_62$responsemetric == "cover", ]

ss_cover_62_cover_native <- ss_cover_62_cover[ss_cover_62_cover$speciesorigin == "native", ]

ss_cover_62_cover_exotic <- ss_cover_62_cover[ss_cover_62_cover$speciesorigin == "exotic", ]

ss_cover_62_cover_native_response_order <- ss_cover_62_cover_native[order(ss_cover_62_cover_native$response), ]

ss_cover_62_cover_exotic_response_order <- ss_cover_62_cover_exotic[order(ss_cover_62_cover_exotic$response), ]
```

### Write Mt. T0lmie cover csv
```{r write-cover-MtTolmie-csv}
# write.csv(ss_cover_62_cover_native_response_order, "data/62_cover_native.csv", row.names = FALSE)
# write.csv(ss_cover_62_cover_exotic_response_order, "data/62_cover_exotic.csv", row.names = FALSE)
```


## plot Mt. Tolmie native vs exotic species
- not showing all species  
- not showing all species - includes multiple replicates/species

### Mt Tolmie Native 
```{r ggplot-ss_cover_62_cover_native_response_order}
ss_cover_62_cover_native_response_order_plot <- ggplot(ss_cover_62_cover_native_response_order, 
                      # text = paste("Site:", Site), 
                    aes(y = speciesid, x = response, color = replicate)) +
                    geom_point(shape = 20, size = 1) +
                    geom_jitter() +
    #scale_x_continuous(limits = c(500, 2000)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_fill_brewer(palette = "Dark2") +
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="bottom") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Mt. Tolmie Native Species Cover %",
                      caption = "Chart by Wendy Anthony \n 2025-09-25",
                      y = "Species", x = "% Cover")

## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), ss_cover_62_cover_native_response_order_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Mt. Tolmie Native Plant % Cover<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Cover (%)'),
        yaxis = list(title = 'Species'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-25",
             showarrow = F, xref='paper', yref='paper',
             xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             font = list(size=8, color="grey"))
        )
```

### Mt Tolmie Exotic 
```{r ggplot-ss_cover_62_cover_exotic_response_order}
ss_cover_62_cover_exotic_response_order_plot <- ggplot(ss_cover_62_cover_exotic_response_order, 
                      # text = paste("Site:", Site), 
                    aes(y = speciesid, x = response, color = replicate)) +
                    geom_point(shape = 20, size = 1) +
                    geom_jitter() +
    #scale_x_continuous(limits = c(500, 2000)) +
    #scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    #scale_fill_brewer(palette = "Dark2") +
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="bottom") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "Mt. Tolmie Exotic Species Cover %",
                      caption = "Chart by Wendy Anthony \n 2025-09-25",
                      y = "Species", x = "% Cover")

## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), ss_cover_62_cover_exotic_response_order_plot) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "Mt. Tolmie Exotic Plant % Cover<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Cover (%)'),
        yaxis = list(title = 'Species'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-25",
             showarrow = F, xref='paper', yref='paper',
             xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             font = list(size=8, color="grey"))
        )
```



## 1-point plot - Temp & Precip
- I think this needs to be first use average of subregion temp and precip - showing too many here ???
```{r GRP-temp-precip-all-point}

GRP_all_site_temp_prec <- ggplot(OM_all_sites, 
                      # text = paste("Site:", Site), 
                    aes(x = precip, y = temp, colour = Subregion)) +
                    geom_point(shape = 18, size = 2) +
    scale_x_continuous(limits = c(500, 2000)) +
    scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    scale_fill_brewer(palette = "Dark2") +
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="bottom") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "GRP Precipitation & Temperature",
                       subtitle = "By Subregion",
                      caption = "Chart by Wendy Anthony \n 2025-09-21",
                      x = "Precipitation", y = "Temperature")

## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), GRP_all_site_temp_prec) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "GRP Annual Precipitation & Temperature by Subregion<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Precipitation (mm)'),
        yaxis = list(title = 'Temperature (°C)'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-21",
             showarrow = F, xref='paper', yref='paper',
             xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             font = list(size=8, color="grey"))
        )
```

## 2-point plot - Temp & Precip
- I think this needs to be first use mean of subregion temp and precip - showing too many here ???

### mean temp-subregion
```{r GRP-temp-subregion-mean}
OM_all_sites_test <- OM_all_sites
colnames(OM_all_sites_test)

subregion_mean_temp <- aggregate(OM_all_sites_test[, 12], list(OM_all_sites_test$Subregion), mean)
subregion_mean_temp # no results for C-GI, C-SP, US-GI

# doesn't work
#OM_all_sites %>% group_by(Subregion) %>% mutate(mean_temp=mean(temp))
```

### mean precip-subregion
- OM_all_sites  all 206 site data  

<details><summary><b><i>Click open to see `subregion precip data`</i></b></summary><p>

```{r GRP-precip-subregion-mean}
OM_all_sites
```
</p></details>

### 2-point plot - Temp & Precip
- OM_all_sites  all 206 site data
```{r GRP-temp-precip-all-point-plot-mean}
# OM_all_sites  all 206 site data
GRP_all_site_temp_prec <- ggplot(OM_all_sites, 
                      # text = paste("Site:", Site), 
                    aes(x = precip, y = temp, colour = Subregion)) +
                    geom_point(shape = 18, size = 2) +
    scale_x_continuous(limits = c(500, 2000)) +
    scale_y_continuous(limits = c(8, 12)) +
    # Replace default palette of pink & blue
    scale_fill_brewer(palette = "Dark2") +
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="bottom") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=1, size = 7)) +
                    labs(title = "GRP Precipitation & Temperature",
                       subtitle = "By Subregion",
                      caption = "Chart by Wendy Anthony \n 2025-09-21",
                      x = "Precipitation", y = "Temperature")

## ggplotly
    ggplotly(tooltip = c("x", "y", "colour"), GRP_all_site_temp_prec) %>%
      config(displayModeBar = FALSE) %>% # This line disables the modebar
      layout(title = list(
        text = "GRP Annual Precipitation & Temperature by Subregion<br><sup>Data Source: GRP</sup>",
        y = 0.95), # Adjust vertical position if needed
        xaxis = list(title = 'Precipitation (mm)'),
        yaxis = list(title = 'Temperature (°C)'),
        annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-21",
             showarrow = F, xref='paper', yref='paper',
             xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
             font = list(size=8, color="grey"))
        )
```

## violin plot - Temp 
- Warning: Removed 67 rows containing non-finite values (`stat_ydensity()`).  
- Warning: Removed 67 rows containing non-finite values (`stat_boxplot()`). 
```{r GRP-temp-sub-violin}

GRP_all_site_temp_violin <- ggplot(OM_all_sites,
      aes(x = Subregion, y = temp)) +
      geom_violin(fill = "seagreen2") +
      geom_boxplot(width = 0.1, fill = "sandybrown") +
      theme_minimal() + #get rid of grey background and tick marks
      theme(legend.position="none") + #remove legend
      theme(axis.text.x = element_text(vjust = 1, size = 9)) +
      labs(title='Plotting Subregion by Temperature Range',
           subtitle='Data: GRP',
           caption = "Chart by Wendy Anthony \n 2025-09-22")
      
GRP_all_site_temp_violin

## ggplotly doesn't seem to work with violin plot??
    # ggplotly(tooltip = c("x", "y", "colour"), GRP_all_site_temp_violin) %>%
    #   config(displayModeBar = FALSE) %>% # This line disables the modebar
    #   layout(title = list(
    #     text = "GRP Annual Temperature by Subregion<br><sup>Data Source: GRP</sup>",
    #     y = 0.95), # Adjust vertical position if needed
    #     yaxis = list(title = 'Temperature (°C)'),
    #     annotations = list(x = 1, y = -0.1, text = "Chart by Wendy Anthony\n2025-09-22",
    #          showarrow = F, xref='paper', yref='paper',
    #          xanchor='right', yanchor= 'auto', xshift=0, yshift=0,
    #          font = list(size=8, color="grey"))
    #     )
```


## violin plot - Precip
- Warning: Removed 67 rows containing non-finite values (`stat_ydensity()`).  
- Warning: Removed 67 rows containing non-finite values (`stat_boxplot()`). 
```{r GRP-precip-subreg-violin}

GRP_all_site_prec_violin <- ggplot(OM_all_sites,
      aes(x = Subregion, y = precip)) +
      geom_violin(fill = "seagreen2") +
      geom_boxplot(width = 0.1, fill = "sandybrown") +
      theme_minimal() + #get rid of grey background and tick marks
      theme(legend.position="none") + #remove legend
      theme(axis.text.x = element_text(vjust = 1, size = 9)) +
      labs(title='Plotting Subregion by Precipitation Range',
           subtitle='Data: GRP',
           caption = "Chart by Wendy Anthony \n 2025-09-22")
      
GRP_all_site_prec_violin
```


# Leaflet Maps

## Leaflet Site Map
```{r leaflet-grp-sites}

# library(dplyr)
library(leaflet)

# Title control
# (Shackelford, et al., 2024). Global Restore Project. Restoration Futures Lab at the University of Victoria
 title <- '<p style="text-align: center; height: 18px; ">
<span style="font-size:9px;font-weight:bold; background-color: rgba(255, 255, 255, 0.9;");>Global Restore Project GOE Monitoring Sites</span><br>
<span style="font-size:6px;font-style:italic; background-color: rgba(255, 255, 255, 0.9;">(Data from Shackelford, et al., 2005-2022)</span></p>'

    # https://rstudio.github.io/leaflet/articles/markers.html
    # Create a palette that maps factor levels to colors
    #####
    # pal <- colorFactor(c("#d7191c", "#2c7bb6"), domain = c("Saanich Peninsula", "Gulf Islands"))
#     pal <- colorFactor(c("#7fc97f", "#beaed4", "#fdc086", "#ffff99", "#386cb0", "#f0027f", "#bf5b17"), domain = c("C-SP", "C-GI", "C-EVII", "C-EVI", "US-GI", "US-Wash", "US-Oreg"
# ))
 # #d95f02 orange; #7570b3 purple; #1b9e77 green
    pal <- colorFactor(c("#1b9e77", "#7570b3", "#7570b3", "#1b9e77", "#d95f02", "#d95f02", "#d95f02"), domain = c("C-SP", "C-GI", "C-EVII", "C-EVI", "US-GI", "US-Wash", "US-Oreg"))
GRP_site_map  <-  leaflet(GRP_Site_data) %>%
      addProviderTiles("Esri.WorldImagery") %>%
      addCircleMarkers(
        ~ longitude, ~ latitude,
        #
        color = ~pal(Subregion),
        # color = "yellow",
        weight = 1, # size of circle border
        stroke = TRUE, fillOpacity = 1, #fillOpacity = 0.5
        radius = 2,
        # OM_site_data$
        # "<b>Global Restore Project GOE Monitoring Sites</b>", "<br>", "<i>(Shackelford, et. al., 2005-2022)</i>", "<br><br>",
        popup = paste0(
          "<b>Site:</b> ", "<b>", GRP_Site_data$Site, "</b>",  "<br>",
          "<b>Subregion:</b> ", "<b>", GRP_Site_data$Subregion, "</b>",  "<br>",
          "<b>Project id:</b> ", "<b>", GRP_Site_data$projectid, "</b>",  "<br>",
          "<b>Landcover:</b> ", "<b>", GRP_Site_data$landcover, "</b>",  "<br>",
          "<b>Elevation:</b> ", "<b>", GRP_Site_data$elevation, " m</b>",  "<br>",
          "<b>Aspect:</b> ", "<b>", GRP_Site_data$aspect, " °</b>",  "<br>",
          "<b>Precipitation:</b> ", "<b>", GRP_Site_data$precip, " mm</b>",  "<br>",
          "<b>Temperature:</b> ", "<b>", GRP_Site_data$temp, " °C</b>",  "<br>",
          "<b>Species Richness (All):</b> ", "<b>", GRP_Site_data$species_richness_all, "</b>",  "<br>",
          "<b>Species Richness (Native):</b> ", "<b>", GRP_Site_data$species_richness_nat, "</b>",  "<br>",
          "<b>Species Richness (Exotic):</b> ", "<b>", GRP_Site_data$species_richness_ex, "</b>"
          
          ))  %>%
      addLegend("bottomright", pal = pal, values = GRP_Site_data$Subregion, title = "Subregion") %>%

      setView(-123.44799, 48.64919, 8) %>%
      # add controls
      #  addMiniMap(width = 150, height = 150, zoomLevelOffset = -4) %>%
      addControl(title, position = "topright")

# Display map
GRP_site_map

# Save map
saveWidget(GRP_site_map, "output/GRP_site_map.html")
```






