---
title: "Project Proposal for Vizualizing HAT's GOE Monitoring Data"
author: "Wendy Anthony"
date: "2025/08/30"
output:
  html_document:
    df_print: paged
    toc: yes 
    toc_depth: 3
    toc_float: yes
    number_sections: yes
  pdf_document: 
    # default
    latex_engine: xelatex
    toc: true
mainfont: Monaco # change from default TimesRoman - same font used by RStudio
fontsize: 11pt
---

```{r, output-option-pdf, , include=FALSE}
# output: pdf_document
```

```{r, output-option-html, , include=FALSE}
# output:
#   html_document:
#     toc: yes 
#     toc_depth: 3
#     toc_float: yes
#     number_sections: yes
```

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Idea for Draft Project Proposal
- I am a student in UVic's Restoration of Natural Systems program interested in making a significant contribution to the local restoration community, and I'm looking to assist an interested community partner through the creation of a Final Restoration Project 
- Nancy Shackelford (RNS Director) suggested that I might consider helping HAT GOE Monitoring Project by creating a tool for vizualizing GOE Monitoring Data. 
- I propose to create a customized-for-HAT automatic coding tool that can be used by non-coders. 
- I have already spent almost 30 hours (during August 2025) reviewing and learning the new-to-me R code I would need to complete this project, and feel confident I will be able to learn any other R code as will be deemed needed after consultation with HAT, and receiving HAT DATA to analyze its structure and formats 

# What I need
## From HAT
- access to HAT GOE Monitoring Data 
- support, input, and feedback in order 
- I met with Vanessa Brownlee 2025-08-27, who asked that I send her further information about my proposed project, and though still new to her position, she seemed to be willing for me to start this project in Fall 2025 semester 

## From RNS
- support and supervision from RNS Nancy Shackelford, particularly before she goes on sabatical July 2026 

# What I can offer
- A keen interest in restoring Garry Oak Ecosystems 
- Extensive knowledge and experience with coding using `R` 
- I am a self-motivated, low maintenance student, very interested in the natural environment, and who loves to continue learning in meaningful ways 
- I can also assist HAT with the monitoring raw data clean-up as needed 


# Proposed Deliverables
## Written report
- The report will provide a record of the work I did to complete this project 
- Report will follow the format, length, and expectations of RNS program 

## Tutorials for using Data Vizualization Tools
- easy to understand 
- customized 

## RMarkdown Data Processing Document 
- for data downloading, cleaning, and formatting data visualizations 
- explain process for preparing data for use 
- can be used to create HTML or pdf reports 
- HTML Document includes a Table of Contents for easy navigation between parts 

## Shiny interactive tool 
- https://goe-interact.shinyapps.io/Test-Shiny-GOE-Monitor/ 
- using data processed with RMarkdown document 
- for choosing data variables to visualize in tables, plot charts, and maps 

## Vizualizations
- data tables, chart plots, maps 

# Data
- Data used for this proposal is from the report that HAT used as a model for their GOE Monitoring Project 
- Malloff, J., & Shackelford, N. (2024). <i>Feeling the Pulse: Monitoring methods and initial outcomes in oak meadow ecosystems.</i> Restoration Futures Lab at the University of Victoria. 


## Data download and formatting
- Data (in excel or csv format) is uploaded 
- Data is formatted, saved, and managed to be useful for creating visualizations 
- The code has been hidden for this report, highlighting only the resulting plots 

```{r, load-libraries, echo=FALSE, results=FALSE, results='hide', message=FALSE}
library(shiny)
library(tidyverse)
library(plotly)
```

```{r, load-data, echo=FALSE, results=FALSE, results='hide', message=FALSE}
# Site Data
geo_eccc <- read.csv("SiteData-TEST-GOE-Monitor.csv", header = TRUE, sep = ",")
# Site Details
Sites_eccc <- read.csv("SiteDetails-TEST-GOE-Monitor.csv", header = TRUE, sep = ",")
```

```{r, echo=FALSE, results=FALSE, results='hide', message=FALSE}
Percentage_ns <- geo_eccc$Proportion_of_native_species * 100
```

```{r, add-column, echo=FALSE, results=FALSE, results='hide', message=FALSE}
geo_eccc <- cbind(geo_eccc, Percentage_ns)
```

```{r, arange-order-df, echo=FALSE, results=FALSE, results='hide', message=FALSE}
Sites_eccc_1 <- Sites_eccc %>% arrange(Sites_eccc$Site)
geo_eccc_1 <- geo_eccc %>% arrange(geo_eccc$Site)
```

```{r, cbind-arranged-df, echo=FALSE, results=FALSE, results='hide', message=FALSE}
geo_eccc_sites <- cbind(Sites_eccc_1, geo_eccc_1)
```

```{r, remove-duplicate-columns, echo=FALSE, results=FALSE, results='hide', message=FALSE}
geo_eccc_sites_1 <- geo_eccc_sites[c(-9,-10)]
```

```{r, rename-columns-baser, echo=FALSE, results=FALSE, results='hide', message=FALSE}
names(geo_eccc_sites_1)[names(geo_eccc_sites_1) == "Land.Manager"] <- "Land_Manager"
names(geo_eccc_sites_1)[names(geo_eccc_sites_1) == "Main.Restoration.Type"] <- "Main_Restoration_Type"
names(geo_eccc_sites_1)[names(geo_eccc_sites_1) == "Restoration.Intensity"] <- "Restoration_Intensity"
names(geo_eccc_sites_1)[names(geo_eccc_sites_1) == "Area.ha"] <- "Area_ha"
```

```{r, create-csv-combined-datasets, echo=FALSE, results=FALSE, results='hide', message=FALSE}
write.csv(geo_eccc_sites_1, "geo_eccc_all_site_data.csv", row.names = FALSE)
```

\newpage

# Data Visualization Examples

## Data Table (Malloff & Shackelford, 2024)
\renewcommand{\arraystretch}{.25}
### Full Data Table

```{r, data table_html, echo=FALSE}
# for html
DT::datatable(geo_eccc_sites_1)
```

### Table Head of Site Information
```{r, data-table-pdf-sites, echo=FALSE, warning=FALSE}

# pdf
kableExtra::kable(
  head(Sites_eccc_1), booktabs = TRUE, linesep = ""
) %>%
  kableExtra::kable_styling(latex_options = "hold_position", font_size = 6)
```

### Head of Site Data
```{r, data-table-pdf-data, echo=FALSE, warning=FALSE}

# pdf
kableExtra::kable(
  head(geo_eccc_1), booktabs = TRUE, linesep = ""
) %>%
  kableExtra::kable_styling(latex_options = "hold_position", font_size = 6)
```

\newpage
## Map of Site Locations

### Leaflet Map with Circle Markers 
- Data from previous data table is summarized in the map marker popups 
- click markers to open site data summary window 

```{r, leaflet_map_circle_markers_sized_radius, echo=FALSE, fig.align = 'center'}

geo_eccc_sites_1 <- read.csv("geo_eccc_all_site_data.csv")

library(dplyr)
library(leaflet)
library(htmlwidgets) # for responsive web map

title <- '<p style="text-align: center; height: 30px; ">
<span style="font-size:10px;font-weight:bold; background-color: rgba(255, 255, 255, 0.9;");>ECCC GOE 2023 Monitoring Sites</span><br>
<span style="font-size:8px;font-style:italic; background-color: rgba(255, 255, 255, 0.9;">(Malloff & Shackelford, 2024)</span></p>'
 
pal <- colorFactor(c("#d7191c", "#2c7bb6"), domain = c("Saanich Peninsula", "Gulf Islands"))

ECCC_map_2 <- leaflet(geo_eccc_sites_1) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addCircleMarkers(
             ~ Lng, ~ Lat, 
         
             radius = ~ifelse(Subregion == "Saanich Peninsula", 8, 8),
    fillColor = ~pal(Subregion),
    color = "black",
    weight = 3, # size of circle border
    stroke = TRUE, fillOpacity = 0.5,
             popup = paste0("<b>ECCC GOE 2023 Monitoring Data</b>", "<br>", "<i>(Malloff & Shackelford, 2024)</i>", "<br><br>",
               "<b>Site:</b> ", "<b>", geo_eccc_sites_1$Site, "</b>", "</b>", "<br><br>", "<b>Subregion:</b> ", geo_eccc_sites_1$Subregion, "<br>",
                        "<b>Latitude:</b> ", geo_eccc_sites_1$Lat, "<br>", "<b>Longitude:</b> ", geo_eccc_sites_1$Lng,  "<br><br>",
                        "<b>Land Manager:</b> ", geo_eccc_sites_1$Land_Manager, "<br>",
                        "<b>Main Restoration:</b> ", geo_eccc_sites_1$Main_Restoration_Type, "<br>",
                        "<b>Restoration Intensity:</b> ", geo_eccc_sites_1$Restoration_Intensity, "<br>",
                        "<b> Area:</b> ", geo_eccc_sites_1$Area_ha, " ha",
                        "<br><br>",
                        "<b>Proportion of native species:</b> ", geo_eccc_sites_1$Proportion_of_native_species, "<br>",
                        "<b>Cultural species richness:</b> ", geo_eccc_sites_1$Cultural_species_richness, "<br>",
                        "<b>Exotic species:</b> ", geo_eccc_sites_1$Exotic_species, "<br>",
                        "<b>Trampling:</b> ", geo_eccc_sites_1$Trampling, "<br>",
                        "<b>Herbivory:</b> ", geo_eccc_sites_1$Herbivory, "<br>",
                        "<b>Composite Index:</b> ", geo_eccc_sites_1$Composite_Index, "<br>")) %>%
  setView(-123.44799, 48.64919, 10) %>%
  # add controls
  addMiniMap(width = 100, height = 100, zoomLevelOffset = -4) %>%
  addControl(title, position = "topright")

# Display map
ECCC_map_2
```

\newpage

## Subregion Plots

```{r, plots_Exotic_species, echo=FALSE, results=FALSE, results='hide', include=FALSE, message=FALSE}
subregion_exsp <- ggplot(geo_eccc, 
                    aes(x = Subregion, y = Exotic_species)) +
                    geom_violin(fill = "seagreen2") +
                    geom_boxplot(width = 0.1, fill = "sandybrown") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(vjust = 1, size = 9)) +
                    labs(title='Plotting Subregion by Exotic Species',
                      subtitle='ECCC',
                      x = "Subregion", y = "Exotic Species")
subregion_exsp
# plotly won't work here
```


```{r, plots_Percentage_ns, echo=FALSE, results=FALSE, results='hide', include=FALSE, message=FALSE}

subregion_nsp <- ggplot(geo_eccc, 
                    aes(x = Subregion, y = Percentage_ns)) +
                    geom_violin(fill = "seagreen2") +
                    geom_boxplot(width = 0.1, fill = "sandybrown") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(vjust = 1, size = 9)) +
                    labs(title='Plotting Subregion by Percentage of Native Species',
                      subtitle='ECCC',
                       x = "Subregion", y = "Percentage Native Species")
subregion_nsp
# plotly won't work here
```

### Violin/Bar Plot: Compare Subregions Exotic and Native Species
```{r, cowplot_violin_subr, echo=FALSE}
library(cowplot)

subregion_exsp_1 <- subregion_exsp + labs(title='Exotic Species',
                      subtitle='ECCC (Malloff & Shackelford, 2024)'
                      )

subregion_nsp_1 <- subregion_nsp + labs(title='Native Species',
                      subtitle='ECCC (Malloff & Shackelford, 2024)',
                      caption = "Chart by Wendy Anthony \n 2025-08-27")

plot_grid(subregion_exsp_1, subregion_nsp_1, labels = "AUTO")
```

\newpage

## Site Plots

### Point Plot: Compare Sites by Restoration Intensity
- click data points for tool-tip of data results 

```{r, plots-site-Restoration-Intensity-order-point, echo=FALSE}

# Step 1
desired_levels <- c("high", "moderate", "low", "minimal")
# create new df to create desired levels
geo_eccc_sites_2 <- geo_eccc_sites_1
# this changes data type from character to factor
geo_eccc_sites_2$Restoration_Intensity <- factor(geo_eccc_sites_1$Restoration_Intensity, levels = desired_levels)

# step 2
geo_eccc_sites_2 <- geo_eccc_sites_2[order(geo_eccc_sites_2$Restoration_Intensity),]

# values in order now -- ?? how to change to reverse order 
site_restInt <- ggplot(geo_eccc_sites_2,
       aes(x = Site, y = reorder(Restoration_Intensity, desc(Restoration_Intensity)), color = Subregion)) +
  # reorder y variable axis
  # https://stackoverflow.com/questions/28391850/reverse-order-of-discrete-y-axis-in-ggplot2s
# y = Restoration_Intensity
                    geom_point(shape = 18, size = 2) +
  #geom_bar(stat="identity") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="bottom") + #remove legend 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Plotting Sites by Restoration Intensity',
                      subtitle='ECCC (Malloff & Shackelford, 2024)',
                      caption = "Chart by Wendy Anthony \n 2025-08-22", x = "Site", y = "Restoration Intensity")

ggplotly(site_restInt)
```


\newpage

### Point Plot: Sites Ordered by Restoration Intensity
- this uses the same plot code as the previous example, changing only the order of the restoration intensity from minimal to high 
- click data points for tool-tip of data results 

```{r, plots_site_Restoration_Intensity_order_point_1, echo=FALSE}

# Step 1
desired_levels <- c("high", "moderate", "low", "minimal")
# create new df to create desired levels
geo_eccc_sites_2 <- geo_eccc_sites_1
# this changes data type from character to factor
geo_eccc_sites_2$Restoration_Intensity <- factor(geo_eccc_sites_1$Restoration_Intensity, levels = desired_levels)

# step 2
geo_eccc_sites_2 <- geo_eccc_sites_2[order(geo_eccc_sites_2$Restoration_Intensity),]

# values in order now -- ?? how to change to reverse order 
site_restInt_x <- ggplot(geo_eccc_sites_2,
       aes(x = reorder(Site, 
                       desc(Restoration_Intensity)), 
           y = reorder(Restoration_Intensity, desc(Restoration_Intensity)), color = Subregion)) +
  # reorder y variable axis
  # https://stackoverflow.com/questions/28391850/reverse-order-of-discrete-y-axis-in-ggplot2s
# y = Restoration_Intensity
                    geom_point(shape = 18, size = 2) +
  #geom_bar(stat="identity") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="bottom") + 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Plot Site Restoration Intensity',
                      subtitle='ECCC (Malloff & Shackelford, 2024)',
                      caption = "Chart by Wendy Anthony \n 2025-08-26",
                      x = "Site", y = "Restoration Intensity")

# ggplotly(site_restInt_x)
leg <- list(
  font = list(
    family = "sans-serif",
    size = 8,
    color = "#000"))
# xanchor = "center", too far left
plot <- ggplotly(site_restInt_x)
plot <- plot %>% layout(legend = list(x = 0.15, y = -0.5, orientation = 'h', size = 0.025))  
plot <- plot %>% layout(legend = leg)
#plot
ggplotly(plot)
```


\newpage

### Point Plot: Individual Site - Uplands Variables of Degradation

```{r, plots-Uplands-Degrade, echo=FALSE}
# Uplands
Uplands <- geo_eccc_1[24, ]
Uplands <- Uplands[, -c(1,2,3,4,8,9,10)]
# names(Uplands)
Uplands_long <- Uplands %>% pivot_longer(everything(), names_to = "variable", values_to = "value")

Uplands_long_plot <- ggplot(Uplands_long,
                aes(x = variable,
                    y = value)) +
      ylim(0, 50) +
      geom_point(shape = 17, 
                 color = "red", 
                 fill = "red",
                 size = 2) +
      theme_minimal() +                        theme(legend.position="none") +
      theme(
        axis.text.x = 
          element_text(angle = 25, 
                       vjust = 1,
                       hjust=1, 
                       size = 7)) +
      labs(title='Plotting Uplands Degradation Variables',
        subtitle='ECCC (Malloff & Shackelford, 2024)',
        caption = "Chart by Wendy Anthony \n 2025-08-30")
Uplands_long_plot
```

