---
title: "R Stats Resource Links"
author: "Collected by Wendy Anthony"
date: "2019-11-17 (updated 2020-01-09)"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
devtools::source_gist("c83e078bf8c81b035e32c3fc0cf04ee8", 
                      filename = 'render_toc.R')
```

```{r toc, echo=FALSE} 
## Table of Contents
#render_toc("R_Stats_Links.Rmd")
```

---

# Cheatsheets
[R Cheatsheets](https://github.com/rstudio/cheatsheets)

[Color Palette Cheatsheet](https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/colorPaletteCheatsheet.pdf)

*** [Spatial Cheatsheet](https://www.maths.lancs.ac.uk/~rowlings/Teaching/UseR2012/cheatsheet.html) ***

---

# R Resources
http://www.r-fiddle.org/

*** [R Resource Links](https://paulvanderlaken.com/2017/08/10/r-resources-cheatsheets-tutorials-books/) ***

*** [R Tips & Tricks](https://paulvanderlaken.com/2018/05/21/r-tips-and-tricks/) ***

[Five Tips to Improve Your R Code - Datacamp](https://www.datacamp.com/community/tutorials/five-tips-r-code-improve)

---

# Presentation

---

## R Markdown
R Markdown Cheatsheet  
https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf

[R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)

[Convert R Markdown (Rmd) Files to R Scripts](https://www.garrickadenbuie.com/blog/convert-r-markdown-rmd-files-to-r-scripts/)

[How to Write Lists Inside Markdown Table](https://stackoverflow.com/questions/19950648/how-to-write-lists-inside-a-markdown-table)

How to use objects from global environment in Rstudio Markdown  
https://stackoverflow.com/questions/34029611/how-to-use-objects-from-global-environment-in-rstudio-markdown

[Advanced R Markdown Workshop](https://arm.rbind.io/)

[R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)

*** [Iterated fact sheets with R Markdown](https://medium.com/@urban_institute/iterated-fact-sheets-with-r-markdown-d685eb4eafce) ***

https://www.r-bloggers.com/r-markdown-reports-on-demand-with-mobiletrigger/

[How to Write Reproducible Papers - Workshop](https://ulyngs.github.io/rmarkdown-workshop-2019/slides/OxfordRUsers.html#1)

*** [Intro to R Markdown](https://m-clark.github.io/Introduction-to-Rmarkdown/getting-started.html) ***

[R Markdown - Getting Started in R &amp; RStudio](http://milton-the-cat.rocks/learnr/r/r_getting_started/#section-r-markdown)

*** [RMarkdown Driven Development](https://emilyriederer.netlify.com/post/rmarkdown-driven-development/) ***

[Markdown in Jupyter Notebook](https://www.datacamp.com/community/tutorials/markdown-in-jupyter-notebook)

[Markdown Guide](https://www.markdownguide.org/getting-started/)

[rmarkdown, bookdown for authoring books](https://bookdown.org)
[blogdown for creating websites](https://github.com/rstudio/blogdown)
[articles for writing journal papers](https://github.com/rstudio/rticles)
[flexdashboard for generating dashboards](https://github.com/rstudio/flexdashboard)
[learnr for tutorials](https://github.com/rstudio/learnr)

xaringan for making slides  
https://github.com/yihui/xaringan 
https://slides.yihui.org/xaringan/#1

http://rmarkdown.rstudio.com

https://www.r-bloggers.com/vitae-dynamic-cvs-with-r-markdown/amp/

[blogdown: Creating Websites with R Markdown](https://bookdown.org/yihui/blogdown/)

[rmd fold code chunks](https://stackoverflow.com/questions/37755037/how-to-add-code-folding-to-output-chunks-in-rmarkdown-html-documents#37839683)

RefManageR::Cite() - Use with rmarkdown  
https://github.com/ropensci/RefManageR

https://libscie.github.io/rmarkdown-workshop/handout.html

https://github.com/hafen/lazyrmd

https://bookdown.org/yihui/rmarkdown/

https://www.r-bloggers.com/r-markdown-reports-on-demand-with-mobiletrigger/

https://github.com/riinuots/markdown/blob/master/instructions/instructions.md

https://rmarkdown.rstudio.com/flexdashboard/index.html

http://svmiller.com/blog/2016/03/svm-r-markdown-cv/

http://www.sthda.com/english/wiki/create-and-format-word-documents-using-r-software-and-reporters-package

https://m-clark.github.io/Introduction-to-Rmarkdown/getting-started.html

https://bookdown.org/yihui/rmarkdown/rmarkdown-site.html

https://rmarkdown.rstudio.com/articles.html

https://rmarkdown.rstudio.com/lesson-11.html

https://cmerow.github.io/RDataScience/07_03_RMarkdown_Tools.html

https://bookdown.org/yihui/rmarkdown/

https://rladiessydney.org/courses/ryouwithme/04-markymark-0/

https://github.com/jenrichmond/RMarkdownThrowdown

https://github.com/jenrichmond/RMarkdownThrowdown/blob/master/Group1_tables_contents/tableofcontents.md

https://github.com/jenrichmond/RMarkdownThrowdown/blob/master/Group1_tables_contents/make_tables.rmd

https://holtzy.github.io/Pimp-my-rmd/

https://rmarkdown.rstudio.com/gallery.html

https://statr.me/2016/08/creating-pretty-documents-with-the-prettydoc-package/  
https://prettydoc.statr.me/cayman.html  
https://prettydoc.statr.me/  
https://github.com/yixuan/prettydoc/  


*** [Intro to R Markdown](https://m-clark.github.io/Introduction-to-Rmarkdown/) ***

[HTML Document](https://bookdown.org/yihui/rmarkdown/html-document.html)

[Storyboards](https://rmarkdown.rstudio.com/flexdashboard/using.html#storyboards)

[flexdashboard: Easy interactive dashboards for R](https://rmarkdown.rstudio.com/flexdashboard/)

[R Markdown from R Studio](https://rmarkdown.rstudio.com/)

https://cran.r-project.org/web/packages/jtools/vignettes/summ.html#table_output_for_word_and_rmarkdown_documents

[package for creating slideshows with remark.js through R Markdown](https://github.com/yihui/xaringan)

*** [R Notebook](https://bookdown.org/yihui/rmarkdown/notebook.html) ***

[Bookdown](*** https://bookdown.org/home/about/)

*** https://bookdown.org/ndphillips/YaRrr/

Blogdown  
https://github.com/rstudio/blogdown

Blogdown  
https://bookdown.org/yihui/blogdown/

https://bookdown.org/yihui/rmarkdown/notebook.html

https://rmarkdown.rstudio.com/formats.html

[R Markdown Python Engine](https://rstudio.github.io/reticulate/articles/r_markdown.html)

Getting Started With R in RStudio Notebooks  
http://www.storybench.org/getting-started-r-rstudio-notebooks/

Materials for the rstudio::conf 2019 Advanced R Markdown workshop  
https://arm.rbind.io  
https://github.com/rstudio-education/arm-workshop-rsc2019

How do I add a URL to R markdown?  
https://stackoverflow.com/questions/29787850/how-do-i-add-a-url-to-r-markdown

The Best Way to Create the Website You Want from Markdown (or Jupyter/RStudio)  
https://sourcethemes.com/academic/

### Markdown in wordpress  
[Jetpack Markdown](https://jetpack.com/support/markdown/)

[Wordpress Markdown Quick Reference](https://en.support.wordpress.com/markdown-quick-reference/)

---

## Shiny
https://shiny.rstudio.com/

https://shiny.rstudio.com/articles/layout-guide.html

*** https://shiny.rstudio.com/gallery/

https://shiny.rstudio.com/gallery/superzip-example.html

https://shiny.rstudio.com/gallery/widget-gallery.html

https://matthew-j-oldach.shinyapps.io/hinuhinu/

http://www.r-gators.com/2018/04/04/creating-a-simple-shiny-app-together/

** https://github.com/ufrmeetup/gruShiny

http://www.r-gators.com/2018/03/28/introduction-to-shiny/

https://github.com/ufrmeetup/shiny-example

https://rstudio.github.io/shiny/tutorial/

https://www.showmeshiny.com/

https://www.shinyapps.io/

https://github.com/rstudio/shiny

https://ikashnitsky.shinyapps.io/us-names/

*** https://www.r-bloggers.com/r-shiny-for-beginners-annotated-starter-code/

https://jakelearnsdatascience.shinyapps.io/twitter_shiny/

*** https://michaelwreed.shinyapps.io/phila_police_map/  
https://github.com/michaelwreed/tidy_tuesday/tree/master/shiny_apps/phila_police_map

https://www.edureka.co/blog/r-shiny-tutorial/

*** https://gqueiroz.shinyapps.io/rshinylady/

https://www.r-bloggers.com/building-a-shiny-app-for-cycling-in-ottawa/

https://rviews.rstudio.com/2019/10/09/building-interactive-world-maps-in-shiny/#disqus_thread

http://enhancedatascience.com/2019/09/05/css-enhances-r-shiny-sass-enhances-css/

*** https://shinyapps.science.psu.edu/

Shiny html templates  
https://rinterface.com/  

https://github.com/nik01010/dashboardthemes  

http://shinyapps.dreamrs.fr/shinyWidgets/  

https://github.com/jienagu/DT-Editor
https://appforiarteam.shinyapps.io/DT_Editor_Example2/

https://github.com/andrewsali/shinycssloaders
https://github.com/andrewsali/shinycssloaders

https://github.com/daattali/shinyalert

https://shiny.rstudio.com/articles/validation.html

https://fontawesome.com/

https://getbootstrap.com/docs/3.3/components/

https://deanattali.com/shinyjs/

https://shiny.rstudio.com/articles/css.html

https://github.com/nanxstats/awesome-shiny-extensions

https://medium.com/@urban_institute/building-a-point-and-click-mapping-tool-in-r-shiny-7aa4e99ac913

https://rstudio.github.io/shinydashboard/index.html  
https://rstudio.github.io/shinydashboard/structure.html

https://bookdown.org/yihui/rmarkdown/shiny-documents.html

https://shiny.rstudio.com/articles/rmarkdown.html

*** https://shiny.rstudio.com/articles/interactive-docs.html

Plotly & Shiny  
https://webinars.cpsievert.me/20180220/#18

*** https://github.com/rstudio/webinars

https://github.com/yihui/shiny-apps/tree/master/face-pi
https://yihui.shinyapps.io/face-pi/

*** Shiny mobile
http://dirk.eddelbuettel.com/cranberries/cran/new/index.html
https://cran.r-project.org/web/packages/shinyMobile/index.html
https://cran.r-project.org/web/packages/shinyMobile/shinyMobile.pdf

*** https://rviews.rstudio.com/2019/10/09/building-interactive-world-maps-in-shiny/

*** HTML templates for Shiny  
http://www.rinterface.com/

*** https://www.r-bloggers.com/the-shiny-developer-series/

---

### Tmap & Leaflet shiny  
*** https://community.rstudio.com/t/integrating-tmap-into-shiny/6036

https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html#leaflet-and-shiny-integration

https://stackoverflow.com/questions/51610161/interactive-map-with-tmap-does-not-appear-in-shiny-app-but-does-show-in-rstudio

*** https://rstudio.github.io/leaflet/shiny.html

--- 

### Interactive data tables
https://glin.github.io/reactable/index.html  
https://glin.github.io/reactable/articles/cookbook/cookbook.html  
https://glin.github.io/reactable/articles/100k-rows.html  
*** https://glin.github.io/reactable/articles/shiny-demo.html  
*** https://glin.github.io/reactable/articles/cookbook/cookbook.html

---

## Jupyter Notebooks
https://www.anaconda.com/distribution/

https://www.datacamp.com/community/tutorials/tutorial-jupyter-notebook

https://jupyter.readthedocs.io/en/latest/install.html

https://jupyter.org/

---

# R Output
Get text from pdfs or images using OCR: a tutorial with {tesseract} and {magick}  
https://www.brodrigues.co/blog/2019-03-31-tesseract/

https://www.brodrigues.co/blog/2018-06-10-scraping_pdfs/

https://www.brodrigues.co/blog/2018-12-15-lubridate_africa/

*** https://www.brodrigues.co/blog/2015-02-22-export-r-output-to-file/

Tables Easily generate information-rich, publication-quality tables from R  
https://gt.rstudio.com  
https://github.com/rstudio/gt

*sVG*
https://www.stat.auckland.ac.nz/~paul/Reports/grImport/grimport-gridsvg-export/grimport-gridsvg-export.html

https://www.r-bloggers.com/2019-02-svg-in-svg-out/

## Slidify - Make slides
http://slidify.org/samples/intro/#10  
http://slidify.org/samples/intro/#1  
https://yintingchou.com/posts/ioslides-vs-slidify-in-r-markdown-presentation/  
http://ramnathv.github.io/carouselDemo/#1  
http://slidify.org/start.html  
https://www.w3.org/Talks/Tools/Slidy2/  

*** http://slidify.github.io/iquiz/#1  
https://codeload.github.com/slidify/iquiz/legacy.zip/gh-pages  
https://github.com/jchamill/jquery-quiz  
https://github.com/search?utf8=✓&q=Jquery+quiz  
https://www.jvcasillas.com/slidify_tutorial/#2  


*** https://stackoverflow.com/questions/14547069/how-to-write-from-r-to-the-clipboard-on-a-mac

*** lets you plot any points on a scatter plot and then download the corresponding data.
http://robertgrantstats.co.uk/drawmydata.html

## Sink - R console output to file
https://www.rdocumentation.org/packages/base/versions/3.6.1/topics/sink  
http://www.endmemo.com/program/R/sink.php  
https://stat.ethz.ch/R-manual/R-devel/library/base/html/sink.html  
https://statisticsglobe.com/r-sink-function-example  

https://stackoverflow.com/questions/14547069/how-to-write-from-r-to-the-clipboard-on-a-mac#14547110

---

# Spatial Analysis

*** How to build a GIF of satellite imagery in R
http://www.storybench.org/how-to-build-a-gif-of-satellite-imagery-in-r/

Dominic Roye - Visualizar el crecimiento urbano  
https://dominicroye.github.io/es/2019/visualizar-el-crecimiento-urbano/  
https://translate.google.com/translate?hl=en&sl=auto&tl=en&u=https%3A%2F%2Fdominicroye.github.io%2Fes%2F2019%2Fvisualizar-el-crecimiento-urbano%2F

*** Geospatial visualization using R
*** http://rpubs.com/bhaskarvk/userR2017-tutorial_proposal
http://rpubs.com/bhaskarvk/userR2017-tutorial_proposal

*** An introduction to R for spatial analysis and mapping (Second Edition)
https://www-tandfonline-com.ezproxy.library.uvic.ca/doi/pdf/10.1080/13658816.2019.1647541?needAccess=true

https://uk.sagepub.com/en-gb/eur/an-introduction-to-r-for-spatial-analysis-and-mapping/book241031#preview

landscapemetrics: an open‐source R tool to calculate landscape metrics
https://cran.r-project.org/web/packages/landscapemetrics/index.html
https://github.com/r-spatialecology/landscapemetrics
https://onlinelibrary.wiley.com/doi/10.1111/ecog.04617

https://github.com/r-spatial/sf/wiki/Migrating

https://github.com/r-spatial/leafem

*** https://www.spatialreference.org/

*** https://github.com/r-spatial/mapview
https://github.com/r-spatial/leafem

https://github.com/r-spatial


## R Packages - Spatial Analysis
https://github.com/spatialanalysis/geodaData

https://cran.r-project.org/web/packages/maps/index.html

https://cran.r-project.org/web/packages/maptools/index.html

https://cran.r-project.org/web/packages/rjson/index.html

https://cran.r-project.org/web/packages/DeducerSpatial/index.html

https://cran.r-project.org/web/packages/RgoogleMaps/index.html

*** Packages to explore the earth ***
CRAN Task View: Analysis of Spatial Data - Spatial Analysis related resources.  
Leaflet - One of the most popular JavaScript libraries interactive maps.  
ggmap - Plotting maps in R with ggplot2.  
REmap - R interface to the JavaScript library ECharts for interactive map data visualization.  
sf - Improved Classes and Methods for Spatial Data.  
sp - Classes and Methods for Spatial Data.  
rgeos - Interface to Geometry Engine - Open Source  
rgdal - Bindings for the Geospatial Data Abstraction Library  
maptools - Tools for Reading and Handling Spatial Objects  
gstat - Spatial and spatio-temporal geostatistical modelling, prediction and simulation.  
spacetime - R classes and methods for spatio-temporal data.  
RColorBrewer - Provides color schemes for maps  
spatstat - Spatial Point Pattern Analysis, Model-Fitting, Simulation, Tests  
spdep - Spatial Dependence: Weighting Schemes, Statistics and Models  
tigris - Download and use Census TIGER/Line shapefiles in R  
GWmodel - Geographically-Weighted Models  

---

### ggplot2

*** https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html

https://www.springer.com/us/book/9780387981413

*** https://wilkelab.org/practicalgg/

http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/

*** https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html

https://ggplot2.tidyverse.org/reference/

https://ggplot2.tidyverse.org/

https://github.com/tidyverse/ggplot2/wiki/Legend-Attributes

https://ggplot2-book.org/

R Bootcamp ggplot  
https://r-bootcamp.netlify.com/chapter1

*** https://www.garrickadenbuie.com/blog/custom-discrete-color-scales-for-ggplot2/

plothelper: New Plots Based on 'ggplot2' and Functions to Create Regular Shapes  
https://cran.r-project.org/web/packages/plothelper/index.html

*** http://zevross.com/blog/2014/08/04/beautiful-plotting-in-r-a-ggplot2-cheatsheet-3/

http://zevross.com/blog/2014/07/16/mapping-in-r-using-the-ggplot2-package/

Twitter #ggplot2 challenge
https://mohdbilal87.github.io/dataviz-r/

https://stackoverflow.com/questions/58899810/r-ggplot-force-facet-grid-margin-to-be-on-the-next-row

https://community.rstudio.com/t/how-to-get-ggplot-in-shiny-to-recognize-input-as-a-vector/46755

*** https://geocompr.github.io/post/2019/ggplot2-inset-maps/

https://medium.com/@urban_institute/why-the-urban-institute-visualizes-data-with-ggplot2-46d8cfc7ee67

https://stackoverflow.com/questions/58977671/how-to-increase-the-ggplot2-figures-resolution

https://cedricscherer.netlify.com/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r/#toc

https://simplystatistics.org/2019/08/28/you-can-replicate-almost-any-plot-with-ggplot2/

https://www.r-bloggers.com/choropleth-map-in-ggplot2/

https://ggplot2tutor.com/streetmaps/streetmaps/

https://www.datasciencecentral.com/profiles/blogs/creating-maps-in-r-using-ggplot2-and-maps-libraries

https://github.com/dkahle/ggmap

[choropleth ggplot - Google Search Results](https://www.google.com/search?q=choropleth+ggplot&oq=choropleth+ggplot&aqs=chrome..69i57j69i61j69i60.314j0j1&sourceid=chrome&ie=UTF-8)

Interactively explore your data by visualizing it with the ggplot2 package  
https://cran.r-project.org/web/packages/esquisse/readme/README.html

/////////////  
R choropleth ggplot
ggplot(data = UK) +  
aes(x = long, y = lay, group = group) +  
geom_polygon(fill = #004529, alpha = 0.75) +  
geom_ point(data = postcodes,  
aes(y = latitude, x = longitude, group = group),  
color = #f7fcb9, size = 0.5) +  
coord_map()  
/////////////

Accelerate your plots with ggforce  
https://rviews.rstudio.com/2019/09/19/intro-to-ggforce/

Chapter 10: GGplot for Scatterplots & Density Plots  
https://ademos.people.uic.edu/Chapter10.html

Boxplots and Bar Graphs  
https://ademos.people.uic.edu/Chapter11.html

---

### sf

http://www.geospatialtrainingsolutions.co.uk/spatial-r-moving-from-sp-to-sf/

https://stackoverflow.com/questions/43456524/how-to-find-which-polygon-a-point-belong-to-via-sf

https://gis.stackexchange.com/questions/282750/identify-polygon-containing-point-with-r-sf-package/318629#318629

https://gis.stackexchange.com/questions/272447/equivalent-of-sp-packages-point-in-polygon-overing-using-sf

https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html

https://stackoverflow.com/questions/58795903/row-wise-operation-with-sf#answer-58801348

https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html

---

## Intro to Spatial Analysis

[Spatial Cheatsheet](https://www.maths.lancs.ac.uk/~rowlings/Teaching/UseR2012/cheatsheet.html)

[Intro to R Spatial Analysis - Nick Bearman](https://github.com/nickbearman/intro-r-spatial-analysis)

[Intro to R Spatial Analysis- workbook - Nick Bearman](https://github.com/nickbearman/intro-r-spatial-analysis/blob/master/workbook.pdf)

*** [R Spatial on github](https://github.com/r-spatial) ***

https://cran.r-project.org/doc/contrib/intro-spatial-rl.pdf

---

## R as GIS

[R QGIS](https://github.com/r-spatial/RQGIS)

http://www.nickeubank.com/gis-in-r/

https://github.com/Pakillo/R-GIS-tutorial/blob/master/R-GIS_tutorial.md

https://www.jessesadler.com/post/gis-with-r-intro/

*** https://north-road.github.io/qgis-processing-r/ ***

---

https://www.google.ca/search?q=r+spatial+tutorial&ie=UTF-8&oe=UTF-8&hl=en-ca&client=safari

https://www.google.ca/search?q=r+gtable&ie=UTF-8&oe=UTF-8&hl=en-ca&client=safari

https://rspatial.org/raster/analysis/8-pointpat.html

https://rspatial.org/_sources/analysis/8-pointpat.R.txt _

https://channel9.msdn.com/Events/useR-international-R-User-conferences/useR-International-R-User-2017-Conference/mapedit-interactive-manipulation-of-spatial-objects

*** https://cran.r-project.org/web/views/Spatial.html ***

https://awesome-r.com/#awesome-r-spatial

https://www.r-bloggers.com/elevation-profiles-in-r/

https://geography.uoregon.edu/GeogR/topics

*** http://www.nickeubank.com/gis-in-r/ ***

*** Geocomputation with R - Robin Lovelace ***  
https://geocompr.robinlovelace.net/

https://geocompr.github.io/geocompkg/articles/us-map.html

https://geocompr.github.io/

https://geocompr.github.io/vignettes/

https://spatialanalysis.github.io/geodaData/

https://cran.r-project.org/doc/Rnews/

https://journal.r-project.org/

*** http://www.spatialanalysisonline.com/index.html ***

*** http://www.spatialanalysisonline.com/HTML/index.html ***

*** https://atlan.com/courses/introduction-to-gis-r/overview/?__hstc=78729675.f49e41239e1d36fecd0c14c710516d14.1567644668294.1567646982413.1572864193809.3&__hssc=78729675.1.1572864193809&submissionGuid=719aeeb1-0877-492e-b00e-e1900eec9a1c ***

*** https://fishandwhistle.net/post/2019/bathymetry-lake-volume-estimation-using-r/ ***

*** https://github.com/PennMUSA/MasterClass2019_3DMappingAndViz/blob/master/MusaMasterclass.md ***

*** https://geodacenter.github.io/workbook/6a_local_auto/lab6a.html#local-moran ***

A method for computing hourly, historical, terrain‐corrected microclimate anywhere on Earth  
https://besjournals.onlinelibrary.wiley.com/doi/pdf/10.1111/2041-210X.13330

---

## Remote Sensing in R

### Whitebox
https://github.com/giswqs/whiteboxR  
https://jblindsay.github.io/wbt_book/available_tools/index.html

---

## Mapping in R

### Cartography and Mapping in R
Making maps with R  
https://geocompr.robinlovelace.net/adv-map.html

How to Create State and County Maps Easily in R — Part 2  
https://medium.com/@urban_institute/how-to-create-state-and-county-maps-easily-in-r-part-2-d21d57310c50

https://github.com/Robinlovelace/Creating-maps-in-R

http://www.molecularecologist.com/2012/09/making-maps-with-r/

https://rgeomatic.hypotheses.org/1205

https://cran.r-project.org/web/packages/cartography/index.html

https://cran.r-project.org/web/packages/cartography/vignettes/cartography.html

https://stackoverflow.com/questions/55924348/how-do-i-plot-the-map-of-east-africa-kenya-uganda-and-tanzania-in-r-with-libr

https://stackoverflow.com/questions/42281443/draw-multiple-object-of-class-tmap-with-just-one-common-legend

*** https://github.com/tts/mapchallenge

*** http://www.molecularecologist.com/2012/09/making-maps-with-r/

*** https://github.com/lina2497/Giftmap

*** https://git.rud.is/hrbrmstr/y2019-30daymapchallenge

https://github.com/tts/mapchallenge/blob/master/globe.R

https://gis.stackexchange.com/questions/48828/making-beautiful-maps-in-r

https://gis.stackexchange.com/questions/3083/seeking-examples-of-beautiful-maps/45518#45518

Telling the story of Indiana lake water quality with interactive web-based mapping and data visualization  
https://github.com/corysauve/NALMS-2019

Plot US roads, coloring each by its designation (e.g. Street, Road, Avenue)  
*** https://github.com/erdavis1/RoadColors

https://www.google.ca/search?q=r+add+inset+map&ie=UTF-8&oe=UTF-8&hl=en-ca&client=safari

https://dominicroye.github.io/en/

*** https://geocompr.robinlovelace.net/adv-map.html  
*** https://geocompr.robinlovelace.net/

---

### tMap
*** https://community.rstudio.com/t/integrating-tmap-into-shiny/6036

https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html#leaflet-and-shiny-integration

https://stackoverflow.com/questions/51610161/interactive-map-with-tmap-does-not-appear-in-shiny-app-but-does-show-in-rstudio

*** https://rdrr.io/cran/tmap/man/renderTmap.html

https://statnmap.com/2019-11-15-30daymapchallenge-building-maps-2-tmap/

*** https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html

https://channel9.msdn.com/Events/useR-international-R-User-conferences/useR-International-R-User-2017-Conference/Exploring-and-presenting-maps-with-tmap

https://stackoverflow.com/questions/21269098/create-an-inset-map-within-a-larger-map-in-r

---

### Leaflet for R

https://github.com/bhaskarvk

http://rpubs.com/bhaskarvk/userR2017-tutorial_proposal

https://justinmillar.github.io/leaflet-intro/#15

https://github.com/bhaskarvk/leaflet.extras

*** https://github.com/r-spatial/leafsync

** https://github.com/r-spatial/leafpop

*** http://rstudio.github.io/leaflet/

https://learn.r-journalism.com/en/mapping/leaflet_maps/leaflet/

https://www.google.ca/search?q=leaflet+r&ie=UTF-8&oe=UTF-8&hl=en-ca&client=safari

*** https://rstudio.github.io/leaflet/shiny.html

Bhaskar V. Karambelkar  - leaflet
https://www.karambelkar.info/  
https://github.com/bhaskarvk  
https://github.com/bhaskarvk?tab=repositories

*** https://www.jla-data.net/eng/leaflet-in-r-tips-and-tricks/index.html  

https://digitalfellows.commons.gc.cuny.edu/2019/11/21/interactively-explore-geographic-data-in-r-using-leaflet/

---

### Choropleth Maps in R
https://rstudio-pubs-static.s3.amazonaws.com/324400_69a673183ba449e9af4011b1eeb456b9.html

https://unconj.ca/blog/choropleth-maps-with-r-and-ggplot2.html

https://rpubs.com/thaufas/375375

https://cengel.github.io/R-spatial/mapping.html

https://www.rdocumentation.org/packages/GISTools/versions/0.7-4/topics/choro.legend

https://www.r-bloggers.com/custom-legend-in-r/

https://stackoverflow.com/questions/32749889/make-a-2d-legend-for-a-plot-bi-variate-choropleth-maps

https://www.r-graph-gallery.com/183-choropleth-map-with-leaflet.html

http://www.alex-singleton.com/GDS_UA_2017/Mapping_Areas_and_Context/

https://www.rdocumentation.org/packages/graphics/versions/3.3/topics/par

---

### Shapefiles in R
*** [Creating GeoJSON out of Shapefile in R](https://blog.exploratory.io/creating-geojson-out-of-shapefile-in-r-40bc0005857d) ***

Shapefiles in R   
https://www.r-bloggers.com/shapefiles-in-r/

How to map point data and polygon shapefiles in R  
http://www.storybench.org/how-to-map-point-data-and-polygon-shapefiles-in-r/

[An Introduction to R for Spatial Analysis and Mapping - Google Books](https://books.google.ca/books?id=zbRkDwAAQBAJ&printsec=frontcover#v=onepage&q&f=false)

---

## Data Frames
R Data Frame: Create, Append, Select, Subset  
https://www.guru99.com/r-data-frames.html

https://stackoverflow.com/questions/58243042/how-to-subset-one-dataframe-and-make-it-equal-to-corresponding-dataframes-in-r

R: Add new column based on condition with multiple columns  
https://stackoverflow.com/questions/58696140/r-add-new-column-based-on-condition-with-multiple-columns

*** https://stackoverflow.com/questions/58833840/making-a-plot-with-geom-lines-of-my-data-frame ***

Introduction to the Subset function  
https://ihaddadenfodil.com/post/introduction_to_the_subset_function/introduction-to-the-subset-function/

---

## Density - KDE
https://vita.had.co.nz/papers/density-estimation.pdf

https://stat.ethz.ch/R-manual/R-devel/library/stats/html/density.html

https://github.com/mountainMath/dotdensity

---

## Cluster Analysis
Cluster Analysis (DataCamp)  
https://www.statmethods.net/advstats/cluster.html

https://data.cdrc.ac.uk/tutorial/creating-a-geodemographic-classification-using-k-means-clustering-in-r

---

## Spatial autocorrelation
https://rspatial.org/raster/rosu/Chapter7.html

https://rspatial.org/raster/rosu/index.html

https://github.com/spatialanalysis/geodaDatav

https://spatialanalysis.github.io/geodaData/

https://github.com/lixun910/rgeoda

*** https://cengel.github.io/R-spatial/ ***

https://www.gislounge.com/free-introduction-to-gis-in-r-online-course/

*** https://geocompr.robinlovelace.net/eco.html#exercises-10 ***

*** https://github.com/r-spatial ***

https://github.com/r-spatial/RSAGA

*** https://github.com/tylerjtran/populationLines/blob/master/populationLinesScript.R ***

https://spatial.ly/2017/04/population-lines-how-and-why-i-created-it/

https://www.datacamp.com/community/tutorials/r-formula-tutorial

https://www.datacamp.com/community/tutorials/r-web-scraping-rvest

https://www.r-spatial.org/

*** https://geocompr.robinlovelace.net/index.html ***

https://github.com/MicheleTobias/R-Projections-Workshop

*** https://cran.r-project.org/doc/contrib/intro-spatial-rl.pdf ***

https://cengel.github.io/rspatial/

https://github.com/nickbearman/intro-r-spatial-analysis

https://github.com/nickbearman/r-geodemographic-analysis-20140710

---

## Regression Analysis & Modeling
https://bookdown.org/ndphillips/YaRrr/regression-analysis.html

*** https://www.r-bloggers.com/r-tutorial-series-simple-linear-regression/

*** http://www.cookbook-r.com/Statistical_analysis/Regression_and_correlation/ ***

https://github.com/matackett/talks/blob/master/presentations/wsds2019-r-git-regression.pdf

http://www.cookbook-r.com/Statistical_analysis/Regression_and_correlation/

*** http://www.r-tutor.com/elementary-statistics/simple-linear-regression

Simple linear regression  
http://rtutorialseries.blogspot.com/2009/11/r-tutorial-series-simple-linear.html

multiple linear regression  
http://rtutorialseries.blogspot.com/2009/12/r-tutorial-series-multiple-linear.html

Graphic Analysis of Regression Assumptions  
http://rtutorialseries.blogspot.com/2009/12/r-tutorial-series-graphic-analysis-of.html

Regression With Interaction Variables  
http://rtutorialseries.blogspot.com/2010/01/r-tutorial-series-regression-with.html

Regression With Categorical Variables  
http://rtutorialseries.blogspot.com/2010/02/r-tutorial-series-regression-with.html

Basic Polynomial Regression  
http://rtutorialseries.blogspot.com/2010/02/r-tutorial-series-basic-polynomial.html

Hierarchical Linear Regression  
http://rtutorialseries.blogspot.com/2010/01/r-tutorial-series-basic-hierarchical.html

Model basics  
*** https://r4ds.had.co.nz/model-basics.html

*** https://www.r-bloggers.com/r-tutorial-series-simple-linear-regression/

*** https://www.r-bloggers.com/how-to-perform-a-logistic-regression-in-r/

*** https://data-flair.training/blogs/r-linear-regression-tutorial/

https://socialsciences.mcmaster.ca/jfox/Books/Companion-1E/index.html

https://socialsciences.mcmaster.ca/jfox/Books/Companion/downloads.html

Fitting Linear Models  
https://socialsciences.mcmaster.ca/jfox/Books/Companion/scripts/chap-4.R

*** Statistical modelling 1: Simple linear models  
https://jacintak.github.io/biostats/biostats-GLM1.html  
https://jacintak.github.io/biostats/biostats-GLM2.html  
https://jacintak.github.io/biostats/biostats-GLM2.html

https://jacintak.github.io/colRoz/

---

## Spatial Data
[BC Data - An R package for searching & retrieving data from the B.C. Data Catalogue](https://github.com/bcgov/bcdata)

---

## Time Series Analysis

https://ademos.people.uic.edu/Chapter23.html

** 14.1 Representing Time Series Data R cookbook pg 356  
https://rc2e.com/timeseriesanalysis

Time Series and Forecasting  
https://www.statmethods.net/advstats/timeseries.html

https://cran.r-project.org/web/views/TimeSeries.html

https://www.analyticsvidhya.com/blog/2015/12/complete-tutorial-time-series-modeling/

https://livebook.manning.com/book/r-in-action-third-edition/chapter-3/v-1/1

https://www.statmethods.net/advstats/timeseries.html

https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html

https://buildmedia.readthedocs.org/media/pdf/a-little-book-of-r-for-time-series/latest/a-little-book-of-r-for-time-series.pdf

http://r-statistics.co/Time-Series-Analysis-With-R.html

https://www.manning.com/books/r-in-action-second-edition?a_bid=5c2b1e1d&a_aid=RiA2ed

*** http://www.cookbook-r.com/Scripts_and_functions/Measuring_elapsed_time/

https://stackoverflow.com/questions/13073686/interpolating-timeseries

http://www.r-gators.com/2018/01/31/wildlife-tracking-data-in-r/

https://www.r-bloggers.com/dangerous-streets-of-bratislava-animated-maps-using-open-data-in-r/

https://fishandwhistle.net/post/2019/depth-time-heatmaps/

https://www.datacamp.com/courses/introduction-to-time-series-analysis

https://www.datacamp.com/community/tutorials/time-series-r

https://github.com/oscarperpinan/spacetime-vis

https://www.crcpress.com/Displaying-Time-Series-Spatial-and-Space-Time-Data-with-R-Second-Edition/Lamigueiro/p/book/9781138089983#googlePreviewContainer

** Cluster multiple time series using K-means  
https://www.brodrigues.co/blog/2019-10-12-cluster_ts/

*** Apps for creating Landsat time series animations  
https://jdbcode.github.io/Snazzy-EE-TS-GIF/

*** https://www.r-bloggers.com/time-series-forecasting-with-random-forest/amp/#click=https://t.co/3yM9AoHLCl

---

## Census
https://github.com/hrecht/censusapi

https://github.com/walkerke/tidycensus

https://www.census.gov/programs-surveys/geography.html

*** https://github.com/mountainMath?tab=repositories

*** https://github.com/mountainMath/cancensus

https://mountainmath.github.io/cancensus/index.html

https://mountainmath.github.io/cancensus/

https://mountainmath.github.io/cancensus/articles/cancensus.html

https://cran.r-project.org/web/packages/cancensus/vignettes/cancensus.html

https://censusmapper.ca/

*** https://onlineacademiccommunity.uvic.ca/mreumcensusdatatutorial/

https://github.com/walkerke/tidycensus

https://github.com/hrecht/censusapi

*** https://news.communitech.ca/how-to-access-statscan-data-through-their-new-api/

https://github.com/mountainMath/every_city_canada/blob/master/R/every_city_canada.R

https://wizard.unbc.ca/record=e1000099%7ES3

*** https://github.com/mountainMath/cansim

https://cran.r-project.org/web/packages/CANSIM2R/CANSIM2R.pdf

[Count trees per census block NY](https://mattherman.info/blog/point-in-poly/)

---

# Color palettes
Comprehensive list of color palettes available in r  
https://github.com/EmilHvitfeldt/r-color-palettes

Paletteer: Collection of most color palettes in a single R package  
https://github.com/EmilHvitfeldt/paletteer

Palettes {grDevices} Create a vector of n contiguous colors  
https://stat.ethz.ch/R-manual/R-devel/library/grDevices/html/palettes.html

The viridis color palettes  
https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html

Using colors in R - Jenny Bryan
https://www.stat.ubc.ca/~jenny/STAT545A/block14_colors.html

a California-inspired package of color palettes  
https://github.com/an-bui/calecopal

https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html

https://cran.r-project.org/web/packages/unikn/vignettes/colors.html

https://github.com/EmilHvitfeldt/r-color-palettes

http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf

*** Top R Color Palettes To Know For Great Data Visualization  
https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/

https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/colorPaletteCheatsheet.pdf

https://www.highcharts.com/docs/export-module/client-side-export

https://jsfiddle.net/gh/get/jquery/1.11.0/highslide-software/highcharts.com/tree/master/samples/highcharts/exporting/offline-download-demo/

*** Custom palette from picture  
https://github.com/AndreaCirilloAC/paletter

https://github.com/aljrico/harrypotter

https://github.com/dill/beyonce

https://nschiett.github.io/fishualize/index.html…

https://github.com/karthik/wesanderson

https://github.com/leonawicz/legocolors

https://vankesteren.github.io/rijkspalette/

https://github.com/bhaskarvk/colormap

https://stackoverflow.com/questions/15282580/how-to-generate-a-number-of-most-distinctive-colors-in-r

https://github.com/nschiett/fishualize

https://nschiett.github.io/fishualize/articles/overview_colors.html

R package to generate colors from a list of 44 pre-defined palettes  
https://github.com/bhaskarvk/colormap  
https://bhaskarvk.github.io/colormap/

EVALUATION OF THE NEW PALETTE() FOR R  
https://nowosad.github.io/post/cbc-bp1/  
https://developer.r-project.org/Blog/public/2019/11/21/a-new-palette-for-r/index.html  
https://github.com/EmilHvitfeldt/quickpalette

Scales  
https://canvas.instructure.com/courses/946353/pages/scales?module_item_id=7585934

https://serialmentor.com/dataviz/color-pitfalls.html

Automatic Generation of Qualitative Color Palettes  
https://jolars.github.io/qualpalr/

http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf

---

# R Books On-line

[Advanced R - Hadley Wickham](https://adv-r.hadley.nz/)

[R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)

[R in Action](https://livebook.manning.com/book/r-in-action-third-edition/chapter-3/v-1/1)

[Intro to R - course booklet](https://www.bioinformatics.babraham.ac.uk/training/Introduction_to_R/Introduction%20to%20R%20course%20booklet.pdf)

[An Introduction to R for Spatial Analysis and Mapping - Google Books](https://books.google.ca/books?id=zbRkDwAAQBAJ&printsec=frontcover#v=onepage&q&f=false)

*** [Beyond Spreadsheets with R](https://www.manning.com/books/beyond-spreadsheets-with-r) ***

[9 Visualizing data: Plotting](https://livebook.manning.com/book/beyond-spreadsheets-with-r/chapter-9/256)

[R for Dummies](http://sgpwe.izt.uam.mx/files/users/uami/gma/R_for_dummies.pdf)

[R for Data Science by Hadley Wickham and Garrett Grolemund](https://r4ds.had.co.nz/)

[Getting started in R and RStudio](http://milton-the-cat.rocks/learnr/r/r_getting_started/)

*** Geocomputation with R - Robin Lovelace ***  
https://geocompr.robinlovelace.net/

The Art of R Programming  
http://diytranscriptomics.com/Reading/files/The%20Art%20of%20R%20Programming.pdf

A Language, not a Letter: Learning Statistics in R  - 27 chapters  
https://ademos.people.uic.edu/index.html

What They Forgot to Teach You About R - Jennifer Bryan  
https://rstats.wtf/

An Introduction to Statistical Learning with Applications in R  
http://faculty.marshall.usc.edu/gareth-james/ISL/  
http://faculty.marshall.usc.edu/gareth-james/ISL/ISLR%20Seventh%20Printing.pdf

YaRrr! The Pirate’s Guide to R  
https://bookdown.org/ndphillips/YaRrr/

A Handbook of Statistical Analyses Using R  
http://www.ecostat.unical.it/tarsitano/Didattica/LabStat2/Everitt.pdf

---

## R Cookbook
[R Graphics Cookbook](http://www.cookbook-r.com/)

R Graphics Cookbook, 2nd edition  
https://r-graphics.org/

[Manipulating Data - R Cookbook](http://www.cookbook-r.com/Manipulating_data/)

*** [Debugging a script or function - R Cookbook](http://www.cookbook-r.com/Scripts_and_functions/Debugging_a_script_or_function/)

*** [Getting a list of functions and objects in a package - R Cookbook](http://www.cookbook-r.com/Scripts_and_functions/Getting_a_list_of_functions_and_objects_in_a_package/)

[Output to a file - R Cookbook](http://www.cookbook-r.com/Graphs/Output_to_a_file/)

[Fonts - R Cookbook](http://www.cookbook-r.com/Graphs/Fonts/)

[Graphs - R Cookbook](http://www.cookbook-r.com/Graphs/)

[Manipulating Data - R Cookbook](http://www.cookbook-r.com/Manipulating_data/)

[Variables - R Cookbook](http://www.cookbook-r.com/Basics/Information_about_variables/)

[Working with NULL and NA - R Cookbook](http://www.cookbook-r.com/Basics/Working_with_NULL_NA_and_NaN/)

---

# R Stats Mentors

---

## Hadley Wickham Resources
[Advanced R - Hadley Wickham](https://adv-r.hadley.nz/)

[R for Data Science, Garrett Grolemund, Hadley Wickham](https://r4ds.had.co.nz/)

*** https://adv-r.hadley.nz/translation.html

*** https://github.com/hadley

https://adv-r.hadley.nz/

*** https://github.com/hadley/stats337

[Video of Hadley Wickham’s Keynote address at rstudio::conf 2017](https://www.rstudio.com/resources/videos/data-science-in-the-tidyverse/)

[Video /slides of Hadley Wickham’s Keynote address at rstudio::conf 2017](https://github.com/rstudio/rstudio-conf/blob/master/2017/The_Tidyverse-Hadley_Wickham/tidyverse.pdf)

[Style Guide - Hadley Wickham](http://adv-r.had.co.nz/Style.html)

The tidyverse style guide - Hadley Wickham  
*** https://style.tidyverse.org/

---

## Jenny Bryan Resources
https://github.com/jennybc  
https://jennybryan.org/code/  
https://stat545.com/

What They Forgot to Teach You About R - Jennifer Bryan  
https://rstats.wtf/

---

# Climate & Weather
GWEX: Multi-Site Stochastic Models for Daily Precipitation and Temperature  
- https://cran.r-project.org/web/packages/GWEX/index.html  
- https://www.rdocumentation.org/packages/GWEX/versions/1.0.2/source  
- https://www.rdocumentation.org/packages/GWEX/versions/1.0.2  
- https://rdrr.io/cran/GWEX/src/R/myParPrecGWEX-data.r


---

# R & Biodiversity

Tree ring data  
https://github.com/AndyBunn/dplR

http://remote-sensing.eu/new-r-package-to-visualize-remote-sensing-data-and-movement-tracks-movevis/  
https://www.google.ca/search?client=safari&hl=en-ca&sxsrf=ACYBGNSYGBtwRssLcvZP7cbbqaYP7PPYbw%3A1573987721748&ei=iSXRXfyVLYGd-gTqo6bgBA&q=R+moveVis&oq=R+moveVis&gs_l=mobile-gws-wiz-serp.3..35i39j0j0i8i30.7461.8036..8509...0.1..0.84.238.3......0....1.........0i71j35i304i39j0i7i30j0i8i7i30.irXMirHqdoY

http://animove.org/  
http://animove.org/software-and-data/

---

## iNaturalist

https://github.com/WendyAnthony/rinat

*** https://cran.r-project.org/web/packages/rinat/vignettes/rinatVignette.html

https://github.com/ropensci/rinat

https://docs.ropensci.org/rinat/

https://www.inaturalist.org/pages/api+reference

https://api.inaturalist.org/v1/docs/

https://ropensci.org/blog/2014/03/26/rinat/

https://rdrr.io/cran/rinat/man/get_inat_taxon_stats.html

https://www.rdocumentation.org/packages/rinat/versions/0.1.4

https://inaturalist.nz/pages/download-data-nz

https://rdrr.io/cran/rinat/man/get_inat_user_stats.html

---

## Plantnet biodiversity app  
*** https://github.com/plantnet  
https://plantnet.org/en/

---

## Phenocam
https://www.neonscience.org/phenocam-intro-series

---

## R Openscience
http://tuijasonkkila.fi/blog/2014/01/mapping-red-listed-rainforest-tree-species/

https://ropensci.org/packages/

https://github.com/ropensci/rebird

https://github.com/ropensci/geojsonio

https://github.com/ropensci/rgbif

https://github.com/ropensci/pdftools

https://github.com/ropenscilabs/syn

*** https://github.com/OpenScienceMOOC/Module-3-Reproducible-Research-and-Data-Analysis/blob/master/key_elements.md

https://github.com/OpenScienceMOOC/Module-5-Open-Research-Software-and-Open-Source/tree/master/content_development

MtreeRing: An R package with graphical user interface for automatic measurement of tree ring widths using image processing techniques  
https://www.sciencedirect.com/science/article/pii/S1125786519301006?dgcid=raven_sd_aip_email  
https://cran.r-project.org/web/packages/MtreeRing/index.html  
https://github.com/ropensci/MtreeRing  
https://ropensci.github.io/MtreeRing/  

---

# Open Air - Pollution
*** Tools for the Analysis of Air Pollution Data
air pollution, hourly time series data  
https://rdrr.io/cran/openair/man/openair.html  
https://cran.r-project.org/web/packages/openair/index.html  
https://cran.r-project.org/web/packages/openair/openair.pdf  
http://dirk.eddelbuettel.com/cranberries/cran/updated/index.html  
https://r-forge.r-project.org/projects/openair/  
http://www.openair-project.org/  
http://davidcarslaw.github.io/openair/  
https://github.com/davidcarslaw/openair  

---

# R Stats
https://stat545.com/  
https://stat545.stat.ubc.ca/  
https://stat545guidebook.netlify.com/  

http://www.r-tutor.com/elementary-statistics

https://rfortherestofus.com/2019/10/how-to-make-functions-in-r/

https://medium.com/analytics-vidhya/advanced-data-wrangling-in-r-4-f98693b92851

https://medium.com/analytics-vidhya/group-manipulation-in-r-3-5554a0c1b544?

https://medium.com/analytics-vidhya/statistical-visualization-in-r-2-9e9070bcdecf?

https://medium.com/analytics-vidhya/many-ways-of-reading-data-into-r-1-52b02825cb27?

---

# Data Prep

*** Introduction to Spatial Data Types in R  
http://www.rpubs.com/cengel248/172289

https://www.computerworld.com/article/2921176/great-r-packages-for-data-import-wrangling-visualization.html

** https://geanders.github.io/RProgrammingForResearch/

*** https://geanders.github.io/RProgrammingForResearch/entering-and-cleaning-data-1.html

4 data wrangling tasks in R for advanced beginners  
https://www.computerworld.com/article/2486425/business-intelligence-4-data-wrangling-tasks-in-r-for-advanced-beginners.html

*** http://milton-the-cat.rocks/learnr/r/r_getting_started/#section-getting-data-into-rstudio

https://ihaddadenfodil.com/post/three_warys_to_get_a_summary_of_your_data/three-ways-to-get-a-summary-of-your-data/

*** https://www.computerworld.com/article/2921176/great-r-packages-for-data-import-wrangling-visualization.html

*** https://geanders.github.io/RProgrammingForResearch/entering-and-cleaning-data-1.html

*** https://www.r-bloggers.com/explore-your-dataset-in-r/

https://github.com/lgellis/MiscTutorial/blob/master/simple_eda/exploreData.Rmd

*** https://www.r-bloggers.com/renaming-all-files-in-a-folder-in-r/

https://www.brodrigues.co/blog/2018-02-16-importing_30gb_of_data/

Best Practices for Using Google Sheets in Your Data Project  
https://matthewlincoln.net/2018/03/26/best-practices-for-using-google-sheets-in-your-data-project.html

## Set seed for random  
https://stats.stackexchange.com/questions/86285/random-number-set-seedn-in-r

## Dates
https://stackoverflow.com/questions/58605726/convert-character-column-to-date-column-in-lubridate

https://community.rstudio.com/t/how-to-write-code-for-getting-start-date-and-end-date/44302/2

https://www.r-bloggers.com/date-formats-in-r/

---

## Na  
https://stackoverflow.com/questions/58547689/eliminate-nas-within-each-group

---

# Python and R
https://m.youtube.com/watch?v=pjDGB9i7kyI

https://r4ds.had.co.nz/

[R Markdown Python Engine](https://rstudio.github.io/reticulate/articles/r_markdown.html)

[R Interface to Python ](https://github.com/rstudio/reticulate)

*** How to track and collaborate on R experiments with Neptune  
https://medium.com/neptune-ml/how-to-track-and-collaborate-on-r-experiments-with-neptune-85d4b8c19596

---

---

# HTML Widgets
*** http://www.htmlwidgets.org/

http://www.htmlwidgets.org/showcase_dygraphs.html

http://www.htmlwidgets.org/showcase_highcharts.html

http://jkunst.com/highcharter/

http://rich-iannone.github.io/DiagrammeR/

---

---

# R Tutorials

Customizing R Startup  
https://www.statmethods.net/interface/customizing.html  

Tutorials for learning R  
*** https://www.r-bloggers.com/how-to-learn-r-2/

*** https://cran.r-project.org/manuals.html

https://cran.r-project.org/doc/manuals/r-release/R-intro.pdf

https://stat545.com/r-basics.html

https://data-flair.training/blogs/learn-r-programming/

https://data-flair.training/blogs/debugging-in-r-programming/

https://data-flair.training/blogs/descriptive-statistics-in-r/

https://data-flair.training/blogs/data-structures-in-r/

Learning R in Seven Simple Steps (DataCamp)  
https://www.datasciencecentral.com/profiles/blogs/learning-r-in-seven-simple-steps

R Tutorial Series: R Beginner’s Guide and R Bloggers Updates  
https://www.r-bloggers.com/r-tutorial-series-r-beginners-guide-and-r-bloggers-updates/

R Tutorial  
https://www.tutorialspoint.com/r/index.htm

swirl teaches you R programming and data science interactively, at your own pace, and right in the R console!  
https://swirlstats.com

swirl_courses  
https://github.com/swirldev/swirl_courses#swirl-courses

https://research.csiro.au/software/r-workshop-notes/

Digital Storytelling With R  
http://www.storybench.org/tag/rstats/

https://github.com/ines/course-starter-r

https://masalmon.eu/2017/12/11/goodrpackages/#short-tutorials

*** https://www.infoworld.com/article/3411819/do-more-with-r-video-tutorials.html

*** Interactive Tutorials for R  
https://rstudio.github.io/learnr/?utm_source=recordnotfound.com

A Guide to Getting International Statistics into R  
https://erikgahner.dk/2019/a-guide-to-getting-international-statistics-into-r/

Statistics with R  
http://zoonek2.free.fr/UNIX/48_R/all.html

Tutorials  
https://www.diigo.com/profile/cengel/R?type=all&sort=updated

R Bootcamp  
https://r-bootcamp.netlify.com/

Free online courses, tutorials and tools - Learn to use earth science and other data in R & Python  
*** https://www.earthdatascience.org/

RStudio Cloud Primers  
*** https://rstudio.cloud/learn/primers

adventr: An Adventure in Statistics interactive tutorials  
*** http://milton-the-cat.rocks/home/adventr.html

An Adventure in Statistics  
*** https://www.discoveringstatistics.com/statistics-hell-p/

Graphical Capabilities of R  
https://hub.packtpub.com/graphical-capabilities-r/

How to write the first for loop in R  
https://www.r-bloggers.com/how-to-write-the-first-for-loop-in-r/

How to build a website with Blogdown in R  
*** http://www.storybench.org/how-to-build-a-website-with-blogdown-in-r/

How to access APIs in R  
http://www.storybench.org/how-to-access-apis-in-r/

Scraping HTML tables and downloading files with R  
http://www.storybench.org/scraping-html-tables-and-downloading-files-with-r/

---

## Quick R (DataCamp)  
https://www.statmethods.net/about/sitemap.html  
https://www.statmethods.net/index.html  

---

## R Studio
R Studio increase Font size  
https://community.rstudio.com/t/font-size-files-plot-packages-help-viewer-tabs/41583

RStudio Addin: Replace in Files  
https://www.r-bloggers.com/rstudio-addin-replace-in-files/

---

## R Environment
R Environment and Scope  
https://www.datamentor.io/r-programming/environment-scope/

How to use objects from global environment in Rstudio Markdown  
https://stackoverflow.com/questions/34029611/how-to-use-objects-from-global-environment-in-rstudio-markdown

---

## R Conference Shares
University of Florida  
https://github.com/ufrmeetup

useR! 2019 - Welcome - Keynote: Julia Stewart Lowndes  
https://m.youtube.com/watch?feature=youtu.be&start=2710&v=Z8PqwFPqn6Y

useR! 2019 - Keynote: Julie Josse  
https://m.youtube.com/watch?feature=youtu.be&start=1816&v=z8IuuDe5oXs

---

# Data Viz
https://www.datacamp.com/community/tutorials/visualize-data-vim-package

*** https://www.datacamp.com/community/tutorials/data-visualization-highcharter-r

Raincloud plots: a multi-platform tool for robust data visualization  
https://wellcomeopenresearch.org/articles/4-63/v1

Customizing Graphics and Creating a Bar Chart and Scatterplot in R  
https://hub.packtpub.com/customizing-graphics-and-creating-bar-chart-and-scatterplot-r/

https://github.com/clauswilke/dataviz

https://serialmentor.com/dataviz/  
https://serialmentor.com/dataviz/geospatial-data.html

---

# ETC ...

Version control  
https://peerj.com/preprints/3159/

Tidy data  
https://www.jstatsoft.org/article/view/v059i10/v59i10.pdf

R Ladies Slack Community  
https://rladies-community-slack.herokuapp.com/

I “only” use R for descriptive stats — and that’s OK  
https://rfortherestofus.com/2018/12/descriptive-stats-r/

---

Query openrouteservice from R  
https://maps.openrouteservice.org/  
https://giscience.github.io/openrouteservice-r/articles/openrouteservice.html#optimization

*** write r-blog post without having a blog  
http://r-posts.com/

Organizing, Clarifying and Communicating the R Data Analyses  
https://hub.packtpub.com/organizing-clarifying-and-communicating-r-data-analyses/

Downloading Fitbit Data Histories with R  
https://www.r-bloggers.com/downloading-fitbit-data-histories-with-r-2/

Draw your Data and Download  
http://drawdata.xyz/

https://www.r-bloggers.com/a-full-rstudio-server-setup-for-data-science-in-5-minutes/

Vectorize In R  
https://wetlandscapes.com/

Topics in Data Visualization > Modules  
https://canvas.instructure.com/courses/946353/modules#module_962002

---

## Music
https://leonawicz.github.io/tabr/

https://community.rstudio.com/t/using-instead-of/43255

---

## Audio
*** Audio from graph
https://coolbutuseless.github.io/2019/10/04/introducing-devoutaudio-an-experimental-package-for-rendering-graphics-as-audio/

---

## Weather
*** https://cran.r-project.org/web/packages/weathercan/index.html

https://www.google.ca/search?client=safari&hl=en-ca&sxsrf=ACYBGNRGYI_on_MQmpJ7GEwSM05gNnRdAQ%3A1573946634282&ei=CoXQXdLuEJC_0PEPzLKiyAE&q=analysis+of+weather+time+series+in+r&oq=analysis+of+weather+time+series+in+r&gs_l=mobile-gws-wiz-serp.3...7501.9948..11336...0.1..0.102.643.7j1......0....1.........0i71j35i304i39j0i13i5i30j0i8i13i30j30i10.6mdD7eNyVIQ

*** https://github.com/ropensci/weathercan

http://ropensci.github.io/weathercan/

---

## CV
[Building a data-driven CV with R](https://livefreeordichotomize.com/2019/09/04/building_a_data_driven_cv_with_r/)

[CV built using RMarkdown and pagedown package- N Strayer](https://github.com/nstrayer/cv)

[CV - Sam Abbott](https://github.com/seabbs/cv)

Templates and tools for making a Resume CV  
https://pkg.mitchelloharawild.com/vitae/

*** https://livefreeordichotomize.com/2019/09/04/building_a_data_driven_cv_with_r/

https://ccgilroy.github.io/resume/

https://github.com/mitchelloharawild/vitae

--- 

## R Help

R Search
https://rseek.org/

R Search Engine - DSC Data Science Central  
https://www.datasciencecentral.com/page/search?q=R&cx=007889287169707156432%3A04zouenp6u0&ie=UTF-8&cof=FORID%3A11%3BNB%3A1&sa=Search

R package? GPS time Is it day or night?  
https://twitter.com/lilybentley/status/1194573171220996097?s=21

https://stackoverflow.com/questions/58605034/increase-plot-layout-and-reduce-the-legend-list

R not reading in projection info on Mac OSX  
https://gis.stackexchange.com/questions/19064/opening-shapefile-in-r

Diagnosing a non-responsive RStudio console  
https://community.rstudio.com/t/diagnosing-a-non-responsive-rstudio-console/4601/9

---

## Twitter
*** https://happygitwithr.com/

https://twitter.com/lucystats/status/1193964858007523328?s=21

https://www.rdocumentation.org/packages/twitteR/versions/1.1.9

http://www.storybench.org/get-twitter-data-rtweet-r/

R package? GPS time Is it day or night?  
https://twitter.com/lilybentley/status/1194573171220996097?s=21

https://twitter.com/jasminehanye/status/1194282511012171777?s=21

21 Recipes for Mining Twitter Data with rtweet  
https://rud.is/books/21-recipes/

---

# R Packages

*** https://www.haktansuren.com/installing-r-package-fixing-package-xxx-is-not-available-for-r-version-x-y-z-warning/

https://rfortherestofus.com/2019/10/how-to-make-functions-in-r/

https://medium.com/analytics-vidhya/advanced-data-wrangling-in-r-4-f98693b92851

https://medium.com/analytics-vidhya/group-manipulation-in-r-3-5554a0c1b544?

https://medium.com/analytics-vidhya/statistical-visualization-in-r-2-9e9070bcdecf?

https://medium.com/analytics-vidhya/many-ways-of-reading-data-into-r-1-52b02825cb27?

https://stackoverflow.com/questions/25721884/how-should-i-deal-with-package-xxx-is-not-available-for-r-version-x-y-z-wa

https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html

## Details
https://www.hvitfeldt.me/pkgcal2019/3/  
https://yonicd.github.io/details/articles/index.html  
https://yonicd.github.io/details/articles/custom.html  
https://yonicd.github.io/details/articles/sessioninfo.html

*** https://www.hvitfeldt.me/packagecalendar/2019/  
https://www.hvitfeldt.me/packagecalendar/2018/

*** https://hilaryparker.com/2014/04/29/writing-an-r-package-from-scratch/

https://hilaryparker.com/2014/05/27/sunsets-in-google-calendar-using-r/

Installing R packages  
https://www.r-bloggers.com/installing-r-packages/

*** List r packages installed before reinstalling 
```{r}
# Before re-installing R, save a list of all your packages (otherwise you'll have to do it by hand)  
#packs <- installed.packages()  
#packs <- mypacks[c(1:2),]  
# save packs as .csv or other format
# reload .csv
#install.packages(packs[,1])
```
*** {beepr} for victory noises when your scripts complete  
https://cran.r-project.org/web/packages/beepr/index.html  
https://www.r-project.org/nosvn/pandoc/beepr.html  
https://cran.r-project.org/web/packages/beepr/beepr.pdf

[The drake R Package User Manual](https://books.ropensci.org/drake/index.html)

learnr  
https://blog.rstudio.com/2017/07/11/introducing-learnr/

CRAN packages  
http://dirk.eddelbuettel.com/cranberries/cran/updated/index.html

https://cran.r-project.org/web/packages/janeaustenr/index.html

https://stat.ethz.ch/R-manual/R-devel/library/base/html/system.html  
https://www.rdocumentation.org/packages/base/versions/3.6.1/topics/system  
https://stat.ethz.ch/R-manual/R-devel/library/base/html/system2.html

https://swirlstats.com/students.html
https://github.com/swirldev/swirl_courses#swirl-courses


## Tables
https://cran.r-project.org/web/packages/expss/index.html  
https://cran.r-project.org/web/packages/expss/expss.pdf  

https://cran.r-project.org/web/packages/huxtable/index.html  
https://cran.r-project.org/web/packages/huxtable/huxtable.pdf

https://cran.r-project.org/web/packages/polite/index.html  
https://cran.r-project.org/web/packages/polite/polite.pdf

https://cran.r-project.org/web/packages/searcher/searcher.pdf

*** Did you know, using "searcher" package, you could automatically to search stackoverflow , google, GitHub and many more sites for errors , packages or topics


RSQLite: 'SQLite' Interface for R  
https://cran.r-project.org/web/packages/RSQLite/

*** The R Commander: A Basic-Statistics GUI for R
https://socialsciences.mcmaster.ca/jfox/Misc/Rcmdr/

Rayshader  
https://github.com/PennMUSA/MasterClass2019_3DMappingAndViz/blob/master/MusaMasterclass.md

https://cran.r-project.org/web/packages/beepr/beepr.pdf

xkcd: Plotting ggplot2 Graphics in an XKCD Style  
https://cran.r-project.org/web/packages/xkcd/index.html

*** curated list of awesome R packages, frameworks and software  
https://awesome-r.com/

API client for the Internet Archive  
https://lincolnmullen.com/software/internetarchive/

ggtree: an r package for visualization and annotation of phylogenetic trees with their covariates and other associated data  
https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12628

assign {base}	- Assign a Value to a Name  
https://stat.ethz.ch/R-manual/R-devel/library/base/html/assign.html

summary {base} Object Summaries  
https://stat.ethz.ch/R-manual/R-devel/library/base/html/summary.html

---

## Dplyr
https://m.youtube.com/watch?reload=9&feature=youtu.be&v=ztcsf7G4S78

dplyr: separate and unite  
https://youtu.be/ztcsf7G4S78

https://github.com/tidyverse/dtplyr

dplyr Tidy Animated Verbs  
https://www.garrickadenbuie.com/project/tidyexplain/

Chapter 5: plyr  
https://ademos.people.uic.edu/Chapter5.html

---

## Tidyverse Resources
Modern R with the tidyverse  
https://b-rodrigues.github.io/modern_R/

Tidylog provides feedback about basic dplyr operations  
*** https://github.com/elbersb/tidylog

Query R data frames with SQL  
https://github.com/ianmcook/tidyquery

The tidyverse style guide - Hadley Wickham  
*** https://style.tidyverse.org/

https://www.tidytuesday.com/6

https://www.tidytuesday.com/episodes

Text Mining with R   
https://www.tidytextmining.com/

https://www.tidytextmining.com/

Getting Started with the Tidyverse in R  
http://www.storybench.org/getting-started-with-tidyverse-in-r/

https://www.tidyverse.org/articles/2019/09/tidyr-1-0-0/

A moderndive into R and the tidyverse  
*** https://moderndive.com/index.html

https://moderndive.com/scripts/04-tidy.R

pivot_longer: Pivot data from wide to long  
https://rdrr.io/a/github/tidyverse/tidyr/man/pivot_longer.html

https://github.com/matloff/TidyverseSkeptic

https://www.r-bloggers.com/why-i-dont-use-the-tidyverse/

---

# Github Resources
*** https://happygitwithr.com/  
https://happygitwithr.com/rmd-test-drive.html

** https://github.com/mountainmath/cancensus

https://github.com/viveksrinivasanss/Data_Science_Using_R/tree/master/code

https://rawcdn.githack.com/EmilHvitfeldt/oRganized-talk/209977121efa5d3a97d4fb86090b9c6dfcd9f6b0/2018-10-29_getting-organized-with-r.html#1

https://rawcdn.githack.com/EmilHvitfeldt/laerug-Text-Mining-2018/29f1b3b2f96c0a2f23fbfea9620e404adc3c85af/slides.html#1

https://github.com/elbersb/tidylog

https://github.com/datacamp/datacamp-light

*** https://coolbutuseless.github.io/2019/10/09/the-miniverse-a-collection-of-document-creation-packages-for-rstats/

https://github.com/jolars/qualpalr

https://rladiesbelgrade.github.io/DSC5_RWS/DSRWay.html#1

https://lindeloev.github.io/mcp/

https://github.com/jennybc/send-email-with-r

Enter URL of the HTML file to preview:  
*** http://htmlpreview.github.io/

https://swcarpentry.github.io/r-novice-inflammation/06-best-practices-R

https://github.com/rstudio-education/arm-workshop-rsc2019

https://github.com/ropensci/geojsonio

https://github.com/ropensci/rtweet

https://github.com/ACCLAB/dabestr

https://m-clark.github.io/data-processing-and-visualization/

https://edwinth.github.io/ADSwR/

*** https://github.com/ropensci/internetarchive

https://jdblischak.github.io/workflowr/articles/wflow-02-customization.html

*** https://jdblischak.github.io/workflowr/

*** https://jdbcode.github.io/Snazzy-EE-TS-GIF/

https://jrnold.github.io/r4ds-exercise-solutions/

https://github.com/qinwf/awesome-R

*** sublime
https://github.com/randy3k/R-Box/

https://rstudio.github.io/r2d3/

https://gist.github.com/mollietaylor

https://github.com/AndyBunn

https://sejdemyr.github.io/#teaching

https://davisvaughan.github.io/almanac/articles/almanac.html

https://github.com/coryc3133/uoregon_r_bootcamp

https://github.com/gadenbuie/tidyexplain/tree/master/R

https://github.com/gadenbuie/tidyexplain

https://www.r-bloggers.com/tidy-evaluation-in-r-simple-examples/

https://github.com/thebioengineer/tidytuesdayR

*** https://github.com/sctyner/memer

*** https://jdblischak.github.io/workflowr/

Bhaskar V. Karambelkar  - leaflet
https://www.karambelkar.info/  
https://github.com/bhaskarvk  
https://github.com/bhaskarvk?tab=repositories

--- 

# DataCamp Resources
*** Convert any blog or website to an interactive learning platform for data science  
https://github.com/datacamp/datacamp-light

https://www.datacamp.com/community/tutorials/detect-anomalies-anomalize-r

https://www.datacamp.com/community/tutorials/merging-datasets-r

https://www.datacamp.com/community/tutorials/tutorial-on-loops-in-r

https://datacamp-community-prod.s3.amazonaws.com/e04c5a6b-4aca-46f5-8cd5-803d975ccc4b

Five Tips to Improve Your R Code - Datacamp  
https://www.datacamp.com/community/tutorials/five-tips-r-code-improve

[Markdown in Jupyter Notebook](https://www.datacamp.com/community/tutorials/markdown-in-jupyter-notebook)

https://www.datacamp.com/community/tutorials/tutorial-jupyter-notebook

Cluster Analysis (DataCamp)  
https://www.statmethods.net/advstats/cluster.html

https://www.datacamp.com/community/tutorials/r-formula-tutorial

https://www.datacamp.com/community/tutorials/r-web-scraping-rvest

https://www.datacamp.com/courses/introduction-to-time-series-analysis

https://www.datacamp.com/community/tutorials/time-series-r

## Quick R (DataCamp)  
Quick R Site map  
https://www.statmethods.net/about/sitemap.html  
Quick R About  
https://www.statmethods.net/index.html  
Customizing R Startup  
https://www.statmethods.net/interface/customizing.html  

https://www.datacamp.com/community/tutorials/visualize-data-vim-package

*** https://www.datacamp.com/community/tutorials/data-visualization-highcharter-r

---

# Update R keep packages
http://pierre.pudlo.perso.luminy.univ-amu.fr/wordpress/keeping-all-installed-packages-while-updating-r/

https://www.r-bloggers.com/upgrade-r-without-losing-your-packages/

https://community.rstudio.com/t/reinstalling-packages-on-new-version-of-r/7670/8

https://www.r-bloggers.com/updating-r-but-keeping-your-installed-packages/

https://stat.ethz.ch/R-manual/R-devel/library/utils/html/installed.packages.html

https://community.rstudio.com/t/issue-with-installed-packages/9458/8

---

# Other R links - unsorted

*** https://www.r-bloggers.com/how-to-learn-r-2/  
*** https://paulvanderlaken.com/2017/08/10/r-resources-cheatsheets-tutorials-books/

*** https://paulvanderlaken.com/2018/05/21/r-tips-and-tricks/

https://blog.r-hub.io/2019/12/03/readmes/

https://bookdown.org/

https://stackoverflow.com/questions/4459969/are-there-any-platforms-for-which-the-file-separator-isnt

http://vita.had.co.nz/papers/model-vis.pdf

*** https://r4ds.had.co.nz/

Forecasting: Principles and Practice  
https://otexts.com/fpp3/

Text Analysis  
http://rpubs.com/bnevt0/AT336106

https://joshuamrosenberg.com/post/2019/11/13/using-r-with-7th-grade-science-students/

https://rsuite.io/RSuite_Download.php

https://h-savran.blogspot.com/2017/08/learn-r-part-4-data-frames.html

http://www.r-fiddle.org/#/fiddle?id=SC5V50Pe&version=1

*** https://blog.rstudio.com/2019/11/06/renv-project-environments-for-r/

https://malco.io/2018/11/05/why-should-i-use-the-here-package-when-i-m-already-using-projects/

https://alan-y.netlify.com/post/objectremover-rstudio-addin/

*** https://towardsdatascience.com/exploring-highcharts-in-r-f754143efda7

https://github.com/gadenbuie/tidyexplain/blob/master/README.md

https://stackoverflow.com/questions/59116030/conditional-replacement-while-match-on-a-variable

*** knitr  
https://www.datafix.com.au/cookbook/  
*** https://github.com/yihui/knitr-examples/blob/master/009-slides.Rmd

https://stackoverflow.com/questions/59102998/passing-numeric-value-into-a-function-that-uses-dplyr

https://www.learnstatswithjamovi.com/

https://stackoverflow.com/questions/59033918/conditionally-aggregate-dataframe-rows-that-dont-contain-certain-values

https://mdsumner.github.io/spex/  
https://github.com/mdsumner/spex

https://www.r-bloggers.com/geocoding-with-tidygeocoder/

https://socialsciences.mcmaster.ca/jfox/Books/Companion/downloads.html  
https://socialsciences.mcmaster.ca/jfox/Books/Companion/index.html

https://www.inwt-statistics.com/read-blog/debugging-in-r.html

https://vickysteeves.gitlab.io/repro-papers/

https://github.com/laRusers/socalr2019

** http://faculty.marshall.usc.edu/gareth-james/ISL/  
http://faculty.marshall.usc.edu/gareth-james/ISL/All%20Labs.txt

*** Hands-On Programming with  
https://rstudio-education.github.io/hopr/

*** https://www.r-bloggers.com/5-ways-to-subset-a-data-frame-in-r/

*** https://livebook.manning.com/book/practical-data-science-with-r-second-edition/welcome/v-6/0

http://www.columbia.edu/~cjd11/charles_dimaggio/DIRE/resources/R/rFunctionsList.pdf

*** https://rich-iannone.github.io/blastula/articles/simple_composition.html  
https://randr.rocks/post/blastula-0-3/

https://www.eddjberry.com/post/writing-your-thesis-with-bookdown/  
*** https://bookdown.org/yihui/bookdown/

https://medium.com/@HadrienD/an-introduction-to-apis-for-r-users-223597a5ce0d

*** https://github.com/bcgov/bcgov-data-science-resources/wiki/Installing-R-&-RStudio

** https://joss.theoj.org/papers/10.21105/joss.01686

*** https://www.datasciencecentral.com/profiles/blogs/the-ultimate-r-cheatsheet

https://www.datasciencecentral.com/profiles/blogs/7-visualizations-you-should-learn-in-r

How to share data for collaboration
https://peerj.com/collections/50-practicaldatascistats/

https://peerj.com/preprints/3139/

https://rstudio.github.io/learnr/

https://github.com/rstudio-education/datascience-box

https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts

https://github.com/ropensci/plotly

https://rstudio.github.io/DT/
https://renkun-ken.github.io/formattable/

https://github.com/carlganz/rintrojs

Visdat when we started loading data!  
https://github.com/rfordatascience/tidytuesday/blob/master/README.md

https://wiki.qcbs.ca/geostatistics

https://blog.rstudio.com/2017/07/11/introducing-learnr/

http://adv-r.had.co.nz/Style.html

https://www.r-bloggers.com/5-ways-to-subset-a-data-frame-in-r/

https://www.r-bloggers.com/how-to-install-packages-on-r-screenshots/

*** https://rladiessydney.org/courses/ryouwithme/

*** https://cedricscherer.netlify.com/top/links/

https://drsimonj.svbtle.com/five-simple-tricks-to-improve-your-r-code

---

This web document were created with R Markdown, using RStudio v.1.1.463 and R v.3.6.1.

---