---
title: "HAT Monitoring Crew Data"
author: "Wendy Anthony for HAT"
date: "2026-03-08"
output:
  html_document:
    highlight: zenburn
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
    # highlight: textmate
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: '3'
# word_document:
  # mainfont: "Arial"      # Your main font (serif or sans-serif)
  # sansfont: "Helvetica"  # The specific sans-serif font for sans-serif contexts
  # fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries
```{r  include=FALSE, load-libraries}
# Load Libraries
#############################################
# uncomment install.packages("") to install a package, then replace comment #
#############################################

# Load Data Wrangling and Visualization Package Libraries
#############################################
# install.packages("tidyverse")
library(tidyverse) # includes ggplot2
# install.packages("plotly")
library(plotly) # interactive plots > ggplot2 requires plotly wrap for crosstalk
# install.packages("DT")
library(DT) # tables
```

# Clean and Wrangle Crew & Volunteer Data

## Data Import

### Crew Hours and Amount Removed Data
- To read and display Indigenous Languages save as CSV UTF-8 and use fileEncoding="UTF-8" to keep Aboriginal characters
```{r data-import-HAT}
# Data Import
################################# 
# To read and display Indigenous Languages save as CSV UTF-8
# and use fileEncoding="UTF-8" to keep Aboriginal characters
crew_data <- read.csv("data/HATRestorationCrewData-2022-2025-sepsp.csv", fileEncoding="UTF-8")

# Matson Field Logs
matson_fieldlog_2022_2025 <- read.csv("data/MatsonMattocksFieldLog-2016-2024/Matson-Field-Log-2022-2025.csv", fileEncoding="UTF-8")
```

# Data Exploration

## HAT Crew Data
```{r explore-data-crew-data}
glimpse(crew_data)

colnames(crew_data)
unique(crew_data$Work_Site)
```

<details>
<summary>Click to view structure of `crew_data`</summary>
```{r}
str(crew_data)
```
</details>

## Matson Volunteer Field log Data 2022-2025
<details>
<summary>Click to view structure of `matson_fieldlog_2022_2025`</summary>
```{r explore-data-volunteer-data}
str(matson_fieldlog_2022_2025)
# need to change date
```
</details>



# Data Wrangling

## Format HAT Crew Dates 
```{r data-wrangling-crew-data-dates}
################################# 
# Format HAT Crew Dates

# separate values
crew_data_sep <- crew_data %>% separate(Date, into = c("Day", "Month"), sep = "-")

# convert month name to number
crew_data_sep$Month <- match(crew_data_sep$Month, month.abb)

# create new column for Date
crew_data_sep$Date <- NA

crew_data_sep$Date <- paste(crew_data_sep$Year, crew_data_sep$Month,crew_data_sep$Day,sep =  "-")
str(crew_data_sep)

crew_data_sep$Date <- as.Date(crew_data_sep$Date)
str(crew_data_sep)
```



## Format Matson Fieldlog Dates
```{r data-wrangling-maston-data-dates}
################################# 
# Format Matson Fieldlog Dates

matson_fieldlog_2022_2025$Date <- as.Date(matson_fieldlog_2022_2025$Date)
str(matson_fieldlog_2022_2025)
```



## HAT Crew Columns - Change Names and Select
```{r rename-cols-crew-data}
colnames(crew_data_sep)
# Rename columns
# names(crew_data_sep)[names(crew_data_sep) == "Work.Site"] <- "Work_Site"
names(crew_data_sep)[names(crew_data_sep) == "Biomass.Removed..m3."] <- "Biomass_Removed_m3"
names(crew_data_sep)[names(crew_data_sep) == "Area.Cleared..m2."] <- "Area_Cleared_m2"
# names(crew_data_sep)[names(crew_data_sep) == "Invasive.Species.Cleared"] <- "Invasive_Species_Cleared"
# names(crew_data_sep)[names(crew_data_sep) == "Crew.Hours"] <- "Crew_Hours"

################################# 
# Columns - Subset Selected Columns
# Get column names
colnames(crew_data_sep)

# Remove unneeded columns
crew_data_sep_sub <- subset(crew_data_sep, select = -c(1,3,4,6,7,9,12,13,14,19))
colnames(crew_data_sep_sub)
str(crew_data_sep_sub)
```

## Matson Volunteer Columns - Change Names and Select
```{r rename-cols-vol-data}
# Matson field log remove unneeded columns
colnames(matson_fieldlog_2022_2025)
matson_fieldlog_2022_2025_sub <- subset(matson_fieldlog_2022_2025, select = -c(1,4,6,8,10,11,14,17:21))
```

## Change Format Numeric 
```{r change-numeric-format}
# change to number format
crew_data_sep_sub$Area_Cleared_m2 <- as.numeric(as.character(crew_data_sep_sub$Area_Cleared_m2))
crew_data_sep_sub$Biomass_Removed_m3 <- as.numeric(as.character(crew_data_sep_sub$Biomass_Removed_m3))
# Warning: NAs introduced by coercion May 6 2025 was a blank cell with no value
str(crew_data_sep_sub)

```

## Save Crew Data subset
```{r save-crew-subset}
crew_data_sep_sub
write.csv(crew_data_sep_sub, "data/crew_data_sep_sub.csv", row.names = FALSE)
```

## Filter by Site HAT Crew
```{r crew-filter-site}
colnames(crew_data_sep_sub)
unique(crew_data_sep_sub$Work_Site)

crew_data_sep_sub_matson <- crew_data_sep_sub[crew_data_sep_sub$Work_Site == "Matson Conservation Area",]
crew_data_sep_sub_OHP <- crew_data_sep_sub[crew_data_sep_sub$Work_Site == "Oak Haven Park",]
crew_data_sep_sub_yates <- crew_data_sep_sub[crew_data_sep_sub$Work_Site == "Yates",]
crew_data_sep_sub_funk <- crew_data_sep_sub[crew_data_sep_sub$Work_Site == "Funk",]
crew_data_sep_sub_ps <- crew_data_sep_sub[crew_data_sep_sub$Work_Site == "Purple Sanicle",]
crew_data_sep_sub_calypso <- crew_data_sep_sub[crew_data_sep_sub$Work_Site == "Calypso Woods",]
crew_data_sep_sub_camas <- crew_data_sep_sub[crew_data_sep_sub$Work_Site == "Camas Hill",]
crew_data_sep_sub_havenwood <- crew_data_sep_sub[crew_data_sep_sub$Work_Site == "Havenwood Park",]
crew_data_sep_sub_bear <- crew_data_sep_sub[crew_data_sep_sub$Work_Site == "Bear Hill",]
```


## Multiple Invasive Species Removed
- will need to make a new column - too many errors to fix, different capitalization, space at end of name
```{r}
length(unique(crew_data_sep_sub$Invasive_Species_Cleared)) #23
unique(crew_data_sep_sub$Invasive_Species_Cleared)
```

## Calculate Sums of Biomass and Hours
```{r}
colnames(crew_data_sep_sub)

crew_data_sep_sub_sum_bio_inv <- crew_data_sep_sub %>%
  group_by(Work_Site, Year, Invasive_Species_Cleared, Data_Change_Assumptions, Work_Volunteer) %>%
  summarise(Biomass_Removed_m3 = sum(Biomass_Removed_m3), 
            Crew_Hours = sum(Crew_Hours),
            Area_Cleared_m2 = sum(Area_Cleared_m2)
            )
crew_data_sep_sub_sum_bio_inv
# Check to ensure unique values of Invasive Species Cleared
unique(crew_data_sep_sub_sum_bio_inv$Invasive_Species_Cleared)
# Save Calculation subset
write.csv(crew_data_sep_sub_sum_bio_inv, "data/crew_data_sep_sub_sum_bio_inv.csv", row.names = FALSE)

```

## **** Calculations Group by Work Site and add work hours and biomass removed
```{r}
colnames(crew_data_sep_sub)
str(crew_data_sep_sub)

# aggregate(crew_data_sep_sub$Work_Site, by=list(Biomass_Removed_m3=crew_data_sep_sub$Biomass_Removed_m3), FUN=sum)
## Error in FUN(X[[i]], ...) : invalid 'type' (character) of argument

# crew_data_sep_sub_sum_by_group_site <- crew_data_sep_sub %>%
#   group_by(Work_Site, Year) %>%
#   summarise(Biomass_Removed_m3 = sum(Biomass_Removed_m3), 
#             Crew_Hours = sum(Crew_Hours),
#             Area_Cleared_m2 = sum(Area_Cleared_m2))
# crew_data_sep_sub_sum_by_group_site
# write.csv(crew_data_sep_sub_sum_by_group_site, "data/crew_data_sep_sub_sum_by_group_site.csv", row.names = FALSE)


crew_data_sep_sub_sum_by_group_site_hr_bio_ratio <- crew_data_sep_sub %>%
  group_by(Work_Site, Year, Work_Volunteer) %>%
  summarise(Biomass_Removed_m3 = sum(Biomass_Removed_m3), 
            Crew_Hours = sum(Crew_Hours),
            Area_Cleared_m2 = sum(Area_Cleared_m2),
            Labour_Hr_Bio_Ratio = round(sum(Crew_Hours)/sum(Biomass_Removed_m3), digits = 1),
            Labour_Bio_Hr_Ratio = round(sum(Biomass_Removed_m3)/sum(Crew_Hours), digits = 1), 
            Area_Bio_Ratio = round(sum(Area_Cleared_m2)/sum(Biomass_Removed_m3), digits = 1), 
            )
crew_data_sep_sub_sum_by_group_site_hr_bio_ratio
# Save Calculation subset
write.csv(crew_data_sep_sub_sum_by_group_site_hr_bio_ratio, "data/crew_data_sep_sub_sum_by_group_site_hr_bio_ratio.csv", row.names = FALSE)
```





# Data Summary Crew Data Import
```{r data-import-crew-bio, include = TRUE}
#############################################
#############################################
# Crew Data
crew_data_sep_sub_sum_by_group_site_1 <- read.csv("data/crew_data_sep_sub_sum_by_group_site_hr_bio_ratio.csv", fileEncoding="UTF-8")
str(crew_data_sep_sub_sum_by_group_site_1)
head(crew_data_sep_sub_sum_by_group_site_1)

crew_data_sep_sub_sum_bio_inv <- read.csv("data/crew_data_sep_sub_sum_bio_inv.csv", fileEncoding="UTF-8")
```

## Data Wrangling Summary Data - Rename site locations & Change column order
```{r crew-rename-site-location-worksite-columns-for-sd-group, include = TRUE}

#############################################
# Crew Data
# Summary data totals per year per site
crew_data_sep_sub_sum_by_group_site_rename_site <- crew_data_sep_sub_sum_by_group_site_1 %>%
  rename(Site = Work_Site)

crew_data_sep_sub_sum_by_group_site_rename_site$Site <- gsub("Matson Conservation Area", "Matson", crew_data_sep_sub_sum_by_group_site_rename_site$Site)
crew_data_sep_sub_sum_by_group_site_rename_site$Site <- gsub("Oak Haven Park", "Oak Haven", crew_data_sep_sub_sum_by_group_site_rename_site$Site)
crew_data_sep_sub_sum_by_group_site_rename_site$Site <- gsub("Havenwood Park", "Havenwood", crew_data_sep_sub_sum_by_group_site_rename_site$Site)

#############################################
write.csv(crew_data_sep_sub_sum_by_group_site_rename_site, "data/crew_data_sep_sub_sum_by_group_site_rename_site.csv")

crew_data_sep_sub_sum_bio_inv_rename_site <- crew_data_sep_sub_sum_bio_inv %>%
  rename(Site = Work_Site)

crew_data_sep_sub_sum_bio_inv_rename_site$Site <- gsub("Matson Conservation Area", "Matson", crew_data_sep_sub_sum_bio_inv_rename_site$Site)
crew_data_sep_sub_sum_bio_inv_rename_site$Site <- gsub("Oak Haven Park", "Oak Haven", crew_data_sep_sub_sum_bio_inv_rename_site$Site)

crew_data_sep_sub_sum_bio_inv_rename_site$Site <- gsub("Havenwood Park", "Havenwood", crew_data_sep_sub_sum_bio_inv_rename_site$Site)

# Top 5 Invasive Species
colnames(crew_data_sep_sub_sum_bio_inv_rename_site)
unique(crew_data_sep_sub_sum_bio_inv_rename_site$Invasive_Species_Cleared)

crew_data_sep_sub_sum_bio_inv_rename_site <- crew_data_sep_sub_sum_bio_inv_rename_site %>% relocate(Year, Site, Area_Cleared_m2, Biomass_Removed_m3, Invasive_Species_Cleared, Crew_Hours, Work_Volunteer, Data_Change_Assumptions)

crew_data_sep_sub_sum_bio_inv_top5 <- crew_data_sep_sub_sum_bio_inv_rename_site %>% filter(Invasive_Species_Cleared %in% c("daphne", "English ivy",  "Himalayan blackberry", "Scotch broom", "shiny geranium"))

colnames(crew_data_sep_sub_sum_bio_inv_rename_site)
```


# Import Crew Data
```{r data-import-crew, include = TRUE}
# Load Data
# Crew Data
################################# 
crew_data_sep_sub_1 <- read.csv("data/crew_data_sep_sub.csv", fileEncoding="UTF-8")

# Crew Data
crew_data_sep_sub_1 <- crew_data_sep_sub_1 %>% rename(Site = Work_Site)

crew_data_sep_sub_1$Site <- gsub("Matson Conservation Area", "Matson", crew_data_sep_sub_1$Site)
crew_data_sep_sub_1$Site <- gsub("Oak Haven Park", "Oak Haven", crew_data_sep_sub_1$Site)
crew_data_sep_sub_1$Site <- gsub("Havenwood Park", "Havenwood", crew_data_sep_sub_1$Site)
colnames(crew_data_sep_sub_1)

# Date format
str(crew_data_sep_sub_1)
crew_data_sep_sub_1$Date <- as.Date(crew_data_sep_sub_1$Date)

str(crew_data_sep_sub_1)

# Filter 3 rows remove 3 rows with NA
# Matson 2025-10-21
# Matson 2024-11-14
# Camas Hill 2025-09-04
crew_data_sep_sub_1_filtered <- crew_data_sep_sub_1 %>%
  filter(Date != as.Date("2025-10-21"),
         Date != as.Date("2024-11-14"),
         Date != as.Date("2025-09-04"))
```

## Columns move for filtered crew data
```{r crew-relocate-columns-for-sd-group, include = TRUE}
#############################################
#############################################
# change column order to same as volunteer_data_matson
colnames(crew_data_sep_sub_1_filtered)
#crew_data_sep_sub_1_filtered <- subset(crew_data_sep_sub_1_filtered, select = -c(3))
crew_data_sep_sub_1_filtered <- crew_data_sep_sub_1_filtered %>%
  relocate(Year, Date, Site, Area_Cleared_m2, Biomass_Removed_m3, Invasive_Species_Cleared, Native_or_Invasive, Crew_Hours, Work_Volunteer, Data_Change_Assumptions)
colnames(crew_data_sep_sub_1_filtered)
## go to combine with volunteer_data_matson_sub crew-vol<-rbind()
write.csv(crew_data_sep_sub_1_filtered, "data/crew_data_sep_sub_1_filtered.csv")
```


# Import Volunteer data
```{r data-import-volunteer, include = TRUE}
# Load Data
################################# 
## Matson Field log volunteer 
volunteer_data_matson <- read.csv("data/MatsonMattocksFieldLog-2016-2024/Matson-Field-Log-2022-2025.csv", fileEncoding="UTF-8")
volunteer_data_matson$Date <- as.Date(volunteer_data_matson$Date)
colnames(volunteer_data_matson)

#############################################
## subset Matson Field log volunteer 
volunteer_data_matson_sub <- subset(volunteer_data_matson, select = -c(1,4,6,8,11,14,17:21))
colnames(volunteer_data_matson_sub)

#############################################
# reorder columns to bind with crew data
volunteer_data_matson_sub <- volunteer_data_matson_sub %>% relocate(Year, Date, Site, Area_Cleared_m2, Biomass_Removed_m3, Invasive_Species_Cleared, Native_or_Invasive, Crew_Hours, Work_Volunteer, Data_Change_Assumptions)
colnames(volunteer_data_matson_sub)
## go to combine crew-vol<-rbind()
write.csv(volunteer_data_matson_sub, "data/volunteer_data_matson_sub.csv")
```

## Save Crew data  before combining
```{r save-crew-before-combine, include = TRUE}
## go to combine with volunteer_data_matson_sub crew-vol<-rbind()
write.csv(crew_data_sep_sub_1_filtered, "data/crew_data_sep_sub_1_filtered.csv")
```

## Save Volunteer data  before combining
```{r save-vol-before-combine, include = TRUE}
## go to combine crew-vol<-rbind()
write.csv(volunteer_data_matson_sub, "data/volunteer_data_matson_sub.csv")
```

## Combine Crew and Volunteer data
```{r combine-crew-vol, include = TRUE}
#############################################
#   COMBINING CREW AND VOLUNTEER DATA
# rbind work and volunteer
crew_vol <- rbind(crew_data_sep_sub_1_filtered, volunteer_data_matson_sub)
write.csv(crew_vol, "data/crew_vol.csv")

unique(crew_vol$Native_or_Invasive)
crew_vol_grs <- crew_vol[crew_vol$Native_or_Invasive == "MixedGrass",]
crew_vol_inv <- crew_vol[crew_vol$Native_or_Invasive == "Invasive",]

write.csv(crew_vol_grs, "data/crew_vol_grs.csv")
write.csv(crew_vol_inv, "data/crew_vol_inv.csv")
```

## Top 5 Invasive Species from Crew Data `crew_vol_inv_top5`
```{r crew-vol-inv-top5, include = TRUE}
#############################################
crew_vol_inv_top5 <- crew_vol %>% filter(Invasive_Species_Cleared %in% c("daphne", "English ivy",  "Himalayan blackberry", "Scotch broom", "shiny geranium"))

colnames(crew_vol_inv_top5)
write.csv(crew_vol_inv_top5, "data/crew_vol_inv_top5.csv")

#############################################
  crew_vol_inv_top5_sum <- crew_vol_inv_top5 %>%
  group_by(Site, Year, Invasive_Species_Cleared) %>%
  summarise(Biomass_Removed_m3 = sum(Biomass_Removed_m3), 
            Crew_Hours = sum(Crew_Hours),
            Area_Cleared_m2 = sum(Area_Cleared_m2)
            )
crew_vol_inv_top5_sum
# Check to ensure unique values of Invasive Species Cleared

#############################################
# Save Calculation subset
write.csv(crew_vol_inv_top5_sum, "data/crew_vol_inv_top5_sum.csv", row.names = FALSE)
```


### `crew_data_sep_sub_sum_bio_inv_top5`
```{r crew-bio-inv-filter-top5, include = TRUE}
#############################################
# Top Invasive Species
crew_data_sep_sub_sum_bio_inv_top5 <- crew_data_sep_sub_sum_bio_inv_rename_site %>% filter(Invasive_Species_Cleared %in% c("daphne", "English ivy",  "Himalayan blackberry", "Scotch broom", "shiny geranium"))
```

