---
title: "HAT Monitoring Cover Data"
author: "Wendy Anthony for HAT"
date: "2026-03-08"
output:
  html_document:
    highlight: zenburn
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
    # highlight: textmate
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: '3'
# word_document:
  # mainfont: "Arial"      # Your main font (serif or sans-serif)
  # sansfont: "Helvetica"  # The specific sans-serif font for sans-serif contexts
  # fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries
```{r load-libraries}
# install.packages("tidyverse")
library(tidyverse)
```

## Import Cover Data 
- Oak Haven 2024, 2025  
- Matson 2024, 2025
```{r import-cover-data}
## Cover data from original Meadow Monitoring Spreadsheet 2025.xlsx & Meadow Monitoring Spreadsheet 2024.xlsx
# separated by year and site
# Oak Haven Cover data 2025
## change invasive grass to mixed grass
cover_OH_2024 <- read.csv("data/HAT-MeadowMonitoring-2024-OakHaven.csv", fileEncoding="UTF-8")
cover_OH_2025 <- read.csv("data/HAT-MeadowMonitoring-2025-OakHaven.csv", fileEncoding="UTF-8")
cover_MCA_2024 <- read.csv("data/HAT-MeadowMonitoring-2024-Matson.csv", fileEncoding="UTF-8")
cover_MCA_2025 <- read.csv("data/HAT-MeadowMonitoring-2025-Matson.csv", fileEncoding="UTF-8")
```



## HAT Cover Data
### rbind 4 file ojbects into 1 in long format
```{r rbind-cover-data}
HAT_cover_OH_MM <- rbind(cover_OH_2024, cover_OH_2025, cover_MCA_2024, cover_MCA_2025)


unique(HAT_cover_OH_MM$Species) #47

# trim trailing while spaces
HAT_cover_OH_MM$Species <- trimws(HAT_cover_OH_MM$Species)
species_unique <- unique(HAT_cover_OH_MM$Species) #39

 
# create a list of comma separated values for list of species names, but in no order
 comma_separated_sp_names <- paste(species_unique, collapse = ", ")
 comma_separated_sp_names
 str(comma_separated_sp_names)
 # these did not sort in any order
#  comma_separated_sp_names_sort <- as_tibble(comma_separated_sp_names) %>%
#  arrange(comma_separated_sp_names)
# comma_separated_sp_names_sort

# update sort ID column
HAT_cover_OH_MM$SortID <- 1:nrow(HAT_cover_OH_MM)


# as.Date
str(HAT_cover_OH_MM)
HAT_cover_OH_MM$Date <- as.Date(HAT_cover_OH_MM$Date)
str(HAT_cover_OH_MM)

# Concatenate QUniqueID
colnames(HAT_cover_OH_MM)
HAT_cover_OH_MM$QUniqueID <- paste(HAT_cover_OH_MM$MonitorID, HAT_cover_OH_MM$Quadrat, HAT_cover_OH_MM$QuadratLocation_m, sep = "_")

# Replace hyphen with underscores in QUniqueID
HAT_cover_OH_MM$QUniqueID <- gsub("-", "_", HAT_cover_OH_MM$QUniqueID)

# remove excess coloumns not needed
# including SortID column
colnames(HAT_cover_OH_MM)
HAT_cover_OH_MM_sub <- subset(HAT_cover_OH_MM, select = -c(1,2,9,10,11,12,16))

# write csv
write.csv(HAT_cover_OH_MM_sub, "data/HAT_cover_OH_MM_sub.csv", row.names = FALSE)
```
<details>
<summary>Click to view `HAT_cover_OH_MM_sub`</summary>
```{r}
HAT_cover_OH_MM_sub
```
</details>



### Filter Subset
#### Native vs Invasive vs MixedGrass
```{r filter-inv-nat-subset}
unique(HAT_cover_OH_MM_sub$Native_or_Invasive)
HAT_cover_OH_MM_sub_nat <- HAT_cover_OH_MM_sub[HAT_cover_OH_MM_sub$Native_or_Invasive == "Native",]

HAT_cover_OH_MM_sub_inv <- HAT_cover_OH_MM_sub[HAT_cover_OH_MM_sub$Native_or_Invasive == "Invasive",]

HAT_cover_OH_MM_sub_mix <- HAT_cover_OH_MM_sub[HAT_cover_OH_MM_sub$Native_or_Invasive == "MixedGrass",]
```

#### Treated vs Untreated
```{r filter-treat-untreat-subset}
unique(HAT_cover_OH_MM_sub$Treated)
HAT_cover_OH_MM_sub_treat <- HAT_cover_OH_MM_sub[HAT_cover_OH_MM_sub$Treated == "Treated",]
HAT_cover_OH_MM_sub_untreat <- HAT_cover_OH_MM_sub[HAT_cover_OH_MM_sub$Treated == "Untreated",]
```

### *** Convert long format to wide format and group by Species
WORKS !!!!
See function that creates this wide format
```{r cover-data-long-to-wide}
colnames(HAT_cover_OH_MM_sub)
unique(HAT_cover_OH_MM_sub$Location)
# filter subset
HAT_cover_OH_MM_sub_treat_OH <- HAT_cover_OH_MM_sub_treat[HAT_cover_OH_MM_sub_treat$Location == "Oak Haven",]
HAT_cover_OH_MM_sub_treat_OH_2025 <- HAT_cover_OH_MM_sub_treat_OH[HAT_cover_OH_MM_sub_treat_OH$Year == "2025",]
HAT_cover_OH_MM_sub_treat_OH_2024 <- HAT_cover_OH_MM_sub_treat_OH[HAT_cover_OH_MM_sub_treat_OH$Year == "2024",]
HAT_cover_OH_MM_sub_untreat_OH <- HAT_cover_OH_MM_sub_untreat[HAT_cover_OH_MM_sub_untreat$Location == "Oak Haven",]
HAT_cover_OH_MM_sub_untreat_OH_2025 <- HAT_cover_OH_MM_sub_untreat_OH[HAT_cover_OH_MM_sub_untreat_OH$Year == "2025",]
HAT_cover_OH_MM_sub_untreat_OH_2024 <- HAT_cover_OH_MM_sub_untreat_OH[HAT_cover_OH_MM_sub_untreat_OH$Year == "2024",]

HAT_cover_OH_MM_sub_treat_M <- HAT_cover_OH_MM_sub_treat[HAT_cover_OH_MM_sub_treat$Location == "Matson",]
HAT_cover_OH_MM_sub_treat_M_2025 <- HAT_cover_OH_MM_sub_treat_M[HAT_cover_OH_MM_sub_treat_M$Year == "2025",]
HAT_cover_OH_MM_sub_treat_M_2024 <- HAT_cover_OH_MM_sub_treat_M[HAT_cover_OH_MM_sub_treat_M$Year == "2024",]
HAT_cover_OH_MM_sub_untreat_M <- HAT_cover_OH_MM_sub_untreat[HAT_cover_OH_MM_sub_untreat$Location == "Matson",]
HAT_cover_OH_MM_sub_untreat_M_2025 <- HAT_cover_OH_MM_sub_untreat_M[HAT_cover_OH_MM_sub_untreat_M$Year == "2025",]
HAT_cover_OH_MM_sub_untreat_M_2024 <- HAT_cover_OH_MM_sub_untreat_M[HAT_cover_OH_MM_sub_untreat_M$Year == "2024",]


colnames(HAT_cover_OH_MM_sub_treat_OH_2025)

# long to wide
HAT_cover_OH_MM_sub_treat_OH_2025_wide <- HAT_cover_OH_MM_sub_treat_OH_2025 %>% 
pivot_wider(names_from = QUniqueID, values_from = PercentCover, values_fill = 0)
HAT_cover_OH_MM_sub_treat_OH_2024_wide <- HAT_cover_OH_MM_sub_treat_OH_2024 %>% 
pivot_wider(names_from = QUniqueID, values_from = PercentCover, values_fill = 0)

# HAT_cover_OH_MM_sub_untreat_OH_2025_wide <- HAT_cover_OH_MM_sub_untreat_OH_2025 %>%
# pivot_wider(names_from = QUniqueID, values_from = PercentCover, values_fill = 0)
# HAT_cover_OH_MM_sub_untreat_OH_2024_wide <- HAT_cover_OH_MM_sub_untreat_OH_2024 %>%
# pivot_wider(names_from = QUniqueID, values_from = PercentCover, values_fill = 0)
# 
# HAT_cover_OH_MM_sub_treat_M_2025_wide <- HAT_cover_OH_MM_sub_treat_M_2025 %>%
# pivot_wider(names_from = QUniqueID, values_from = PercentCover, values_fill = 0)
# HAT_cover_OH_MM_sub_treat_M_2024_wide <- HAT_cover_OH_MM_sub_treat_M_2024 %>%
# pivot_wider(names_from = QUniqueID, values_from = PercentCover, values_fill = 0)
# 
# HAT_cover_OH_MM_sub_untreat_M_2025_wide <- HAT_cover_OH_MM_sub_untreat_M_2025 %>%
# pivot_wider(names_from = QUniqueID, values_from = PercentCover, values_fill = 0)
# HAT_cover_OH_MM_sub_untreat_M_2024_wide <- HAT_cover_OH_MM_sub_untreat_M_2024 %>%
# pivot_wider(names_from = QUniqueID, values_from = PercentCover, values_fill = 0)


### SOME PROBLEM WITH DUPLICATE DATA HAT_cover_OH_MM_sub_untreat_OH_2024
# HAT_cover_OH_MM_sub_untreat_OH_2024_wide <- HAT_cover_OH_MM_sub_untreat_OH_2024 %>% 
# pivot_wider(names_from = QUniqueID, values_from = PercentCover, values_fill = 0)

# colnames(HAT_cover_OH_MM_sub_untreat_OH_2024_wide)

# sum all percentage cover per species over all Quadrats
HAT_cover_OH_MM_sub_treat_OH_2025_wide_agg <- HAT_cover_OH_MM_sub_treat_OH_2025_wide %>% 
group_by(Year, Date, Location, Treated, Species, Native_or_Invasive) %>% 
  summarise(across(c(T_DL_Q1_4.3, T_DL_Q2_3.3, T_DL_Q3_8.3, T_CH_Q3_8.3, T_CH_Q1_4.3, T_CH_Q2_3.3), sum))
 
colnames(HAT_cover_OH_MM_sub_treat_OH_2025_wide_agg)

```



### *** Calculate Mean Cover Percent
WORKS !!!!
```{r cover-mean}
### *** Calculate Mean Cover Percent
HAT_cover_OH_MM_sub_treat_OH_2025_wide_agg_mean <- HAT_cover_OH_MM_sub_treat_OH_2025_wide_agg %>%
   group_by(Year, Date, Location, Treated, Species) %>%
  summarise(mean_value = mean(c(T_DL_Q1_4.3, T_DL_Q2_3.3, T_DL_Q3_8.3, T_CH_Q3_8.3, T_CH_Q1_4.3, T_CH_Q2_3.3)), 
            sd_value = sd(c(T_DL_Q1_4.3, T_DL_Q2_3.3, T_DL_Q3_8.3, T_CH_Q3_8.3, T_CH_Q1_4.3, T_CH_Q2_3.3))) %>%  
  # round
  mutate_if(is.numeric, round, 2)

 HAT_cover_OH_MM_sub_treat_OH_2025_wide_agg_mean
 
colnames(HAT_cover_OH_MM_sub_treat_OH_2025_wide_agg_mean)

write.csv(HAT_cover_OH_MM_sub_treat_OH_2025_wide_agg_mean, "data/HAT_cover_OH_MM_sub_treat_OH_2025_wide_agg_mean.csv", row.names = FALSE)

```


## Function Cover Subset
- in preparation for creating average mean percentage cover
```{r fun-cover-sub}

unique(HAT_cover_OH_MM_sub$Location)


HAT_cover_subset_fun <- function(data, col_name, location, col_name_2, treat, col_name_3, year){
   
  data  <- data %>% dplyr::filter(
    {{col_name}} == location,
    {{col_name_2}} == treat,
    {{col_name_3}} == year)
  
  return(data)
  data_location_treat_year
  
}
HAT_cover_subset_fun_OH_T_2025 <- HAT_cover_subset_fun(HAT_cover_OH_MM_sub, Location, "Oak Haven", Treated, "Treated", Year, "2025")
HAT_cover_subset_fun_OH_U_2025 <- HAT_cover_subset_fun(HAT_cover_OH_MM_sub, Location, "Oak Haven", Treated, "Untreated", Year, "2025")
HAT_cover_subset_fun_OH_T_2024 <- HAT_cover_subset_fun(HAT_cover_OH_MM_sub, Location, "Oak Haven", Treated, "Treated", Year, "2024")
HAT_cover_subset_fun_OH_U_2024 <- HAT_cover_subset_fun(HAT_cover_OH_MM_sub, Location, "Oak Haven", Treated, "Untreated", Year, "2024")
HAT_cover_subset_fun_M_T_2025 <- HAT_cover_subset_fun(HAT_cover_OH_MM_sub, Location, "Matson", Treated, "Treated", Year, "2025")
HAT_cover_subset_fun_M_U_2025 <- HAT_cover_subset_fun(HAT_cover_OH_MM_sub, Location, "Matson", Treated, "Untreated", Year, "2025")
HAT_cover_subset_fun_M_T_2024 <- HAT_cover_subset_fun(HAT_cover_OH_MM_sub, Location, "Matson", Treated, "Treated", Year, "2024")
HAT_cover_subset_fun_M_U_2024 <- HAT_cover_subset_fun(HAT_cover_OH_MM_sub, Location, "Matson", Treated, "Untreated", Year, "2024")
```

## Function Cover Subset pivot wide
- in preparation for creating average mean percentage cover
```{r fun-cover-sub-pivot-wide, echo=F, results='hide'}

#unique(HAT_cover_OH_MM_sub$Location)


HAT_cover_subset_fun_pivot_wide <- function(data, col_name, location, col_name_2, treat, col_name_3, year){
   
  data  <- data %>% dplyr::filter(
    {{col_name}} == location,
    {{col_name_2}} == treat,
    {{col_name_3}} == year) %>% 
    pivot_wider(names_from = QUniqueID, values_from = PercentCover, values_fill = 0)  
}
#    pivot_wider(names_from = QUniqueID, values_from = PercentCover, values_fill = 0)  
## Error in `pivot_wider()`:
# ! Can't convert `fill` <double> to <list>.
# # HAT_cover_subset_fun_pivot_wide()
```


### Populate pivot wide using function Cover Subset pivot wide
```{r fun-subset-pivot-wide}
HAT_cover_subset_fun_OH_T_2025_wide <- HAT_cover_subset_fun_pivot_wide(HAT_cover_OH_MM_sub, Location, "Oak Haven", Treated, "Treated", Year, "2025")
HAT_cover_subset_fun_OH_U_2025_wide <- HAT_cover_subset_fun_pivot_wide(HAT_cover_OH_MM_sub, Location, "Oak Haven", Treated, "Untreated", Year, "2025")
HAT_cover_subset_fun_OH_T_2024_wide <- HAT_cover_subset_fun_pivot_wide(HAT_cover_OH_MM_sub, Location, "Oak Haven", Treated, "Treated", Year, "2024")

## unable to pivot wide and fill to 0 ## ????????? had to wrangle some duplicate values and quadrat ID
HAT_cover_subset_fun_OH_U_2024_wide <- HAT_cover_subset_fun_pivot_wide(HAT_cover_OH_MM_sub, Location, "Oak Haven", Treated, "Untreated", Year, "2024")

HAT_cover_subset_fun_M_T_2025_wide <- HAT_cover_subset_fun_pivot_wide(HAT_cover_OH_MM_sub, Location, "Matson", Treated, "Treated", Year, "2025")
HAT_cover_subset_fun_M_U_2025_wide <- HAT_cover_subset_fun_pivot_wide(HAT_cover_OH_MM_sub, Location, "Matson", Treated, "Untreated", Year, "2025")
HAT_cover_subset_fun_M_T_2024_wide <- HAT_cover_subset_fun_pivot_wide(HAT_cover_OH_MM_sub, Location, "Matson", Treated, "Treated", Year, "2024")


HAT_cover_subset_fun_M_U_2024_wide <- HAT_cover_subset_fun_pivot_wide(HAT_cover_OH_MM_sub, Location, "Matson", Treated, "Untreated", Year, "2024")
```




### Calculate Mean Cover Percentage from pivot wide function results
```{r fun-col-names}
# gather column names for each data object to paste the UniqueID values into mean calculations 
comma_separated_col_names <- function(data){
  col  <- data

cols <- colnames(col)

comma_separated_col_names <- paste(cols, collapse = ", ")

print(comma_separated_col_names)

}

##### COPY COLUMN SEPARATED NAMES FOR UNIQUE QUADRATS OT REPLACE C() VALUES
# HAT_cover_subset_fun_OH_T_2025_wide
comma_separated_col_names(HAT_cover_subset_fun_OH_T_2025_wide)
HAT_cover_subset_fun_OH_T_2025_wide_mean <- HAT_cover_subset_fun_OH_T_2025_wide %>%
   group_by(Year, Date, Location, Treated, Species, Native_or_Invasive) %>%
  summarise(mean_value = mean(c(T_DL_Q1_4.3, T_DL_Q2_3.3, T_DL_Q3_8.3, T_CH_Q1_4.3, T_CH_Q2_3.3, T_CH_Q3_8.3)), 
            sd_value = sd(c(T_DL_Q1_4.3, T_DL_Q2_3.3, T_DL_Q3_8.3, T_CH_Q1_4.3, T_CH_Q2_3.3, T_CH_Q3_8.3))) %>%  
  mutate_if(is.numeric, round, 2) # round

# HAT_cover_subset_fun_OH_U_2025_wide
comma_separated_col_names(HAT_cover_subset_fun_OH_U_2025_wide)
HAT_cover_subset_fun_OH_U_2025_wide_mean <- HAT_cover_subset_fun_OH_U_2025_wide %>%
   group_by(Year, Date, Location, Treated, Species, Native_or_Invasive) %>%
  summarise(mean_value = mean(c(U_DL_Q1_9, U_DL_Q2_13.3, U_DL_Q3_2.6, U_CH_Q1_9, U_CH_Q2_13.3, U_CH_Q3_2.6)), 
            sd_value = sd(c(U_DL_Q1_9, U_DL_Q2_13.3, U_DL_Q3_2.6, U_CH_Q1_9, U_CH_Q2_13.3, U_CH_Q3_2.6))) %>%  
  mutate_if(is.numeric, round, 2) # round

# HAT_cover_subset_fun_OH_T_2024_wide
comma_separated_col_names(HAT_cover_subset_fun_OH_T_2024_wide)
HAT_cover_subset_fun_OH_T_2024_wide_mean <- HAT_cover_subset_fun_OH_T_2024_wide %>%
   group_by(Year, Date, Location, Treated, Species, Native_or_Invasive) %>%
  summarise(mean_value = mean(c(T_CL_Q2_3.3, T_CL_Q1_4.3, T_HD_Q1_4.3, T_HD_Q3_8.3, T_HD_Q2_3.3, T_CL_Q3_8.3)), 
            sd_value = sd(c(T_CL_Q2_3.3, T_CL_Q1_4.3, T_HD_Q1_4.3, T_HD_Q3_8.3, T_HD_Q2_3.3, T_CL_Q3_8.3))) %>%  
  mutate_if(is.numeric, round, 2) # round

# HAT_cover_subset_fun_OH_U_2024_wide
comma_separated_col_names(HAT_cover_subset_fun_OH_U_2024_wide)
HAT_cover_subset_fun_OH_U_2024_wide_mean <- HAT_cover_subset_fun_OH_U_2024_wide %>%
   group_by(Year, Date, Location, Treated, Species, Native_or_Invasive) %>%
  summarise(mean_value = mean(c(U_CL_Q1_9, U_CL_Q3_2.6, U_CL2_Q3_2.6, U_CL2_Q1_9, U_CL2_Q2_13.3, U_CL_Q2_13.3)), 
            sd_value = sd(c(U_CL_Q1_9, U_CL_Q3_2.6, U_CL2_Q3_2.6, U_CL2_Q1_9, U_CL2_Q2_13.3, U_CL_Q2_13.3))) %>%  
  mutate_if(is.numeric, round, 2) # round

# HAT_cover_subset_fun_M_T_2025_wide
comma_separated_col_names(HAT_cover_subset_fun_M_T_2025_wide)
HAT_cover_subset_fun_M_T_2025_wide_mean <- HAT_cover_subset_fun_M_T_2025_wide %>%
   group_by(Year, Date, Location, Treated, Species, Native_or_Invasive) %>%
  summarise(mean_value = mean(c(T_LJ_Q1_8, T_LJ_Q3_4, T_DH_Q3_8, T_DH_Q1_4, T_LJ_Q2_1, T_DH_Q2_1)), 
            sd_value = sd(c(T_LJ_Q1_8, T_LJ_Q3_4, T_DH_Q3_8, T_DH_Q1_4, T_LJ_Q2_1, T_DH_Q2_1))) %>%  
  mutate_if(is.numeric, round, 2) # round

# HAT_cover_subset_fun_M_U_2025_wide
comma_separated_col_names(HAT_cover_subset_fun_M_U_2025_wide)
HAT_cover_subset_fun_M_U_2025_wide_mean <- HAT_cover_subset_fun_M_U_2025_wide %>%
   group_by(Year, Date, Location, Treated, Species, Native_or_Invasive) %>%
  summarise(mean_value = mean(c(U_LJ_Q3_9, U_LJ_Q2_2, U_DH_Q1_2, U_DH_Q2_4, U_DH_Q3_9, U_LJ_Q1_4)), 
            sd_value = sd(c(U_LJ_Q3_9, U_LJ_Q2_2, U_DH_Q1_2, U_DH_Q2_4, U_DH_Q3_9, U_LJ_Q1_4))) %>%  
  mutate_if(is.numeric, round, 2) # round

# HAT_cover_subset_fun_M_T_2024_wide
comma_separated_col_names(HAT_cover_subset_fun_M_T_2024_wide)
HAT_cover_subset_fun_M_T_2024_wide_mean <- HAT_cover_subset_fun_M_T_2024_wide %>%
   group_by(Year, Date, Location, Treated, Species, Native_or_Invasive) %>%
  summarise(mean_value = mean(c(T_HCDD_Q3_8, T_HCDD_Q1_2, T_HCDD_Q2_5)), 
            sd_value = sd(c(T_HCDD_Q3_8, T_HCDD_Q1_2, T_HCDD_Q2_5))) %>%  
  mutate_if(is.numeric, round, 2) # round

# HAT_cover_subset_fun_OH_U_2025_wide
comma_separated_col_names(HAT_cover_subset_fun_M_U_2024_wide)
HAT_cover_subset_fun_M_U_2024_wide_mean <- HAT_cover_subset_fun_M_U_2024_wide %>%
   group_by(Year, Date, Location, Treated, Species, Native_or_Invasive) %>%
  summarise(mean_value = mean(c(U_HCDD_Q1_3.66, U_HCDD_Q2_5.7, U_HCDD_Q3_7.3)), 
            sd_value = sd(c(U_HCDD_Q1_3.66, U_HCDD_Q2_5.7, U_HCDD_Q3_7.3))) %>%  
  mutate_if(is.numeric, round, 2) # round

```


### Rbind mean cover data
```{r mean-cover-merge}
HAT_cover_mean <- rbind(HAT_cover_subset_fun_OH_T_2025_wide_mean, HAT_cover_subset_fun_OH_U_2025_wide_mean, HAT_cover_subset_fun_OH_T_2024_wide_mean, HAT_cover_subset_fun_OH_U_2024_wide_mean, HAT_cover_subset_fun_M_T_2025_wide_mean, HAT_cover_subset_fun_M_U_2025_wide_mean, HAT_cover_subset_fun_M_T_2024_wide_mean, HAT_cover_subset_fun_M_U_2024_wide_mean)

write.csv(HAT_cover_mean, "data/HAT_cover_mean.csv", row.names = FALSE)
```


## Report: Species Richness Values
```{r mean-cover-merge-species-richness, results = TRUE, echo = FALSE}
# colnames(HAT_cover_mean)
# unique(HAT_cover_mean$Native_or_Invasive)

## Native vs Invasive
nat_cover_mean <- subset(HAT_cover_mean, Native_or_Invasive == "Native")
inv_cover_mean <- subset(HAT_cover_mean, Native_or_Invasive == "Invasive")  

nat_cover_mean_OH <- subset(nat_cover_mean, Location == "Oak Haven")
inv_cover_mean_OH <- subset(inv_cover_mean, Location == "Oak Haven")
nat_cover_mean_M <- subset(nat_cover_mean, Location == "Matson")
inv_cover_mean_M <- subset(inv_cover_mean, Location == "Matson")

# Treated vs Untreated
nat_cover_T_mean <- subset(nat_cover_mean, Treated == "Treated")
nat_cover_U_mean <- subset(nat_cover_mean, Treated == "Untreated")  
inv_cover_T_mean <- subset(inv_cover_mean, Treated == "Treated")
inv_cover_U_mean <- subset(inv_cover_mean, Treated == "Untreated")  


nat_cover_T_mean_OH <- subset(nat_cover_T_mean, Location == "Oak Haven")
nat_cover_U_mean_OH <- subset(nat_cover_U_mean, Location == "Oak Haven")
inv_cover_T_mean_OH <- subset(inv_cover_T_mean, Location == "Oak Haven")
inv_cover_U_mean_OH <- subset(inv_cover_U_mean, Location == "Oak Haven")

nat_cover_T_mean_M <- subset(nat_cover_T_mean, Location == "Matson")
nat_cover_U_mean_M <- subset(nat_cover_U_mean, Location == "Matson")
inv_cover_T_mean_M <- subset(inv_cover_T_mean, Location == "Matson")
inv_cover_U_mean_M <- subset(inv_cover_U_mean, Location == "Matson")
```

### In Oak Haven
- There are `r length(unique(nat_cover_mean_OH$Species))` unique native plant species, and `r   length(unique(inv_cover_mean_OH$Species))` invasive Plant species   
- `r length(unique(nat_cover_T_mean_OH$Species))` unique native plant species in the Treated, and `r length(unique(nat_cover_U_mean_OH$Species))` unique native plant species in the Untreated monitoring locations
- `r length(unique(inv_cover_T_mean_OH$Species))` unique invasive plant species in the Treated, and `r length(unique(inv_cover_U_mean_OH$Species))` unique invasive plant species in the Untreated monitoring locations
- The **native plant species** are: `r unique(nat_cover_mean_OH$Species)`  
- The **invasive plant species** are: `r unique(inv_cover_mean_OH$Species)`  

### In Matson
- There are `r length(unique(nat_cover_mean_M$Species))` unique native plant species, and `r   length(unique(inv_cover_mean_M$Species))` invasive Plant species   
- `r length(unique(nat_cover_T_mean_M$Species))` unique native plant species in the Treated, and `r length(unique(nat_cover_U_mean_M$Species))` unique native plant species in the Untreated monitoring locations
- `r length(unique(inv_cover_T_mean_M$Species))` unique invasive plant species in the Treated, and `r length(unique(inv_cover_U_mean_M$Species))` unique invasive plant species in the Untreated monitoring locations
- The **native plant species** are: `r unique(nat_cover_mean_M$Species)`  
- The **invasive plant species** are: `r unique(inv_cover_mean_M$Species)`  

```{r add-values-to-report, results = FALSE, echo = TRUE}
"- There are `r length(unique(HAT_cover_mean$Species))` unique native plant species, 
and `r length(unique(HAT_cover_mean$Species))` invasive Plant species 
found in both the Treated and Untreated monitoring locations at Oak Haven Park"
"- The **native plant species** are: `r unique(HAT_cover_mean$Species)`  "
"- The **invasive plant species** are: `r unique(HAT_cover_mean$Species)`  "
```


---

# Prepare Cover Data to be imported into Webfilter

## Import Cover Data
```{r data-import-cover}
# Load Data
################################# 
# Cover data
cover_data <- read.csv("data/HAT_cover_mean.csv", fileEncoding="UTF-8")
```

### Explore Cover Data
```{r explore-cover-data}
str(cover_data)
```

## Prepare Cover Data
```{r cover-rename-site-location-worksite-columns-for-sd-group, include = FALSE}
#############################################
# Cover Data
cover_data_rename_site <- cover_data %>%
  rename(Site = Location)

cover_data_rename_site$Site <- gsub("Matson Conservation Area", "Matson", cover_data_rename_site$Site)
cover_data_rename_site$Site <- gsub("Oak Haven Park", "Oak Haven", cover_data_rename_site$Site)
cover_data_rename_site$Site <- gsub("Havenwood Park", "Havenwood", cover_data_rename_site$Site)

colnames(cover_data_rename_site)
cover_data_rename_site_sub <- subset(cover_data_rename_site, select = -c(9,10,11))

cover_data_rename_site_sub$Date <- as.Date(cover_data_rename_site_sub$Date)

## Add upper and lower bounds mean +/- sd

cover_data_rename_site_sub <- cover_data_rename_site_sub %>% 
                                      mutate(lower = mean_value - sd_value)

cover_data_rename_site_sub <- cover_data_rename_site_sub %>% 
                                      mutate(upper = mean_value + sd_value)

str(cover_data_rename_site_sub)

## ROund all numbers to 2 decimals
cover_data_rename_site_sub <- cover_data_rename_site_sub %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))
write.csv(cover_data_rename_site_sub, "data/cover_data_rename_site_sub.csv")

#############################################
cov_nat_top5 <- cover_data_rename_site_sub %>% filter(Species %in% c("camas", "delphinium",  "Pacific sanicle", "licorice fern", "Oregon grape"))
write.csv(cov_nat_top5, "data/cov_nat_top5.csv")
#############################################
unique(cover_data_rename_site_sub$Native_or_Invasive)
cover_data_rename_site_sub_grs <- cover_data_rename_site_sub[cover_data_rename_site_sub$Native_or_Invasive == "MixedGrass",]
cover_data_rename_site_sub_inv <- cover_data_rename_site_sub[cover_data_rename_site_sub$Native_or_Invasive == "Invasive",]
cover_data_rename_site_sub_nat <- cover_data_rename_site_sub[cover_data_rename_site_sub$Native_or_Invasive == "Native",]

write.csv(cover_data_rename_site_sub_grs, "data/cover_data_rename_site_sub_grs.csv")
write.csv(cover_data_rename_site_sub_inv, "data/cover_data_rename_site_sub_inv.csv")
write.csv(cover_data_rename_site_sub_nat, "data/cover_data_rename_site_sub_nat.csv")
```




## Import Species Count
```{r data-import-species-count}
# Load Data
################################# 
# Species counts
species_count <- read.csv("data/species_counts.csv", fileEncoding = "UTF-8")

#############################################
# drop unknown species
species_count <- species_count[-c(25), ]
species_count <- species_count %>%
  rename(Site = Location)

write.csv(species_count, "data/species_count_wrangled.csv")
```








