---
title: "HAT Monitoring Spatial Data"
author: "Wendy Anthony for HAT"
date: "23/02/2026"
output:
  html_document:
    highlight: zenburn
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
    # highlight: textmate
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: '3'
# word_document:
  # mainfont: "Arial"      # Your main font (serif or sans-serif)
  # sansfont: "Helvetica"  # The specific sans-serif font for sans-serif contexts
  # fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load Libraries
```{r load-libraries}
library(htmlwidgets) # to save leaflet map as html
```



# Data Import 
- st_read()  
```{r data-import-st-data-download, include = FALSE, cache = TRUE}
## Import spatial Data
OH_2024_ShinyG_Burning_Priorities <- st_read("data/data_spatial/OH_2024_ShinyG_Burning_Priorities/OH_2024_ShinyG_Burning_Priorities.shp")
OH_2024_Invasive_Species_Removed <- st_read("data/data_spatial/OH_2024_Invasive_Species_Removed/OH_2024_Invasive_Species_Removed.shp")
OH_2024_Identified_Invasive_Species_Patches <- st_read("data/data_spatial/OH_2024_Identified_Invasive_Species_Patches/OH_2024_Identified_InvasiveSpecies_Patches.shp")
OH_2024_Invasive_Species_Points <- st_read("data/data_spatial/OH_2024_Invasive_Species_Points/OH_2024_Invasive_Species_Points.shp")
OHP_2025_Completed_polygons <- st_read("data/data_spatial/OHP_2025_Completed_polygons/OHP_2025_Completed polygons.shp")
OHP_2025_Experimental_test_plots <- st_read("data/data_spatial/OHP_2025_Experimental_test_plots/OHP_2025_Experimental test plots.shp")
```

## Drop Z Dimension to map in R
- st_zm()  
```{r shape-data-dropz, include = FALSE, cache = TRUE}
## Drop Z Dimension of Spatial Data to map
# https://gis.stackexchange.com/questions/253898/adding-a-linestring-by-st-read-in-shiny-leaflet
# When trying to create leaflet map Error in if (length(nms) != n || any(nms == "")) stop("'options' must be a fully named list, or have no names (NULL)") : missing value where TRUE/FALSE needed

## Drop Z Dimension of Spatial Data to map
OH_2024_ShinyG_Burning_Priorities_dropz <- st_zm(OH_2024_ShinyG_Burning_Priorities, drop = T, what = "ZM")
OH_2024_Invasive_Species_Removed_dropz <- st_zm(OH_2024_Invasive_Species_Removed, drop = T, what = "ZM")
OH_2024_Identified_Invasive_Species_Patches_dropz <- st_zm(OH_2024_Identified_Invasive_Species_Patches, drop = T, what = "ZM")
OH_2024_Invasive_Species_Points_dropz <- st_zm(OH_2024_Invasive_Species_Points, drop = T, what = "ZM")
OHP_2025_Completed_polygons_dropz <- st_zm(OHP_2025_Completed_polygons, drop = T, what = "ZM")
OHP_2025_Experimental_test_plots_dropz <- st_zm(OHP_2025_Experimental_test_plots, drop = T, what = "ZM")
```

# Subset Polygon Data
- rename Site  
- reorder columns

## OH_2024_ShinyG_Burning_Priorities
```{r spatial-subset-save-ShinyG_Burning_Priorities, include = FALSE, cache = TRUE}
### Subset and save spatial data for use in datatable
# add new columns "Site"   "Year"      "Data"

#################################
colnames(OH_2024_ShinyG_Burning_Priorities_dropz)
# [1] "Name"       "FolderPath" "SymbolID"   "AltMode"    "Base"       "Clamped"    "Extruded"   "Snippet"    "PopupInfo"  "Shape_Leng"
# [11] "Shape_Area" "INSVNAME"   "AreaNumber" "Priority"   "geometry"
OH_2024_ShinyG_Burning_Priorities_dropz$Site <- "Oak Haven"
OH_2024_ShinyG_Burning_Priorities_dropz$Year <- "2024"
OH_2024_ShinyG_Burning_Priorities_dropz$Data <- "OH_2024_ShinyG_Burning_Priorities"
ncol(OH_2024_ShinyG_Burning_Priorities_dropz)

# reorder names for priority so polygon data can be bound by row
OH_2024_ShinyG_Burning_Priorities_dropz <- OH_2024_ShinyG_Burning_Priorities_dropz[, c("Name", "FolderPath", "SymbolID", "AltMode", "Base", "Clamped", "Extruded", "Snippet", "PopupInfo", "Shape_Leng", "Shape_Area", "INSVNAME", "AreaNumber", "geometry", "Site", "Year", "Data", "Priority" )]
colnames(OH_2024_ShinyG_Burning_Priorities_dropz)
ncol(OH_2024_ShinyG_Burning_Priorities_dropz)
nrow(OH_2024_ShinyG_Burning_Priorities_dropz) #13
#  [1] "Name"       "FolderPath" "SymbolID"   "AltMode"    "Base"       "Clamped"    "Extruded"   "Snippet"    "PopupInfo"  "Shape_Leng"
# [11] "Shape_Area" "INSVNAME"   "AreaNumber" "geometry"   "Site"       "Year"       "Data"       "Priority"

write.csv(OH_2024_ShinyG_Burning_Priorities_dropz, "data/OH_2024_ShinyG_Burning_Priorities_dropz.csv")
```

## OH_2024_Invasive_Species_Removed
```{r spatial-subset-save-OH_2024_Invasive_Species_Removed, include = FALSE, cache = TRUE}
#################################
colnames(OH_2024_Invasive_Species_Removed_dropz)
# [1] "Name"       "FolderPath" "SymbolID"   "AltMode"    "Base"       "Clamped"    "Extruded"   "Snippet"    "PopupInfo"  "Shape_Leng"
# [11] "Shape_Area" "INSVNAME"   "geometry"   "Site"       "Year"       "Data"       "Priority"
OH_2024_Invasive_Species_Removed_dropz$Site <- "Oak Haven"
OH_2024_Invasive_Species_Removed_dropz$Year <- "2024"
OH_2024_Invasive_Species_Removed_dropz$Data <- "OH_2024_Invasive_Species_Removed"
OH_2024_Invasive_Species_Removed_dropz$AreaNumber <- ""
OH_2024_Invasive_Species_Removed_dropz$Priority <- "Priority"
# reorder names for priority so polygon data can be bound by row
OH_2024_Invasive_Species_Removed_dropz <- OH_2024_Invasive_Species_Removed_dropz[, c("Name", "FolderPath", "SymbolID", "AltMode", "Base", "Clamped", "Extruded", "Snippet", "PopupInfo", "Shape_Leng", "Shape_Area", "INSVNAME", "AreaNumber", "geometry", "Site", "Year", "Data", "Priority" )]
ncol(OH_2024_Invasive_Species_Removed_dropz)
nrow(OH_2024_Invasive_Species_Removed_dropz) #10

write.csv(OH_2024_Invasive_Species_Removed_dropz, "data/OH_2024_Invasive_Species_Removed_dropz.csv")
```

## OH_2024_Identified_Invasive_Species_Patches
```{r spatial-subset-save-OH_2024_Identified_Invasive_Species_Patches, include = FALSE, cache = TRUE}
#################################
colnames(OH_2024_Identified_Invasive_Species_Patches_dropz)
# [1] "Name"       "FolderPath" "SymbolID"   "AltMode"    "Base"       "Clamped"    "Extruded"   "Snippet"    "PopupInfo"  "Shape_Leng"
# [11] "Shape_Area" "INSVNAME"   "AreaNumber" "geometry"
OH_2024_Identified_Invasive_Species_Patches_dropz$Site <- "Oak Haven"
OH_2024_Identified_Invasive_Species_Patches_dropz$Year <- "2024"
OH_2024_Identified_Invasive_Species_Patches_dropz$Data <- "OH_2024_Identified_Invasive_Species_Patches"
OH_2024_Identified_Invasive_Species_Patches_dropz$Priority <- "Priority"
ncol(OH_2024_Identified_Invasive_Species_Patches_dropz)
nrow(OH_2024_Identified_Invasive_Species_Patches_dropz) #13

write.csv(OH_2024_Identified_Invasive_Species_Patches_dropz, "data/OH_2024_Identified_Invasive_Species_Patches_dropz.csv")
```

## OHP_2025_Completed_polygons
```{r spatial-subset-save-OHP_2025_Completed_polygons, include = FALSE, cache = TRUE}

#################################
colnames(OHP_2025_Completed_polygons_dropz)
 #[1] "Name"       "FolderPath" "SymbolID"   "AltMode"    "Base"       "Clamped"    "Extruded"   "Snippet"    "PopupInfo"  "Shape_Leng"
#[11] "Shape_Area" "geometry"   "Site"       "Year"       "Data"       "Priority" "
OHP_2025_Completed_polygons_dropz$Site <- "Oak Haven"
OHP_2025_Completed_polygons_dropz$Year <- "2025"
OHP_2025_Completed_polygons_dropz$Data <- "OHP_2025_Completed_polygons"
OHP_2025_Completed_polygons_dropz$Priority <- "Priority"
OHP_2025_Completed_polygons_dropz$INSVNAME <- ""
OHP_2025_Completed_polygons_dropz$AreaNumber <- ""
ncol(OHP_2025_Completed_polygons_dropz)
nrow(OHP_2025_Completed_polygons_dropz) #7
# reorder names for priority so polygon data can be bound by row
OHP_2025_Completed_polygons_dropz <- OHP_2025_Completed_polygons_dropz[, c("Name", "FolderPath", "SymbolID", "AltMode", "Base", "Clamped", "Extruded", "Snippet", "PopupInfo", "Shape_Leng", "Shape_Area", "INSVNAME", "AreaNumber", "geometry", "Site", "Year", "Data", "Priority" )]

write.csv(OHP_2025_Completed_polygons_dropz, "data/OHP_2025_Completed_polygons_dropz.csv")
```

## Bind spatial polygons for Datatable Display
```{r rbind-spatial-polygons, include = FALSE, cache = TRUE}
#################################
# rbind polygons
OH_spatial_polygons_all <- rbind(OH_2024_ShinyG_Burning_Priorities_dropz, OH_2024_Invasive_Species_Removed_dropz, OH_2024_Identified_Invasive_Species_Patches_dropz, OHP_2025_Completed_polygons_dropz)
nrow(OH_spatial_polygons_all) #43

colnames(OH_spatial_polygons_all)
# remove unused columns
OH_spatial_polygons_all_sub <- subset(OH_spatial_polygons_all, select = -c(2:8, 10, 11))
colnames(OH_spatial_polygons_all_sub)

write.csv(OH_spatial_polygons_all_sub, "data/OH_spatial_polygons_all_sub.csv")
```


# Subset Point Data
- rename Site  
- reorder columns

# OH_2024_Invasive_Species_Points
```{r spatial-subset-save-OH_2024_Invasive_Species_Points, include = FALSE, cache = TRUE}
#################################
#################################

colnames(OH_2024_Invasive_Species_Points_dropz)
# [1] "Id"        "Name"      "Descript"  "Type"      "Comment"   "Symbol"    "DateTimeS" "Elevation" "INSVNAME"  "geometry"
OH_2024_Invasive_Species_Points_dropz$Site <- "Oak Haven"
OH_2024_Invasive_Species_Points_dropz$Year <- "2024"
OH_2024_Invasive_Species_Points_dropz$Data <- "OH_2024_Invasive_Species_Points"

write.csv(OH_2024_Invasive_Species_Points_dropz, "data/OH_2024_Invasive_Species_Points_dropz.csv")
```

# OHP_2025_Experimental_test_plots
```{r spatial-subset-save-OHP_2025_Experimental_test_plots, include = FALSE, cache = TRUE}
#################################
colnames(OHP_2025_Experimental_test_plots_dropz)
# [1] "Id"        "Name"      "Descript"  "Type"      "Comment"   "Symbol"    "DateTimeS" "Elevation" "geometry"
OHP_2025_Experimental_test_plots_dropz$Site <- "Oak Haven"
OHP_2025_Experimental_test_plots_dropz$Year <- "2025"
OHP_2025_Experimental_test_plots_dropz$Data <- "OHP_2025_Experimental_test_plots"
OHP_2025_Experimental_test_plots_dropz$INSVNAME <- ""

# reorder columns
OHP_2025_Experimental_test_plots_dropz <- OHP_2025_Experimental_test_plots_dropz[, c("Id", "Name", "Descript", "Type", "Comment", "Symbol", "DateTimeS", "Elevation", "INSVNAME", "geometry", "Site", "Year", "Data")]

colnames(OHP_2025_Experimental_test_plots_dropz)
colnames(OH_2024_Invasive_Species_Points_dropz)

write.csv(OHP_2025_Experimental_test_plots_dropz, "data/OHP_2025_Experimental_test_plots_dropz.csv")
```

## Bind spatial points for Datatable Display
```{r rbind-spatial-points, include = FALSE, cache = TRUE}
#################################
# rbind points
OH_spatial_points_all <- rbind(OHP_2025_Experimental_test_plots_dropz, OH_2024_Invasive_Species_Points_dropz)
colnames(OH_spatial_points_all)
OH_spatial_points_all_sub <- subset(OH_spatial_points_all, select = -c(1,4,5,6))
colnames(OH_spatial_points_all_sub)

write.csv(OH_spatial_points_all_sub, "data/OH_spatial_points_all_sub.csv")
```

# Leaflet Map with Spatial Layers
## Oak Haven Spatial Data Map
```{r leaflet-OH-spatial, echo = FALSE, message = FALSE, include = TRUE}
title <- '<p style="text-align: center; height: 12px;"><span style="font-size:12px;font-weight:bold; background-color: rgba(255, 255, 255, 0.9;");>HAT GOE Monitoring Covenant Sites Map</span></p>'

map_OH_sp <- leaflet() %>%
  # addTiles() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(data = OHP_2025_Completed_polygons_dropz,
              group = "OH 2025 Completed polygons",
              color = "#AA4499", #AA4499 purple
              #fillColor = "purple",
              fillOpacity = 0.5,
              popup = paste0("<b>Name: </b>", OHP_2025_Completed_polygons_dropz$Name, "<br>",
                             "<b>Description: </b>", OHP_2025_Completed_polygons_dropz$PopupInfo)) %>%
  addCircleMarkers(data = OHP_2025_Experimental_test_plots_dropz,
              group = "OH 2025 Experimental test plot points",
              color = "#EECC66", #EECC66 light yellow
              #fillColor = "purple",
              fillOpacity = 0.8,
              weight = 1,
              radius = 3,
              popup = paste0("<b>Name: </b>", OHP_2025_Experimental_test_plots_dropz$Name, "<br>",
                            "<b>Description: </b>", OHP_2025_Experimental_test_plots_dropz$Descript, "<br>",
                            "<b>Date: </b>", OHP_2025_Experimental_test_plots_dropz$DateTimeS)) %>%
  addPolygons(data = OH_2024_ShinyG_Burning_Priorities_dropz,
              # covers same area as Invasive_Species_Patches
              group = "OH 2024 ShinyG Burning Priorities",
              # use a lighter colour so lighter Shiny will show over Species patches
              color = "#EE7733", #EE7733 orange
              #color = "blue",
              #fillColor = "blue",
              fillOpacity = 0.3,
              popup = paste0("<b>Name: </b>", OH_2024_ShinyG_Burning_Priorities_dropz$Name, "<br>",
                             "<b>Priority: </b>", OH_2024_ShinyG_Burning_Priorities_dropz$Priority, "<br>",
                             "<b>Description: </b>", OH_2024_ShinyG_Burning_Priorities_dropz$PopupInfo, "<br>",
                             "<b>Area Number: </b>", OH_2024_ShinyG_Burning_Priorities_dropz$AreaNumber, "<br>",
                             "<b>Species: </b>", OH_2024_ShinyG_Burning_Priorities_dropz$INSVNAME)) %>% # Optional: add popups
  addPolygons(data = OH_2024_Invasive_Species_Removed_dropz,
              group = "OH 2024 Invasive Species Removed",
              color = "#EE3377", # magenta #EE3377
              #color = "yellow",
              #fillColor = "yellow",
              fillOpacity = 0.6,
              popup = paste0("<b>Name: </b>", OH_2024_Invasive_Species_Removed_dropz$Name, "<br>",
                             "<b>Description: </b>", OH_2024_Invasive_Species_Removed_dropz$PopupInfo)) %>%
  addPolygons(data = OH_2024_Identified_Invasive_Species_Patches_dropz,
              # covers same area as Shiny g
              group = "OH 2024 Invasive Species Patches",
              # use a darker colour so lighter Shiny will show
              color = "#33BBEE", #cyan #33BBEE
              #fillColor = "orange",
              fillOpacity = 0.4,
              popup = paste0("<b>Name: </b>", OH_2024_Identified_Invasive_Species_Patches_dropz$Name, "<br>",
                             "<b>Description: </b>", OH_2024_Identified_Invasive_Species_Patches_dropz$PopupInfo, "<br>",
                             "<b>Area: </b>", OH_2024_Identified_Invasive_Species_Patches_dropz$AreaNumber, "<br>",
                             "<b>Species: </b>", OH_2024_Identified_Invasive_Species_Patches_dropz$INSVNAME)
              ) %>%
  addCircleMarkers(data = OH_2024_Invasive_Species_Points_dropz,
              group = "OH 2024 Invasive Species Points",
              color = "#CC3311", # CC3311 red
              #fillColor = "red",
              fillOpacity = 0.5,
              radius = 3,
              popup = paste0("<b>Name: </b>", OH_2024_Invasive_Species_Points_dropz$Name, "<br>",
                            "<b>Description: </b>", OH_2024_Invasive_Species_Points_dropz$Descript, "<br>",
                            "<b>Species: </b>", OH_2024_Invasive_Species_Points_dropz$INSVNAME, "<br>",
                            "<b>Elevation: </b>", OH_2024_Invasive_Species_Points_dropz$Elevation, "<br>",
                            "<b>Date: </b>", OH_2024_Invasive_Species_Points_dropz$DateTimeS)
              ) %>%
  # add legend to overlay group toggled off
  addLegend(
              "bottomright",
              colors = c("#AA4499", "#EECC66", "#EE3377", "#33BBEE", "#EE7733", "#CC3311"),
              labels = c("OH 2025 Completed polygons", "OH 2025 Experimental test plot points", "OH 2024 Invasive Species Removed", "OH 2024 Invasive Species Patches", "OH 2024 ShinyG Burning Priorities", "OH 2024 Invasive Species Points"),
              title = "OH Spatial Data",
              group = "OH Spatial Polygon Color Legend"
              ) %>%
  setView(-123.4471, 48.56419, 17) %>%
  addControl(title, position = "topright") %>% 
  addMeasure(
    position = "topleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#DDAA33",
    completedColor = "#BB5566") %>% 
  addLayersControl(
    overlayGroups = c("OH 2025 Completed polygons", "OH 2025 Experimental test plot points", "OH 2024 Invasive Species Removed", "OH 2024 Invasive Species Patches", "OH 2024 ShinyG Burning Priorities", "OH 2024 Invasive Species Points", "OH Spatial Polygon Color Legend"),
    options = layersControlOptions(collapsed = TRUE) # Makes control box always visible
  ) %>%
  hideGroup(c("OH 2024 ShinyG Burning Priorities", "OH 2024 Invasive Species Removed", "OH 2024 Invasive Species Patches", "OH 2024 Invasive Species Points", "OH Spatial Polygon Color Legend")) %>%
  addScaleBar("topleft")  
# %>%
#       # add controls
#        addMiniMap(width = 100, height = 100, zoomLevelOffset = -4)

# Print the map
map_OH_sp

# Save map as webpage
saveWidget(map_OH_sp, "output/HAT_spatial_site_map.html")

```

### Geolocate Oak Haven Park
```{r leaflet-OH-spatial_geoloc, echo = FALSE, message = FALSE, include = TRUE}
################################################################################
# Interactive Geolocation map Developed by Wendy Anthony for HAT 2026-03-03
################################################################################
library(leaflet.extras) # for addControlGPS

title <- '<p style="text-align: center; height: 12px;"><span style="font-size:12px;font-weight:bold; background-color: rgba(255, 255, 255, 0.9;");>HAT GOE Monitoring Covenant Sites Map</span></p>'

map_OH_sp_geoloc <- leaflet() %>%
  # addTiles() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(data = OHP_2025_Completed_polygons_dropz,
              group = "OH 2025 Completed polygons",
              color = "#AA4499", #AA4499 purple
              #fillColor = "purple",
              fillOpacity = 0.5,
              popup = paste0("<b>Name: </b>", OHP_2025_Completed_polygons_dropz$Name, "<br>",
                             "<b>Description: </b>", OHP_2025_Completed_polygons_dropz$PopupInfo)) %>%
  addCircleMarkers(data = OHP_2025_Experimental_test_plots_dropz,
              group = "OH 2025 Experimental test plot points",
              color = "#EECC66", #EECC66 light yellow
              #fillColor = "purple",
              fillOpacity = 0.8,
              weight = 1,
              radius = 3,
              popup = paste0("<b>Name: </b>", OHP_2025_Experimental_test_plots_dropz$Name, "<br>",
                            "<b>Description: </b>", OHP_2025_Experimental_test_plots_dropz$Descript, "<br>",
                            "<b>Date: </b>", OHP_2025_Experimental_test_plots_dropz$DateTimeS)) %>%
  addPolygons(data = OH_2024_ShinyG_Burning_Priorities_dropz,
              # covers same area as Invasive_Species_Patches
              group = "OH 2024 ShinyG Burning Priorities",
              # use a lighter colour so lighter Shiny will show over Species patches
              color = "#EE7733", #EE7733 orange
              #color = "blue",
              #fillColor = "blue",
              fillOpacity = 0.3,
              popup = paste0("<b>Name: </b>", OH_2024_ShinyG_Burning_Priorities_dropz$Name, "<br>",
                             "<b>Priority: </b>", OH_2024_ShinyG_Burning_Priorities_dropz$Priority, "<br>",
                             "<b>Description: </b>", OH_2024_ShinyG_Burning_Priorities_dropz$PopupInfo, "<br>",
                             "<b>Area Number: </b>", OH_2024_ShinyG_Burning_Priorities_dropz$AreaNumber, "<br>",
                             "<b>Species: </b>", OH_2024_ShinyG_Burning_Priorities_dropz$INSVNAME)) %>% # Optional: add popups
  addPolygons(data = OH_2024_Invasive_Species_Removed_dropz,
              group = "OH 2024 Invasive Species Removed",
              color = "#EE3377", # magenta #EE3377
              #color = "yellow",
              #fillColor = "yellow",
              fillOpacity = 0.6,
              popup = paste0("<b>Name: </b>", OH_2024_Invasive_Species_Removed_dropz$Name, "<br>",
                             "<b>Description: </b>", OH_2024_Invasive_Species_Removed_dropz$PopupInfo)) %>%
  addPolygons(data = OH_2024_Identified_Invasive_Species_Patches_dropz,
              # covers same area as Shiny g
              group = "OH 2024 Invasive Species Patches",
              # use a darker colour so lighter Shiny will show
              color = "#33BBEE", #cyan #33BBEE
              #fillColor = "orange",
              fillOpacity = 0.4,
              popup = paste0("<b>Name: </b>", OH_2024_Identified_Invasive_Species_Patches_dropz$Name, "<br>",
                             "<b>Description: </b>", OH_2024_Identified_Invasive_Species_Patches_dropz$PopupInfo, "<br>",
                             "<b>Area: </b>", OH_2024_Identified_Invasive_Species_Patches_dropz$AreaNumber, "<br>",
                             "<b>Species: </b>", OH_2024_Identified_Invasive_Species_Patches_dropz$INSVNAME)
              ) %>%
  addCircleMarkers(data = OH_2024_Invasive_Species_Points_dropz,
              group = "OH 2024 Invasive Species Points",
              color = "#CC3311", # CC3311 red
              #fillColor = "red",
              fillOpacity = 0.5,
              radius = 3,
              popup = paste0("<b>Name: </b>", OH_2024_Invasive_Species_Points_dropz$Name, "<br>",
                            "<b>Description: </b>", OH_2024_Invasive_Species_Points_dropz$Descript, "<br>",
                            "<b>Species: </b>", OH_2024_Invasive_Species_Points_dropz$INSVNAME, "<br>",
                            "<b>Elevation: </b>", OH_2024_Invasive_Species_Points_dropz$Elevation, "<br>",
                            "<b>Date: </b>", OH_2024_Invasive_Species_Points_dropz$DateTimeS)
              ) %>%
  # add legend to overlay group toggled off
  addLegend(
              "bottomright",
              colors = c("#AA4499", "#EECC66", "#EE3377", "#33BBEE", "#EE7733", "#CC3311"),
              labels = c("OH 2025 Completed polygons", "OH 2025 Experimental test plot points", "OH 2024 Invasive Species Removed", "OH 2024 Invasive Species Patches", "OH 2024 ShinyG Burning Priorities", "OH 2024 Invasive Species Points"),
              title = "OH Spatial Data",
              group = "OH Spatial Polygon Color Legend"
              ) %>%
  setView(-123.4471, 48.56419, 17) %>%
    # leaflet.extras
  addControlGPS(
    options = gpsOptions(
      position = "topleft",
      activate = TRUE,
      autoCenter = TRUE,
      maxZoom = 10,
      setView = TRUE
    )
  ) %>%
  addControl(title, position = "topright") %>% 
  addMeasure(
    position = "topleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#DDAA33",
    completedColor = "#BB5566")  %>%
  hideGroup(c("OH 2024 ShinyG Burning Priorities", "OH 2024 Invasive Species Removed", "OH 2024 Invasive Species Patches", "OH 2024 Invasive Species Points", "OH Spatial Polygon Color Legend")) %>%
  addScaleBar("topleft")  %>%
  addDrawToolbar(
    targetGroup = "Points",
    markerOptions = TRUE,
    polylineOptions = FALSE,
    polygonOptions = FALSE,
    circleOptions = FALSE,
    rectangleOptions = FALSE,
    circleMarkerOptions = FALSE,
    editOptions = editToolbarOptions(selectedPathOptions = selectedPathOptions())
  ) %>%
  ##### this layer.getLatLng() function only works for points
   onRender("
    function(el, x) {
      this.on('draw:created', function(e) {
        var layer = e.layer;
        var coords = layer.getLatLng();
        console.log('Lat: ' + coords.lat + ', Lng: ' + coords.lng);
        // Add to a global variable or custom HTML element here
        alert('Point Saved: ' + coords.lat + ', ' + coords.lng);
      });
    }
    ") %>% 
    addLayersControl(
      overlayGroups = c("OH 2025 Completed polygons", "OH 2025 Experimental test plot points", "OH 2024 Invasive Species Removed", "OH 2024 Invasive Species Patches", "OH 2024 ShinyG Burning Priorities", "OH 2024 Invasive Species Points", "OH Spatial Polygon Color Legend", "Points"),
      options = layersControlOptions(collapsed = TRUE) # Makes control box always visible
    ) 

################################################################################
# lat long on click
# these get in the way of drawing polygons
# js_code <- "
# function(el, x) {
#   this.on('click', function(e) {
#     var coords = e.latlng;
#     var lat = coords.lat;
#     var lng = coords.lng;
#     var popup = L.popup()
#       .setLatLng(coords)
#       .setContent('Latitude: ' + lat.toFixed(4) + ', Longitude: ' + lng.toFixed(4))
#       .openOn(this);
#   });
# }
# "
# 
# map_OH_sp_geoloc <- map_OH_sp_geoloc %>%
#   onRender(js_code)

map_OH_sp_geoloc

# Save map as webpage
saveWidget(map_OH_sp_geoloc, "output/HAT_spatial_site_map_geoloc_draw.html")

```


